### 2022-01-20 replication_data/rmd/02_format_and_clean.rmd

---
title: "02_format_and_clean"
output:
  html_document:
    df_print: paged
  toc: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

`%>%` <- magrittr::`%>%`
```

# Load and transform wide to long

read in icb_coder_crisis_sentence_event_aligned.Rds
```{r}
icb_wide_aligned_raw <- readRDS(file=paste0(here::here(),"/replication_data/out/icb_coder_crisis_sentence_event_aligned.Rds")) %>% 
                    dplyr::mutate(crisno3 = stringr::str_pad(crisno, 3, pad = "0"), sentence_number_int_aligned3 =
                                    stringr::str_pad(sentence_number_int_aligned, 3, pad = "0")) %>%
                    dplyr::mutate(email_crisis_sent_event =
                                    paste0(email_id,"_",crisno3,"_",sentence_number_int_aligned3,"_",event_number_int)) %>%
                    dplyr::select(-crisno3, -sentence_number_int_aligned3) %>%
                    dplyr::arrange(crisno, sentence_number_int, event_number_int) %>%

                    ####################Where we kill unhelpful events
                    #There are 2 legitimate kinds of rows
                    dplyr::filter(
                      event_number_int==0 & sentence_events %in% c('Background', 'Covered by Previous Sentence','Error','No Action') |
                      event_number_int>0 & event_type %in% c('Action', 'Speech','Thought' )
                    ) #%>%

dim(icb_wide_aligned_raw) #87710    85
```

Remove variables used earlier for compiling the individual codings
```{r}
icb_wide_aligned <- icb_wide_aligned_raw %>% dplyr::ungroup() %>% 
  dplyr::select(
               # -lastviewed, 
               # -input_crisis, 
                -sentence_number_int, -raw_crisis, -raw_crisno, -raw_sentence, -savefile_crc32, -raw_crisno_sentno, -raw_crisno_version_sentno, -crisisno_crisis_version) %>%
  dplyr::mutate(icb_survey_version = ifelse(is.na(icb_survey_version), 1.1, icb_survey_version))
dim(icb_wide_aligned) #52929    78

```

Initial coding edits (spelling, formatting, etc)
```{r}
icb_wide_aligned <- icb_wide_aligned %>% 
                      dplyr::rename(act_deescalate = act_descalate) %>% 
                      dplyr::rename(interact_deescalate = interact_descalate) %>% 
                      dplyr::rename(condition_act_deescalate = condition_act_descalate) %>% 
                      dplyr::rename(condition_interact_deescalate = condition_interact_descalate) %>% 
                      dplyr::mutate(sayintkind_react = ifelse(sayintkind_react=="This Sentence", raw_sentence_number_int,
                                                                         stringr::str_replace(sayintkind_react, "Sentence ","")))
dim(icb_wide_aligned) #52929    78

```

Where a coder looked at the same crisis twice, drop the earlier version and the earlier date
```{r}
icb_wide_aligned <- icb_wide_aligned %>% 
                      dplyr::group_by(email_crisis_sent_event) %>% 
                      dplyr::slice_max(icb_survey_version) %>%
                      dplyr::ungroup()

icb_wide_aligned <- icb_wide_aligned %>% 
                    dplyr::group_by(email_crisis_sent_event) %>% 
                    dplyr::slice_max(lastviewed) %>%
                    dplyr::ungroup()
dim(icb_wide_aligned) #87710    77
```

Drop a small subset where it appears that a coder looked at the same crisis twice but one of the times was logged incorrectly
```{r}
icb_wide_aligned <- icb_wide_aligned %>% dplyr::left_join(
                    icb_wide_aligned %>% dplyr::select(email_crisis_sent_event, event_number_int) %>% dplyr::filter(
                    (icb_wide_aligned %>% dplyr::select(email_crisis_sent_event, event_number_int) %>% 
                       duplicated()) == TRUE) %>% 
                      dplyr::mutate(duplicate = 1)) %>% 
                    dplyr::mutate(mismatch = crisis != input_crisis) %>% 
                    dplyr::filter(!(mismatch %in% TRUE & duplicate %in% 1)) %>%
                    dplyr::select(-mismatch, -duplicate)

dim(icb_wide_aligned) #51098    78

```

Pivot to long form 
```{r}

icb_long <- icb_wide_aligned %>% dplyr::ungroup() %>% 
             dplyr::relocate(sentence_events, .after = event_number_int) %>%
             tidyr::pivot_longer(
               cols = c(sentence_events:interact_units), #  changed this to be inclusive of event type and sentence_events 
               names_to = "variable",
               #names_prefix = "wk",
               values_to = "value",
               values_drop_na = TRUE
             ) %>% dplyr::ungroup() %>% 
  dplyr::mutate(value=value %>% tolower %>% trimws())
dim(icb_long) #491757     15

```

Now split values with a list (separated by a semicolon)

```{r}

icb_long_split <- icb_long %>% 
                  dplyr::mutate(value = strsplit(as.character(value), ";")) %>% 
                  tidyr::unnest(value) %>%
                  dplyr::mutate(value=value %>% tolower %>% trimws())
dim(icb_long_split) #536397     15

```

# Define subsets of variables

```{r}

vars <- c(
"sentence_events", "event_type", "raterconfidence", "raterconfidence_reason",
"raterconfidence_reason_survey", "date", "think_sentence_events", "think_actor_a", "think_actor_a_other", 
"thinkkind", "say_actor_a", "say_actor_a_other", "say_actor_b", "say_actor_b_other",
"sayintkind", "sayintkind_react", "consequence", "condition", "do_actor_a",
 "do_actor_a_other", "do_actor_b", "do_actor_b_other", "condition_do_actor_a",
"condition_do_actor_a_other", "condition_do_actor_b", "condition_do_actor_b_other",
 "condition_act_cooperative", "condition_act_deescalate", "condition_act_escalate",
"condition_act_uncooperative", "condition_do_interact_kind", "condition_do_kind",
"condition_interact_decreasecoop",  "condition_interact_deescalate",
"condition_interact_domains", "condition_interact_escalate", "condition_interact_fatalities",
"condition_interact_forces", "condition_interact_geoscope", "condition_interact_increasecoop",
"condition_interact_location", "condition_interact_location_other", "condition_interact_territory",
"condition_interact_units", "do_duration", "do_interact_kind", "do_kind", "do_timing",
"do_timing_reorder", "act_cooperative", "act_deescalate", "act_escalate",
"act_uncooperative", "interact_decreasecoop", "interact_deescalate", "interact_domains",
"interact_escalate", "interact_fatalities", "interact_forces", "interact_geoscope",
"interact_increasecoop",  "interact_location", "interact_location_other", "interact_territory",
"interact_units" )

vars_actor <- c("think_actor_a",
             "think_actor_a_other",
             "say_actor_a",
             "say_actor_a_other",
             "say_actor_b",
             "say_actor_b_other",
             "do_actor_a",
             "do_actor_a_other",
             "do_actor_b",
             "do_actor_b_other", 
             "condition_do_actor_a",
             "condition_do_actor_a_other",
             "condition_do_actor_b",
             "condition_do_actor_b_other")

vars_act <- c("act_cooperative",
              "act_uncooperative",
              "act_deescalate",
              "act_escalate",
              "interact_decreasecoop",
              "interact_increasecoop",
              "interact_deescalate",
              "interact_escalate",
              "condition_act_cooperative",
              "condition_act_uncooperative",
              "condition_act_deescalate",
              "condition_act_escalate",
              "condition_interact_decreasecoop",
              "condition_interact_increasecoop",
              "condition_interact_deescalate",
              "condition_interact_escalate")

# saveRDS(vars_actor, paste0(here::here(), "/replication_data/temp/","vars_actor.Rds"))
# saveRDS(vars_act, paste0(here::here(), "/replication_data/temp/","vars_act.Rds"))

vars_say <- c("sayintkind", "sayintkind_react","consequence","condition")
vars_think <- c("think_sentence_events", "thinkkind", "thinkkind_clean")

# Duplicate variables, same options for each in these types
vars_act_kind <- c("do_kind", "condition_do_kind", "do_interact_kind", "condition_do_interact_kind")
vars_fatalities <- c("interact_fatalities", "condition_interact_fatalities")
vars_forces <- c("interact_forces", "condition_interact_forces")
vars_geoscope <- c("interact_geoscope", "condition_interact_geoscope")
vars_location <- c("interact_location", "condition_interact_location")
vars_location_other <- c("interact_location_other", "condition_interact_location_other")
vars_territory <- c("interact_territory", "condition_interact_territory")

# NEED TO ADD TO CODEBOOK
vars_domains <- c("interact_domains", "condition_interact_domains")
vars_units <- c("interact_units", "condition_interact_units")

```

```{r}
# are all variables in vars?
setdiff(unique(icb_long_split$variable), vars)

# do the subsets cover everything in vars?
setdiff(vars, c(vars_actor, vars_act, vars_act_kind, vars_fatalities, vars_forces, vars_geoscope, vars_location, vars_location_other, vars_territory, vars_domains, vars_units, vars_say, vars_think))
```

# Load and apply normalizations

Load normalization dictionaries

```{r message=FALSE}
target_file <- paste0(here::here(),"/replication_data/in/icb_manual_recoding_master_sheet.xlsx")
downloaded_file <- googledrive::drive_download(file="https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", path=target_file, type = "xlsx", overwrite=T)
# in not working, run googledrive::drive_auth() and select new token

dictionary_actions  <- readxl::read_excel(target_file, sheet="actions") %>% dplyr::distinct()
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno,value, value_normalized=value_normalized_wikidata_id) %>% dplyr::distinct()

#Locations and dates are both really hard to do
dictionary_locations <- readxl::read_excel(target_file, sheet="locations", na = c("", "NA")) 
dictionary_dates <- readxl::read_excel(target_file, sheet="dates", col_types="text") #This is the tricky one and why he had to add values

```

```{r}
# Create dateset to reconnect location dictionary to individual rows through 'sentence_number_int_aligned'
location_merge_directory <- icb_wide_aligned %>% 
    dplyr::mutate(crisno3 = stringr::str_pad(crisno, 3, pad = "0"), 
                  sentence_number_int_aligned3 = stringr::str_pad(sentence_number_int_aligned, 3, pad = "0")) %>% 
  dplyr::mutate(email_crisis_sent_event = paste(email_id,crisno3,sentence_number_int_aligned3,event_number_int, sep="_")) %>%
  dplyr::select(-crisno3, -sentence_number_int_aligned3) %>%
  dplyr::mutate(crisis_sent = paste0(crisno,"_",sentence_number_int_aligned)) %>%
  dplyr::select(crisno, email_crisis_sent_event, do_actor_a, do_actor_b, interact_location, interact_location_other) %>% 
  dplyr::mutate(do_actor_a=strsplit(do_actor_a, ";")) %>% tidyr::unnest(do_actor_a) %>%
  dplyr::mutate(do_actor_b=strsplit(do_actor_b, ";")) %>% tidyr::unnest(do_actor_b) %>%
  dplyr::mutate(interact_location_other=strsplit(interact_location_other, ";")) %>% tidyr::unnest(interact_location_other) %>%
  dplyr::mutate(do_actor_a=tolower(do_actor_a), do_actor_b=tolower(do_actor_b),
                interact_location_other=trimws(tolower(interact_location_other))) %>%
  dplyr::left_join(dictionary_locations %>%
                     dplyr::mutate() %>% 
                     dplyr::mutate(interact_location_other = tolower(interact_location_other)) %>% 
                     dplyr::select(do_actor_a, do_actor_b, interact_location, interact_location_other, norm_name)) %>%
  dplyr::mutate(crisno = as.numeric(crisno)) %>% 
  unique() %>% 
  dplyr::filter(!is.na(interact_location_other))

```

Apply the cleaning dictionaries to the relevant subsets

```{r}
  #ACTOR
icb_long_actor_clean <- icb_long_split %>% 
  dplyr::filter(variable %in% vars_actor) %>%
    dplyr::left_join(tibble::tibble(variable=vars_actor) %>% 
                       dplyr::mutate(temp="temp") %>%
                       dplyr::left_join(dictionary_actors %>% dplyr::mutate(temp="temp")) %>%
                       dplyr::select(-temp) %>% 
                       dplyr::select(crisno, variable, value, value_normalized) %>%
                     dplyr::mutate(crisno = as.character(crisno)) %>%
                     # dplyr::mutate(value_normalized = ifelse(is.na(value_normalized), "no_cleaning", value_normalized)) %>% 
                       unique(),
                     by=c('crisno','variable','value')) %>%
    dplyr::mutate(variable_normalized = paste0(variable, "_clean")) %>% 
  dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";")) %>% 
  tidyr::unnest(value_normalized) %>%
  dplyr::mutate(value_normalized=value_normalized %>% trimws())

# check for unmatched
icb_long_actor_clean %>% dplyr::filter(is.na(value_normalized)) %>% dplyr::select(crisno, variable, value)
```

```{r}
  #ACTION
icb_long_action_clean <- icb_long_split %>% 
                        dplyr::filter(variable %in% vars_act) %>% 
                        dplyr::left_join(dplyr::bind_rows(dictionary_actions,
                                         dictionary_actions %>% 
                                           dplyr::mutate(variable = paste0("condition_", variable)) %>% 
                                           dplyr::mutate(variable_normalized = ifelse(is.na(variable_normalized), 
                                                                                            variable_normalized, 
                                                                                            paste0("condition_", variable_normalized)))) %>% 
                                           dplyr::select(variable, value, variable_normalized, value_normalized) %>%
                                           dplyr::mutate(variable_normalized = ifelse(!is.na(value_normalized) & is.na(variable_normalized), 
                                                                                   variable, variable_normalized)),
                                         by=c('variable','value'))

# check for unmatched
icb_long_action_clean %>% dplyr::filter(is.na(value_normalized)) %>% dplyr::select(crisno, variable, value, variable_normalized, value_normalized) 
```

```{r}
 #LOCATION
icb_long_location_clean <- icb_long_split %>% 
  dplyr::filter(variable == "interact_location_other") %>% 
  dplyr::full_join(location_merge_directory %>%
                     dplyr::mutate(crisno = as.character(crisno)) %>% 
                     dplyr::rename(value = interact_location_other)) %>% 
  dplyr::rename(value_normalized = norm_name) %>%
  dplyr::mutate(variable_normalized = "interact_location_other_clean") %>%   
  dplyr::select(-do_actor_a, -do_actor_b, -interact_location) %>% 
  dplyr::mutate(value_normalized = ifelse(is.na(value_normalized), "no_cleaning", value_normalized))
```

```{r}
  #DATES
icb_long_dates_clean <- icb_long_split %>% 
  dplyr::filter(variable=="date") %>% 
  dplyr::left_join(dictionary_dates %>% 
                   dplyr::select(-coder, -notes) %>%
                   dplyr::rename(value=date_orig)) %>%
  tidyr::pivot_longer(cols = point_in_time:latest_date2,
               names_to = "variable_normalized",
               values_to = "value_normalized",
               values_drop_na = TRUE) %>%
  dplyr::mutate(variable_normalized = ifelse(variable_normalized %in% c("point_in_time2", "point_in_time3", "point_in_time4", "point_in_time5"), "point_in_time", variable_normalized))

```

```{r}
# combine cleaned variables

icb_long_clean <- icb_long_split %>% 
  dplyr::left_join(icb_long_action_clean, by=names(icb_long_split)) %>% 
  dplyr::left_join(icb_long_actor_clean, by=names(icb_long_split), suffix = c("", ".actor")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.actor, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.actor, value_normalized))) %>%
      dplyr::select(-variable_normalized.actor, -value_normalized.actor) %>% 
  dplyr::left_join(icb_long_location_clean, by=names(icb_long_split), suffix = c("", ".location")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.location, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.location, value_normalized))) %>%
      dplyr::select(-variable_normalized.location, -value_normalized.location) %>% 
  dplyr::full_join(icb_long_dates_clean, by=names(icb_long_split), suffix = c("", ".date")) %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable_normalized.date, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value_normalized.date, value_normalized))) %>%
      dplyr::select(-variable_normalized.date, -value_normalized.date)

```

```{r}
# As the dictionaries translate some groups into multiple values, need to split out rows with multiple values

icb_long_clean <- icb_long_clean %>% 
  dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";")) %>% 
  tidyr::unnest(value_normalized) %>%
  dplyr::mutate(value_normalized=value_normalized %>% trimws()) %>% 
  dplyr::mutate(crisno=as.numeric(crisno)) %>% 
  dplyr::mutate(value=tolower(value))

dim(icb_long_clean) #507699
```

```{r}
# Fix several typos:

print(paste("change 'premption' to 'preemption'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "premption")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "premption", "preemption"))

print(paste("change 'disaprove' to 'disapprove'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "disaprove")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "disaprove", "disapprove"))

print(paste("change 'uncoperative' to 'uncooperative'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "uncoperative")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "uncoperative", "uncooperative"))

print(paste("change 'capitol' to 'capital'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "capitol")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "capitol", "capital"))

print(paste("change 'survailance' to 'surveillance'", icb_long_clean %>% dplyr::filter(stringr::str_detect(value, "survailance")==TRUE) %>% nrow(), "times."))
icb_long_clean <- icb_long_clean %>% dplyr::mutate(value = stringr::str_replace(value, "survailance", "surveillance"))

```

Create variable "preceding_event"
Create a lead / lag that points to the previous event by group

Create variable "previous_event_timing"
```{r}
alignment_key <- icb_wide_aligned_raw %>% 
  dplyr::ungroup() %>% 
  dplyr::select(email_id, crisno, sentence, raw_sentence_number_int, sentence_number_int_aligned, crisisno_crisis_version) %>% 
  dplyr::mutate(crisno = as.numeric(crisno))
```

```{r}
icb_long_preceding_event <- icb_long_clean %>% 
  # get do_timing_reorder and do_timing variables
  dplyr::filter(variable=="do_timing_reorder") %>% 
  dplyr::mutate(value = as.numeric(stringr::str_remove(value, "sentence "))) %>% 
  dplyr::rename(prior_sent_raw = value) %>% 
  # merge in "crisisno_crisis_version""
  dplyr::left_join(alignment_key %>% 
                     dplyr::select(email_id, crisno, sentence, crisisno_crisis_version)) %>% 
  dplyr::left_join(alignment_key %>% 
                     dplyr::select(-sentence) %>%
                     dplyr::rename(prior_sent_raw = raw_sentence_number_int, 
                                   prior_sent_aligned = sentence_number_int_aligned) %>% 
                     dplyr::mutate(prior_sent_raw = as.numeric(prior_sent_raw))) %>%
  dplyr::rename(value = prior_sent_raw) %>% 
  dplyr::mutate(value = as.character(value)) %>% 
  dplyr::mutate(variable_normalized = "do_timing_reorder_clean") %>% 
  dplyr::mutate(value_normalized = ifelse(value==1, "drop", prior_sent_aligned)) %>%
  dplyr::select(-c(crisisno_crisis_version, prior_sent_aligned)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert5_397_080_1" & variable_normalized=="do_timing_reorder_clean", 60)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert5_403_064_1" & variable_normalized=="do_timing_reorder_clean", 60)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert5_403_099_1" & variable_normalized=="do_timing_reorder_clean", 53)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_115_006_1" & variable_normalized=="do_timing_reorder_clean", 4)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_115_006_2" & variable_normalized=="do_timing_reorder_clean", 4)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_119_031_1" & variable_normalized=="do_timing_reorder_clean", 22)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_123_010_1" & variable_normalized=="do_timing_reorder_clean", 8)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_123_011_1" & variable_normalized=="do_timing_reorder_clean", 8)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_149_017_1" & variable_normalized=="do_timing_reorder_clean", 11)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_149_018_1" & variable_normalized=="do_timing_reorder_clean", 12)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_152_029_1" & variable_normalized=="do_timing_reorder_clean", 22)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_152_029_2" & variable_normalized=="do_timing_reorder_clean", 22)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_152_031_1" & variable_normalized=="do_timing_reorder_clean", 22)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_342_013_2" & variable_normalized=="do_timing_reorder_clean", 9)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_186_018_1" & variable_normalized=="do_timing_reorder_clean", 15)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_233_009_1" & variable_normalized=="do_timing_reorder_clean", 3)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_240_009_1" & variable_normalized=="do_timing_reorder_clean", 14)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_241_015_1" & variable_normalized=="do_timing_reorder_clean", 13)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_252_020_1" & variable_normalized=="do_timing_reorder_clean", 7)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_256_026_1" & variable_normalized=="do_timing_reorder_clean", 16)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_264_025_1" & variable_normalized=="do_timing_reorder_clean", 16)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_264_026_1" & variable_normalized=="do_timing_reorder_clean", 16)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_299_020_1" & variable_normalized=="do_timing_reorder_clean", 18)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_330_015_1" & variable_normalized=="do_timing_reorder_clean", 8)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_330_015_2" & variable_normalized=="do_timing_reorder_clean", 8)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_348_042_1" & variable_normalized=="do_timing_reorder_clean", 25)) %>%
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="undergrada31ef865_410_025_1" & variable_normalized=="do_timing_reorder_clean", 24))

```

```{r}
icb_long_think_event <- icb_long_clean %>% 
  # get do_timing_reorder and do_timing variables
  dplyr::filter(variable=="think_sentence_events") %>% 
  dplyr::mutate(value = trimws(stringr::str_remove(value, "sentence "))) %>% 
  # dplyr::mutate(value = if(value %in% c("act in this sentence", "speech in this sentence")), raw_sentence_number_int, value) %>%
  dplyr::rename(think_sent_raw = value) %>% 
  # merge in "crisisno_crisis_version""
  dplyr::left_join(alignment_key %>% 
                     dplyr::select(email_id, crisno, sentence, crisisno_crisis_version)) %>% 
  dplyr::left_join(alignment_key %>% 
                     dplyr::select(-sentence) %>%
                     dplyr::rename(think_sent_raw = raw_sentence_number_int, 
                                   think_sent_aligned = sentence_number_int_aligned) %>% 
                     dplyr::mutate(think_sent_raw = trimws(as.character(think_sent_raw)))) %>%
  # dplyr::filter(prior_sent_raw != 1) %>% # drop if prior_sentence==1 which seems to be GUI default
  dplyr::rename(value = think_sent_raw) %>% 
  dplyr::mutate(value = as.character(value)) %>% 
  dplyr::mutate(variable_normalized = "think_sentence_events_clean") %>% 
  dplyr::mutate(value_normalized = ifelse(value %in% c("act in this sentence", "speech in this sentence"), 
                                          sentence_number_int_aligned, think_sent_aligned)) %>%
  dplyr::mutate(value_normalized = as.character(value_normalized)) %>%
  dplyr::select(-c(crisisno_crisis_version, think_sent_aligned)) %>% 
  # clean up some NAs
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_392_017_1" & variable_normalized=="think_sentence_events", 16)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert3_007_007_1" & variable_normalized=="think_sentence_events", 7)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_173_012_1" & variable_normalized=="think_sentence_events", 11)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert4_203_004_1" & variable_normalized=="think_sentence_events", 3)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="expert2_253_012_1" & variable_normalized=="think_sentence_events", 9)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="undergrad32faf6f6_004_016_1" & variable_normalized=="think_sentence_events", 15)) %>% 
  dplyr::mutate(value_normalized=replace(value_normalized, email_crisis_sent_event=="undergraddb11c211_092_008_1" & variable_normalized=="think_sentence_events", 6))
```

```{r}
icb_long_clean  <- dplyr::bind_rows(
  icb_long_clean %>% 
    #dplyr::mutate(event_number_int = as.character(event_number_int)) %>% 
    dplyr::filter(variable!="do_timing_reorder") %>% 
    dplyr::filter(variable!="think_sentence_events"),
  icb_long_preceding_event,
  icb_long_think_event
)

rm(icb_long_preceding_event, icb_long_think_event)

```

```{r}
icb_long_clean <- icb_long_clean %>% 
      dplyr::mutate(variable_normalized = (ifelse(is.na(variable_normalized), variable, variable_normalized))) %>%
      dplyr::mutate(value_normalized = (ifelse(is.na(value_normalized), value, value_normalized)))
```

Join in google sheet of tree simplification and drop variables that we replace

```{r message=FALSE}

full_tree <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/10tZGzjYgmvrbgQTV3oadVEV8m1LxLEDJEq9pzqq90Fc/edit?usp=sharing", sheet="Codebook in Tree Form")

full_tree <- full_tree %>% dplyr::filter(!is.na(leaf) & leaf!='')

#The leafs aren't unique here because you can get to the same branch multiple ways

#This ignores conditions FYI
#Behavior # 3 bits
#think_aggression #1 bit
#say_aggression #1 bit
#do_aggression #1 bit
#Do_Armed_Unarmed #1 bit
#do_Act_Interact #1 bit
#full_tree$leaf_Simplified %>% unique() %>% length() #66 bits

#Only 70 bits total

#That's only another 29 bits

full_tree$leaf_think_simplified %>% unique() %>% length() #6
full_tree$leaf_say_simplified %>% unique() %>% length() #5
full_tree$leaf_do_simplified %>% unique() %>% length() #18

# full_tree$think_type_l0 %>% unique() %>% length() #6
# full_tree$say_type_l0 %>% unique() %>% length() #5
# full_tree$do_type_l0 %>% unique() %>% length() #18

```

```{r}

drops <- c(
'interact_escalate',
'interact_deescalate',
'act_uncooperative',
'act_deescalate',
'act_cooperative',
'do_interact_kind',
'do_kind',
'interact_decreasecoop',
'interact_increasecoop',
'interact_deescalate',
'act_escalate',
'condition_interact_deescalate',
'condition_interact_escalate',
'condition_interact_increasecoop',
'condition_interact_decreasecoop',
'condition_act_cooperative',
'condition_act_deescalate',
'condition_act_escalate',
'condition_act_uncooperative',
'do_interact_kind',
'interact_escalate',
'sayintkind',
'sayintkind_react',
'interact_deescalate',
'interact_decreasecoop',
'act_escalate',
'act_uncooperative',
'act_deescalate',
'act_cooperative',
'act_descalate', 
'act_escalate',
'act_uncooperative',
'interact_decreasecoop',
'interact_descalate',
'interact_deescalate', #there's a mispelling here
'interact_escalate',
'interact_increasecoop',
'thinkkind',
'sayintkind'
) %>% unique()

icb_long_clean2 <- icb_long_clean %>% 
                            dplyr::distinct() %>%
                            dplyr::mutate(variable_normalized= variable_normalized %>%
                                            stringr::str_replace("_other_clean","")) %>% 
  #We want to collapse others in with regulars
                            dplyr::mutate(variable_normalized= variable_normalized %>% 
                                            stringr::str_replace("_clean","")) %>% 
  #These only exist with actors now
                            dplyr::mutate(variable_normalized= variable_normalized %>% 
                                            stringr::str_replace("_other","")) %>% 
  #Only really matters for location_other

                            dplyr::filter(value_normalized!='drop') %>%
                            dplyr::filter(email_id!='expert3') %>% #expert3 Looks broken all the time
                            #I'm going to move to a structure where event numbers can have a subsection that is condition rather than keeping it in the variable names
                            dplyr::mutate(event_number_int_condition=variable_normalized %>%
                                            stringr::str_detect('condition_') %>% as.numeric() ) %>%
                            dplyr::mutate(variable_normalized=variable_normalized %>%
                                            stringr::str_replace('condition_',''))

temp_wide <- icb_long_clean2 %>%
             dplyr::select(icb_survey_version, email_id,  crisno, sentence_number_int_aligned, section, sentence,event_number_int,  event_number_int_condition, variable_normalized, value_normalized) %>%
             dplyr::mutate(value_normalized= value_normalized %>% trimws() %>% tolower()) %>%
             dplyr::mutate(variable_normalized= variable_normalized %>% trimws() %>% tolower()) %>%
             dplyr::filter(variable_normalized %in% (c( 'sentence_events', drops ) %>% trimws() %>% tolower()) ) %>%
             dplyr::distinct() %>% 
             dplyr::left_join(full_tree %>% 
                       dplyr::rename(value_normalized=leaf)  %>% 
                       dplyr::mutate(value_normalized= value_normalized %>% trimws() %>% tolower()) 
                       )

temp_long <- temp_wide %>% 
         dplyr::select(-variable_normalized,-value_normalized) %>%
         tidyr::pivot_longer(-c(icb_survey_version, email_id, crisno, sentence_number_int_aligned, section, sentence,   event_number_int,event_number_int_condition),
                      names_to='variable_normalized' , values_to= 'value_normalized')  %>%
         dplyr::filter(!is.na(value_normalized))

icb_long_clean_final <- icb_long_clean2 %>%
                   dplyr::select(icb_survey_version, email_id, crisno, sentence_number_int_aligned, section,  sentence,  event_number_int, event_number_int_condition, variable_normalized, value_normalized) %>%
                  rbind(temp_long) %>% dplyr::distinct() %>% 
                  dplyr::filter(!variable_normalized %in% drops)

```

Save icb_long_clean!
```{r}
saveRDS(icb_long_clean_final, paste0(here::here(), "/replication_data/temp/icb_long_clean.Rds"))
```

# create wide file

```{r}

listtostring <- function(x){
  x %>% unlist() %>% paste(., collapse="; ")
}

paste_rex <- function(x){ x %>% sort() %>% paste( sep=";", collapse=";")   } 
icb_wide_clean <- icb_long_clean_final %>% 
  #dplyr::select(-variable,-value) %>%
  tidyr::pivot_wider(names_from=variable_normalized, values_from=value_normalized, values_fn=paste_rex)  %>%

  dplyr::filter(!(event_number_int==0 & sentence_events %in% c('1','2','3'))) #It does not appear that we ever need an e0 if number of events is greater than zero
  #There's also a weird situation where they sometimes created events without tagging it as a think, do,say

saveRDS(icb_wide_clean, paste0(here::here(), "/replication_data/temp/icb_wide_clean.Rds"))

```