### 2022-01-20 replication_data/rmd/03_aggregation.Rmd

---
title: "03_aggregation"
output:
  html_document:
    df_print: paged
  toc: yes
editor_options:
  chunk_output_type: inline
---

# Intro

## Library Loads

```{r}
library(flextable)
library(tidyverse)
library(googlesheets4)

`%>%` <- magrittr::`%>%`

rex_paste <- function(x) { x %>% unique() %>% na.omit() %>% sort() %>% paste0(sep=";", collapse=";")}
rex_count <- function(x) { x %>% unique() %>% na.omit() %>% sort() %>% paste0(sep=";", collapse=";")}

```

## Hand coded aggregations

```{r}

icb_long_clean <- readRDS(file=paste0(here::here(), "/replication_data/temp/icb_long_clean.Rds"))

#icb_long_clean_original_393_43 <- icb_long_clean_original %>% filter(crisno==393 & sentence_number_int_aligned==43 ) %>% janitor::remove_empty() 
#Choose the most common sentence for each
sentences <- icb_long_clean %>% dplyr::count(crisno,sentence_number_int_aligned,sentence) %>% arrange(crisno, sentence_number_int_aligned, desc(n)) %>% 
  group_by(crisno, sentence_number_int_aligned) %>% filter(row_number()==1) %>% dplyr::select(-n)

```

# Sentence Level Agreement

Unanimous exact agreement- every expert coder chose the exact same options across all events for that question
Majority exact agreement- at least 2/3 of expert coders chose exactly the same options across all event for that question
This aggregates across conditions too

```{r}

keep_crisis=393
keep_sentence=38

options(warn = -1) #Groups that have no values generate max errors
#This doesn't have this yet
#event_number_int,  event_number_int_condition
rex_paste <- function(x){ paste0(x  %>% trimws() %>% sort() %>% unique() %>% na.omit(), collapse=";")  }
temp_long <- 
  icb_long_clean %>% 
    #filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence )  %>% #for debugging
    dplyr::select(email_id,crisno,sentence_number_int_aligned, section, sentence, event_number_int, event_number_int_condition,variable_normalized, value_normalized) %>% 
    dplyr::filter(value_normalized!='' & !is.na(value_normalized)) %>%
    dplyr::group_by(crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
    dplyr::select(-event_number_int) %>%
    dplyr::distinct() %>%
    dplyr::mutate(
      email_id = email_id %>% as.factor(), 
      #crisno= crisno %>% as.factor(),
      #sentence_number_int_aligned= sentence_number_int_aligned %>% as.factor(),
      crisno_sentence_number_int_aligned= paste0(crisno, "_", sentence_number_int_aligned) %>% as.factor(), #make sure these are the only factors or it'll try to criss cross sentence number and crisno etc.
      section= section %>% as.character(),
      sentence= sentence %>% as.character(),
      variable_normalized= variable_normalized %>% as.factor()
    ) %>%
    dplyr::select(email_id , crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
    distinct() %>%
    dplyr::group_by(email_id , crisno,sentence_number_int_aligned, event_number_int_condition, 
             variable_normalized,.drop=FALSE) %>% 
      #This drop false means every coder gets a '' whether they filled in that variable or not #section, sentence going to have to ignore the sentence text for now
      #You'll get a warning about adding it as a group
      summarise(value_normalized=value_normalized %>% rex_paste() ) %>%
    #that will add in dead ones now we want to split them back out
    dplyr::mutate(value_normalized = value_normalized %>% stringr::str_replace_all(";;",";")) %>% 
    dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";", fixed=T)) %>% 
    tidyr::unnest(value_normalized) %>%
    dplyr::distinct() %>%
    #add a confidence score for every token
    dplyr::mutate(confidence=ifelse(variable_normalized=='raterconfidence', value_normalized, NA)) %>%

    dplyr::mutate(confidence=confidence %>% factor(levels=c('none','low','high','complete')) %>% as.numeric() ) %>%
    dplyr::mutate(confidence_expert= email_id %>% stringr::str_detect('expert') %>% ifelse(confidence,NA) %>% max(na.rm=T) ) %>%
    dplyr::mutate(confidence_undergrad= email_id %>% stringr::str_detect('undergrad') %>% ifelse(confidence,NA) %>% max(na.rm=T) ) %>%

    dplyr::group_by(crisno,sentence_number_int_aligned) %>%
      dplyr::mutate(
        total_coders=email_id %>% unique() %>% length(),
        total_coders_expert=email_id %>% unique() %>% stringr::str_detect('expert') %>% sum() ,
        total_coders_undergrad=email_id %>% unique() %>% stringr::str_detect('undergrad') %>% sum()
      ) %>%
    dplyr::ungroup() %>%

    #Selected By
    dplyr::group_by(crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
      dplyr::mutate(
        selected_by_experts=email_id %>% unique() %>% stringr::str_detect('expert') %>% sum() ,
        selected_by_undergrads=email_id %>% unique() %>% stringr::str_detect('undergrad') %>% sum()
      ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(selected_by_any=selected_by_experts+selected_by_undergrads) %>%
    dplyr::left_join(sentences) %>%
    dplyr::mutate(
        selected_by_any_perc=selected_by_any/total_coders,
        selected_by_experts_perc=selected_by_experts/total_coders_expert ,
        selected_by_undergrads_perc=selected_by_undergrads/total_coders_undergrad ,
    ) %>%
    dplyr::ungroup()

```

```{r}

#We have to compile confidence separately
sentence_confidence_scores <- temp_long %>% 
  dplyr::select(email_id,crisno,sentence_number_int_aligned, confidence,confidence_expert,confidence_undergrad) %>% distinct() %>% 
  mutate(confidence= confidence %>% replace(., !is.finite(.), NA),
         confidence_expert = confidence_expert %>% replace(., !is.finite(.), NA),
         confidence_undergrad = confidence_undergrad %>% replace(., !is.finite(.), NA)
         ) %>% 
  dplyr::select(-email_id) %>%
  group_by(crisno, sentence_number_int_aligned) %>%
  summarise_all(mean, na.rm=T) %>%
  mutate(confidence= confidence %>% replace(., !is.finite(.), NA),
         confidence_expert = confidence_expert %>% replace(., !is.finite(.), NA),
         confidence_undergrad = confidence_undergrad %>% replace(., !is.finite(.), NA)
         )

codings_long <- temp_long %>%
                mutate(
                       expert_majority=(selected_by_experts>(total_coders_expert/2) & selected_by_experts>=2) %>% as.integer() ,
                       undergrad_majority= (selected_by_undergrads > (total_coders_undergrad/2) & total_coders_undergrad>=2 ) %>% as.integer()  ,
                       expert_or_undergrad_majority= ( expert_majority | (undergrad_majority & selected_by_experts>=1) ) %>% as.integer() 
                ) %>%
                group_by(crisno,sentence_number_int_aligned, variable_normalized, value_normalized) %>%
                  mutate(
                        total_coders_expert=total_coders_expert %>% replace(., is.nan(.), NA) %>% mean(na.rm=T) %>% replace(., is.nan(.), NA),
                        total_coders_undergrad=total_coders_undergrad %>% replace(., is.nan(.), NA) %>% mean(na.rm=T) %>% replace(., is.nan(.), NA)
                  ) %>%
                ungroup() %>%
                dplyr::select(-email_id) %>%
                dplyr::select(-confidence,-confidence_expert,-confidence_undergrad) %>%
                distinct() %>% #bc expert majority is maxed now these rows should be fine
                left_join(sentence_confidence_scores) %>%
                arrange(crisno,sentence_number_int_aligned, event_number_int_condition, 
                        variable_normalized, expert_majority %>% desc(), undergrad_majority %>% desc(), undergrad_majority %>% desc(), selected_by_any %>% desc() ) #event_number_int

options(warn = 1)

```

# Agreed

```{r}

#this has to be temp because we actually do a lot of filtering at the wide stage and so need to go back to long
#What we could do is say if there's
codings_long_agreement <- codings_long %>%

                          mutate(keep_majority_expert= 
                                    (selected_by_experts>=2 & #selected by at least two experts
                                     expert_majority==1)   %>% as.numeric()
                           ) %>%
                           mutate(keep_majority_undergrad= (selected_by_experts>=1 & expert_majority==0 & undergrad_majority==1)  %>% as.numeric() ) %>%

                           mutate(keep_majority= (keep_majority_expert | keep_majority_undergrad) %>% as.integer() ) %>%

                           group_by(crisno, sentence_number_int_aligned, event_number_int_condition, variable_normalized) %>%
                                mutate(keep_majority_max=keep_majority %>% max(na.rm=T)) %>%
                                mutate(selected_by_any_max= (selected_by_any* (value_normalized!='')) %>% max(na.rmt=T) ) %>% #only consider non missing
                           ungroup() %>%
                           mutate(keep_leftstanding = (keep_majority_max==0 & #Wasn't picked by a majority
                                                         selected_by_experts>=1 & #was picked by a least one expert
                                                         (value_normalized!='' & selected_by_any_max %in% selected_by_any) #Has the most total votes
                                                       ) %>% as.numeric() ) %>%
                           mutate(keep = ( keep_majority==1 | keep_leftstanding==1 ) %>% as.numeric() )

codings_long_agreement_393_43 <- codings_long_agreement %>% filter(crisno==393 & sentence_number_int_aligned==43 ) %>% janitor::remove_empty() 
codings_long_agreement_393_59 <- codings_long_agreement %>% filter(crisno==393 & sentence_number_int_aligned==59 )
codings_long_agreement_393_38 <- codings_long_agreement %>% filter(crisno==393 & sentence_number_int_aligned==38 )
codings_long_agreement_284_16 <- codings_long_agreement %>% filter(crisno==284 & sentence_number_int_aligned==16 )

codings_long_agreed <- codings_long_agreement %>%
                           filter(value_normalized!='')  %>%
                           filter(keep==1) %>% 
                           group_by(crisno, sentence_number_int_aligned) %>%
                              mutate(no_events= max( variable_normalized=="Event_Any" & value_normalized=="no") ) %>% #make sure you do the filter or this won't twork
                           ungroup()

dim(codings_long_agreed) #249,840     19  tokens that have agreement
codings_long_agreed %>% dplyr::select(keep_majority_expert, keep_majority_undergrad, keep_majority, keep_leftstanding) %>% summarise_all(sum)

(table(expert_majority=codings_long$expert_majority,
      undergrad_majority=codings_long$undergrad_majority) / nrow(codings_long) ) %>% round(2)
#               undergrad_majority
#expert_majority    0    1
#              0 0.40 0.04
#              1 0.40 0.16

#
#
#
#Maybe we use this as a screen on the original codings and then collect what survived to make subevents?
rex_paste <- function(x){ paste0(x %>% unlist() %>% trimws() %>% sort() %>% unique() %>% na.omit(), collapse=";")  }
codings_wide_agreed <- icb_long_clean  %>% arrange(crisno,sentence_number_int_aligned) %>%
                        #filter(crisno==59) %>%
                        inner_join(codings_long_agreed) %>% #only keep ones that show up in both
                        filter(email_id %>% str_detect('expert') ) %>%
                        dplyr::select(email_id,crisno,sentence_number_int_aligned, event_number_int, event_number_int_condition, sentence,variable_normalized, value_normalized) %>%
                        distinct() %>%
                        pivot_wider(id_cols=c(email_id,crisno,sentence_number_int_aligned, sentence,event_number_int, event_number_int_condition), 
                                    names_from=variable_normalized, values_from=value_normalized, values_fn=rex_paste) %>%
                        distinct() %>%
                        group_by(crisno,sentence_number_int_aligned, event_number_int_condition, sentence, say_actor_a, say_actor_b, do_actor_a, do_actor_b) %>%
                          dplyr::summarise_all(rex_paste)
                        #filter(crisno==59 & sentence_number_int_aligned==3) 
codings_wide_agreed[codings_wide_agreed == ""]<-NA
#do the filters here after guaranteeing NAs

#Kill degenerate events
#We have a problem in that we want to reject certain kinds of events but we need to switch them to no events not drop the sentence all together
#This pruning basically works as intended but we need to go back and fill in null codings with a no event coding
codings_wide_agreed_pruned <- codings_wide_agreed %>% 
                              arrange(crisno,sentence_number_int_aligned) %>%
                              #filter(crisno==2 ) %>% #& sentence %>% str_detect('assassinated')

                              #A number of things stop making sense at this level
                              dplyr::select(-behavior) %>%
                              dplyr::select(-event_type) %>%
                              dplyr::select(-event_any) %>%
                              dplyr::select(-event_count) %>%
                              dplyr::select(-raterconfidence) %>%
                              dplyr::select(-event_number_int) %>%
                              dplyr::select(-email_id) %>%

                              #mutate(reject_eventbutnobehavior=!event_any %in% "no" & is.na(behavior)) %>% #it's pre NA conversion below so ''
                              #there's a degenerate kind of actor group where behavior wasn't agreed upon by everyone but there's still some info. This effectively requires each sentence to only have one kind of event, think, say, do but the details might differ slightly by different sets of actors involved.   
                              #I don't know why but the actos are infact NAs, I guess because of the group
                              mutate(reject_degendactorsdo  = !is.na(do_actor_b) & is.na(do_actor_a)     )  %>% 
                              mutate(reject_degendactorssay = !is.na(say_actor_b) & is.na(say_actor_a)   )  %>% 
                              mutate(reject_degendactorsall = is.na(do_actor_b) & is.na(do_actor_a) & is.na(say_actor_b) & is.na(say_actor_a) & is.na(think_actor_a)  ) %>%  #I guess we're just going to reject background codings all together

                              mutate(reject_degenleafdo    = !is.na(do_actor_a) & is.na(leaf_do_original)          ) %>% #kill off pairings that have a do but no do actor a
                              mutate(reject_degenleafsay   = !is.na(say_actor_a) & is.na(leaf_say_original)        ) %>%
                              mutate(reject_degenleafthink = !is.na(think_actor_a) & is.na(leaf_think_original)    ) %>%

                              mutate(reject_degenleafdo2   = is.na(do_actor_a) & !is.na(leaf_do_original)  ) %>% #kill of pairings that have a leaf but no actor
                              mutate(reject_degenleafsay2  = is.na(say_actor_a) & !is.na(leaf_say_original) ) %>%
                              mutate(reject_degenleafthink2= is.na(think_actor_a) & !is.na(leaf_think_original) ) %>%

                              #arrange(crisno,sentence_number_int_aligned) %>%
                              #mutate(crisno_sent= paste0(crisno,"_",sentence_number_int_aligned)) %>%
                              #group_by(crisno_sent) %>%
                              #  mutate(event_any_max= max( !event_any %in% 'no' ) ) %>%
                              #  mutate(reject_nullevent = event_any_max>0 & event_any %in% c("no")  ) %>% #If any group has events then drop any groups without events
                              #ungroup() %>%
                              rowwise() %>%
                                mutate(reject= 
                                               #reject_eventbutnobehavior +
                                               reject_degendactorsdo +
                                               reject_degendactorssay +
                                               reject_degendactorsall +
                                               reject_degenleafdo +
                                               reject_degenleafsay +
                                               reject_degenleafthink +
                                               reject_degenleafdo2 +
                                               reject_degenleafsay2 +
                                               reject_degenleafthink2 #+
                                               #reject_nullevent
                                         ) %>% 
                              ungroup()

#codings_wide_agreed_pruned  %>% t() %>% View() #%>% filter(sentence_number_int_aligned==12 )
#rejects <- codings_wide_agreed_pruned %>% filter(reject>0 ) 
#dim(rejects)   #19346    60
#rejects %>% t() %>% View()
#codings_wide_agreed_pruned %>% filter(crisno==2 & sentence_number_int_aligned==11) 
#codings_wide_agreed %>% filter(crisno==2 & sentence_number_int_aligned==11)

#codings_wide_agreed_384_43 <- codings_wide_agreed_pruned %>% filter(crisno==384 & sentence_number_int_aligned==43)  
#codings_wide_agreed_284_24 <- codings_wide_agreed_pruned %>% filter(crisno==284 & sentence_number_int_aligned==24)  
#codings_wide_agreed_447_14 <- codings_wide_agreed_pruned %>% filter(crisno==447 & sentence_number_int_aligned==14)

#codings_wide_agreed_2_11 <- codings_wide_agreed_pruned %>% filter(crisno==2 & sentence_number_int_aligned==11)

#codings_wide_agreed %>% filter(crisno==265 & sentence_number_int_aligned==28)

codings_wide_agreed_pruned_final <- codings_wide_agreed_pruned %>% 
                                    filter(reject==0) %>% dplyr::select(-starts_with("reject")) %>%
                                    right_join(codings_wide_agreed_pruned %>% dplyr::select(crisno,sentence_number_int_aligned,event_number_int_condition,sentence) %>% distinct() ) %>%

                                    #Generate final event ordering
                                    rowwise() %>%
                                    mutate(final_event_ordering=ifelse(!is.na(do_timing_reorder),
                                                                       #If more than one get past filter we take the smaller
                                                                       (do_timing_reorder %>% str_split(pattern=';', simplify = T) %>% unlist() %>% as.numeric() %>% min(na.rm=T)) +0.5,

                                                                       as.numeric(sentence_number_int_aligned) ) ) %>%
                                    ungroup() %>%
                                    arrange(crisno, final_event_ordering, sentence_number_int_aligned) %>% relocate(final_event_ordering, .after = sentence_number_int_aligned) %>%
                                    group_by(crisno) %>%
                                    mutate(final_event_ordering=row_number()) %>%
                                    ungroup()

```

# Codings wide agree labeled

```{r}

library(glue)

glue_rex <- function(x) glue(x, .open = "{|",.close = "|}", .envir = parent.frame()) #escape variables in the more complicated {||} syntax
target_file <- paste0(here::here(),"/replication_data/in/icb_manual_recoding_master_sheet.xlsx")
dictionary_actors_labels    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno, value_normalized=value_normalized_wikidata_id, value_normalized_label) %>% dplyr::distinct() %>% na.omit()
convert_qcode_to_english <- function(crisis, qcodes){
    crisis_actors <- dictionary_actors_labels %>% filter(crisno==crisis)
    crisis_actors_set <- crisis_actors$value_normalized_label
    names(crisis_actors_set) <- crisis_actors$value_normalized
    actors_english <- crisis_actors_set[qcodes %>% str_split(";") %>% unlist()]  %>% na.omit() %>% paste0(collapse=";")
    if(length(actors_english)==0){actors_english=NA}
    return(actors_english)
}
codings_wide_agreed_pruned_labeled <- NULL
codings_wide_agreed_pruned_labeled <-  codings_wide_agreed_pruned_final %>% 
  mutate(crisno_sent=paste0(crisno,'.',sentence_number_int_aligned)) %>%
  rowwise() %>%
  mutate(say_actor_a_labels=  convert_qcode_to_english(crisno, say_actor_a)) %>%
  mutate(say_actor_b_labels=  convert_qcode_to_english(crisno, say_actor_b)) %>%
  mutate(do_actor_a_labels=  convert_qcode_to_english(crisno, do_actor_a)) %>%
  mutate(do_actor_b_labels=  convert_qcode_to_english(crisno, do_actor_b)) %>%
  mutate(think_actor_a_labels=  convert_qcode_to_english(crisno, think_actor_a)) %>%
  mutate(think_sentence=ifelse(
      !is.na(leaf_think_original) & !is.na(think_sentence_events) , 
      glue_rex("[{|think_actor_a_labels|}]{.underline text.color='black'} experienced a 
                          [{|leaf_think_original|}]{.underline text.color='black'} about 
                          [{|think_sentence_events|}]{.underline text.color='black'}."),
            glue_rex("[{|think_actor_a_labels|}]{.underline text.color='black'} experienced a 
                          [{|leaf_think_original|}]{.underline text.color='black'}.")
  )) %>%
  mutate(say_sentence=ifelse(
      !is.na(leaf_say_original) , 
      glue_rex("[{|say_actor_a_labels|}]{.underline text.color='black'} initiated
                        [{|leaf_say_original|}]{.underline text.color='black'} toward  
                        [{|say_actor_b_labels|}]{.underline text.color='black'}.") ,
      ''
  )) %>%

  mutate(consequence_sentence=ifelse(
    consequence %in% "will happen" , 
    glue_rex("That the following will [happen]{.underline text.color='black'}.") ,
    glue_rex("That if the following [won't happen]{.underline text.color='black'}.")
  )) %>%

  mutate(condition_sentence=ifelse(
    condition %in% "happens" , 
    glue_rex("If the following [happens]{.underline text.color='black'}.") ,
    glue_rex("If the following [does not happen]{.underline text.color='black'}.")
  ))  %>%

  mutate(do_sentence=ifelse(
    !is.na(leaf_do_original) & !is.na(do_actor_b), 
    glue_rex("[{|do_actor_a_labels|}]{.underline text.color='black'} 
              initiated 
              [{|leaf_do_original|}]{.underline text.color='black'}
              toward 
              [{|do_actor_b_labels|}]{.underline text.color='black'}." ) ,
    glue_rex("[{|do_actor_a_labels|}]{.underline text.color='black'} 
              initiated 
              [{|leaf_do_original|}]{.underline text.color='black'}")
   )) %>%

   mutate(codings_sentence=NA) %>%

   #
   mutate(codings_sentence=ifelse( is.na(leaf_say_original) & is.na(leaf_do_original) & is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("No events.")
                                   ,codings_sentence)) %>%

   mutate(codings_sentence=ifelse( is.na(leaf_say_original) & is.na(leaf_do_original) & !is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|think_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & is.na(leaf_do_original) & is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|say_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( is.na(leaf_say_original) & !is.na(leaf_do_original) & is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|do_sentence|}")
                                   ,codings_sentence)) %>%

    #Condition sentence
    mutate(codings_sentence=ifelse( is.na(leaf_say_original) & !is.na(leaf_do_original) & is.na(leaf_think_original) & is.na(consequence) & !is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("[Condition] {|do_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( is.na(leaf_say_original) & !is.na(leaf_do_original) & !is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|think_sentence|} {|do_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & is.na(leaf_do_original) & !is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|think_sentence|} {|say_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & !is.na(leaf_do_original) & !is.na(leaf_think_original) & is.na(consequence) & is.na(condition)  & event_number_int_condition %in% 0,
                                   glue_rex("{|think_sentence|} {|say_sentence|} {|do_sentence|}")
                                   ,codings_sentence)) %>%

    mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & !is.na(leaf_do_original) & is.na(leaf_think_original) & is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|say_sentence|} {|do_sentence|}")
                                   ,codings_sentence)) %>%

   mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & !is.na(leaf_do_original) & is.na(leaf_think_original) & !is.na(consequence) & is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|say_sentence|} {|consequence_sentence|} {|do_sentence|}")
                                   ,codings_sentence))  %>%

    mutate(codings_sentence=ifelse( !is.na(leaf_say_original) & !is.na(leaf_do_original) & is.na(leaf_think_original) & !is.na(consequence) & !is.na(condition) & event_number_int_condition %in% 0,
                                   glue_rex("{|say_sentence|} {|consequence_sentence|} {|do_sentence|} {|condition_sentence|}")
                                   ,codings_sentence))

#codings_wide_agreed_labeled %>% dplyr::select(codings_sentence, contains('actor'), contains('_original'), consequence, condition, event_number_int_condition) #%>% View()

#codings_wide_agreed_labeled %>% dplyr::select(contains('actor'), contains('_original'), consequence, condition, event_number_int_condition, contains('_sentence')) #%>% View()

(codings_wide_agreed_pruned_labeled_2_11 <- codings_wide_agreed_pruned_labeled %>% filter(crisno==2, sentence_number_int_aligned==11) %>% 
                                    dplyr::select(crisno, sentence_number_int_aligned, contains('actor'), contains('_original'), consequence, condition, event_number_int_condition, contains('_sentence')) ) #%>% View()

(codings_wide_agreed_pruned_labeled_28_5 <- codings_wide_agreed_pruned_labeled %>% filter(crisno==28, sentence_number_int_aligned==5) %>% 
                                    dplyr::select(crisno, sentence_number_int_aligned, contains('actor'), contains('_original'), consequence, condition, event_number_int_condition, contains('_sentence')) ) #%>% View()

(codings_wide_agreed_pruned_labeled_34_17 <- codings_wide_agreed_pruned_labeled %>% filter(crisno==34, sentence_number_int_aligned==17) %>% 
                                    dplyr::select(crisno, sentence_number_int_aligned, contains('actor'), contains('_original'), consequence, condition, event_number_int_condition, contains('_sentence')) ) #%>% View()

```

Iterate and visualize sentences until convinced the translation is right

```{r, eval=F}

codings_wide_agreed_labeled_sample_ft <- 
      codings_wide_agreed_pruned_labeled %>%
                group_by(crisno, sentence_number_int_aligned, crisno_sent, sentence) %>%
                summarize(codings_sentence = paste(codings_sentence, collapse="\n")) %>%
                dplyr::select(crisno_sent, sentence, codings_sentence)  %>%
                dplyr::filter(codings_sentence!="No events.") %>%
                group_by(crisno) %>%
                sample_n(size=1) %>%
                ungroup() %>%
                dplyr::select(crisno_sent, sentence, codings_sentence) %>%
                as_flextable() %>%
                colformat_md() %>% 
                flextable::fontsize(size = 7, part = "all") %>%
                flextable::bg( i = which( 1:nrow(codings_wide_agreed_pruned_labeled) %% 2 == 1)  , j=1:ncol(events_df), part = "body", bg = "#EFEFEF") %>%
                flextable::width(j = 1, width=0.3, unit = "in") %>%
                flextable::width(j = 2, width=3, unit = "in") %>%
                flextable::width(j = 3, width=3, unit = "in")

codings_wide_agreed_labeled_sample_ft  %>% saveRDS(paste0(here::here(), '/replication_paper/tables/codings_wide_agreed_labeled_sample_ft.Rds'))
codings_wide_agreed_labeled_sample_ft

```

# Validation Tests

```{r, eval=F}

keep_crisis=393
keep_sentence=38

(a <- icb_long_clean_original %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()  ) #%>% View()
(b <- temp_long %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()  ) #%>% View()
(c <- temp_wide %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty() )#%>% t()  ) #%>% View()
(d <- icb_long_clean %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()  ) #%>% View()
(e <- codings_long %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()  ) #%>% View()
(f <- codings_long_agreement %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()   ) #%>% View()
(g <- codings_long_agreed %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence ) %>% janitor::remove_empty()   ) #%>% View()
(e <- codings_wide_agreed  %>% filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence )  %>% janitor::remove_empty()  ) %>% t() #%>% View()

```

```{r, eval=T}

saveRDS(codings_long, paste0(here::here(), "/replication_data/temp/codings_long.Rds")) 
saveRDS(codings_long_agreement, paste0(here::here(), "/replication_data/temp/codings_long_agreement.Rds"))

saveRDS(codings_long_agreed, paste0(here::here(), "/replication_data/out/codings_long_agreed.Rds")) 
saveRDS(codings_wide_agreed_pruned_labeled, paste0(here::here(), "/replication_data/out/codings_wide_agreed.Rds"))

```