### 2022-01-20 replication_paper/rmd/ICBEdataset_paper_appendixPlayersandAgents.Rmd

---
title: "Appendix ICBe Players and Agents"
output: html_document
---

```{r, echo=F, results='asis', message=F, warning=F, eval=T , cache=F}
library(pacman)
p_load(tidyverse)
p_load(flextable)
p_load(ftExtra)
p_load(googlesheets4)
library(googlesheets4)

```

```{r, eval=T, echo=F, results='markup', include=T, message=F, cache=F, ft.arraystretch=0.75}
target_file <- paste0(here::here(),"/data/icb_manual_recoding_master_sheet.xlsx")
downloaded_file <- googledrive::drive_download(file="https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", path=target_file, type = "xlsx", overwrite=T)
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors")

```

```{r, eval=T, echo=F, results='markup', include=T, message=F, cache=F, ft.arraystretch=0.75}

actor_agents <- dictionary_actors %>%
  #dplyr::select(Player=value_normalized_label, Qcode=value_normalized_wikidata_id, Piece=value_disaggregated_wikidata_label, Qcode_=value_disaggregated_wikidata_id) %>% 
  filter(!is.na(value_normalized) & value_normalized!='') %>%
  mutate(
    value_normalized_label = value_normalized_label %>% str_replace_all("[^A-Za-z0-9 ]",""),   #Throws a latex error  #inputenc Error: Unicode char \u8: not set up for use with LaTeX
    value_disaggregated_wikidata_label = value_disaggregated_wikidata_label %>% str_replace_all("[^A-Za-z0-9 ]",""),   #Throws a latex error  #inputenc Error: Unicode char \u8: not set up for use with LaTeX
  ) %>%
  mutate(Player=paste0(value_normalized_label, " (",value_normalized_wikidata_id,")"),
         Piece=paste0(value_disaggregated_wikidata_label, " (",value_disaggregated_wikidata_id,")") 
         ) %>%
  mutate(Piece= ifelse(Piece %in% "NA (NA)" |  Piece==Player, "",Piece) ) %>%
  dplyr::select(Player,Piece) %>% 
  arrange(Player,Piece) %>% 
  distinct() %>%
  group_by(Player) %>%
  summarise(Piece= paste0(Piece, collapse=" ; ")) %>%
  mutate(Piece= Piece %>% str_replace_all(" ;  ; ","") ) %>%
  mutate(Piece= Piece %>% str_replace_all("^ ; | ; $","") ) %>%
  mutate(Player_Pieces = paste0(Player , " {", Piece, "}")) %>%
  mutate(Player_Pieces= Player_Pieces %>% str_replace_all(" \\{\\}","") )   %>%

  dplyr::select(Player_Pieces)

library(officer) #needed for fp_border(color="gray")
actor_agents_ft <- 
actor_agents %>% 
      dplyr::select(Player_Pieces) %>%
      flextable::as_flextable() %>%
      bg( i = which( 1:nrow(actor_agents) %% 2 == 1), part = "body", bg = "#EFEFEF") %>%
      #merge_v(part = "body") %>% 
      flextable::width( j = 1, width=7) %>%
      flextable::fontsize(size = 6, part = "body") %>%
      #flextable::line_spacing( space = 0.5, part = "body") %>% #i=1:nrow(lit_review_alt),
      flextable::padding(padding = 0, part = "body") %>%
      #hline(i=which( actor_agents$Player!=c(actor_agents$Player[2:length(actor_agents$Player)], NA) ), border = fp_border(color="gray", width = 0.25), part = "body")

      #add_header( "IR Player_Pieces"="Citations"

      #            top = TRUE ) %>%
      set_header_labels(
        values = list("Player_Pieces" = "Players {Agents}" ) 
        )

actor_agents_ft  %>% saveRDS(paste0(here::here(), '/paper/tables/actor_agents_ft.Rds'))
actor_agents_ft

#https://cran.r-project.org/web/packages/ftExtra/vignettes/format_columns.html

```