### 2022-01-20 replication_paper/rmd/ICBEdataset_paper_appendixVerbDomains.Rmd

---
title: "Appendix Verb Domains"
output: html_document
---

```{r, echo=F, results='asis', message=F, warning=F, eval=T , cache=F}
library(pacman)
p_load(tidyverse)
p_load(flextable)
p_load(ftExtra)
p_load(googlesheets4)
library(googlesheets4)

```

```{r}
icb_wide_clean <- readRDS(file=paste0(here::here(), "/data/icb_wide_clean.Rds"))
```

```{r , message=F}

icb_text <- icb_wide_clean %>% dplyr::select(crisno,sentence_number_int_aligned, sentence) %>% distinct() %>% dplyr::mutate_at(vars(sentence_number_int_aligned), as.numeric) %>% arrange(crisno,sentence_number_int_aligned)

# writeLines(icb_text$sentence %>% na.omit() %>% iconv("UTF-8", "ASCII", "?"), paste0(here::here(), "/replication_data/temp/sentences.txt"), useBytes = T)

fromscratch=F
if(fromscratch){

  p_load("spacyr")
  #spacy_install() 
  spacy_initialize(model = "en_core_web_sm")

  parsed <- spacy_parse(icb_text$sentence, nounphrase = TRUE)
  parsed_consolidated <- nounphrase_consolidate(parsed)
  parsed$doc_id_num <- parsed$doc_id %>% str_replace_all("text","") %>% as.numeric()
  parsed$crisno <- icb_text$crisno[parsed$doc_id_num]
  parsed$sentence_number_int_aligned <- icb_text$sentence_number_int_aligned[parsed$doc_id_num]  %>% as.numeric()
  parsed$sentence <- icb_text$sentence[parsed$doc_id_num]

  parsed %>% saveRDS(paste0(here::here(), "/data/parsed.Rds"))

  entities <- entity_extract(parsed, type = "all")
  entities %>% saveRDS(paste0(here::here(), "/data/entities.Rds"))

} else {
  parsed <- readRDS(paste0(here::here(), "/data/parsed.Rds"))
  entities <- readRDS(paste0(here::here(), "/data/entities.Rds"))

}

entities$doc_id_num <- entities$doc_id %>% str_replace_all("text","") %>% as.numeric()
entities$crisno <- icb_text$crisno[entities$doc_id_num]
entities_unique <- entities %>% dplyr::select(entity,entity_type) %>% mutate(entity = entity %>% str_to_lower() %>% str_replace_all("[^A-Za-z0-9]","") ) %>% distinct()

target_file <- paste0(here::here(),"/data/icb_manual_recoding_master_sheet.xlsx")
downloaded_file <- googledrive::drive_download(file="https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", path=target_file, type = "xlsx", overwrite=T)
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors")

unique_agent_q_codes <- dictionary_actors$value_disaggregated_wikidata_id %>% na.omit() %>% unique() %>% str_split(";") %>% unlist() %>% trimws() %>% unique()
unique_actor_q_codes <- dictionary_actors$value_normalized_wikidata_id %>% na.omit() %>% unique() %>% str_split(";") %>% unlist() %>% trimws() %>% unique()

verbs_sentence <- parsed %>% filter(pos=="VERB")

verbs <- parsed %>% filter(pos=="VERB") %>% dplyr::select(lemma, pos) %>% distinct()

entities_per_sentence <- entities %>% 
  count(doc_id) %>% 
  summarise(
    entities_per_sentence_min=min(n),
    entities_per_sentence_mean=mean(n),
    entities_per_sentence_max=max(n),
  )

```

```{r, messages=F, results='hide', warnings=F, include=FALSE}

#wordnet rdf (in tripple format)
fromscratch=F

wordnet_rdf <- readRDS(paste0(here::here(), "/replication_data/in/wordnet_rdf.Rds") )

subject_headings <- wordnet_rdf %>% as_tibble() %>% filter(value %>% str_detect('http://purl.org/dc/terms/subject'))
subject_headings_verbs <- subject_headings %>%
                          filter(value %>% str_detect('-v>')) %>% 
                          mutate(value=value %>% str_replace("<http://wordnet-rdf.princeton.edu/id/","2")) %>% 
                          mutate(value=value %>% str_replace("-v> <http://purl.org/dc/terms/subject> ","\t")) %>%
                          separate(value, c("a", "b"), extra = "drop", fill = "right", sep="\t") %>%
                          mutate(b = b %>% str_replace_all('\\"| \\.',""))

#subject_headings <- wordnet_rdf %>% as_tibble() %>% filter(value %>% str_detect('00002137')) #the first digit is for part of speech so we can remove that if we subset to just verbs

#wordnet csv
filenames <- list.files(path=paste0(here::here(), "/data_replication/in/wncsv-master/csv/") )
filepaths <- list.files(path=paste0(here::here(), "/data_replication/in/wncsv-master/csv/"), full.names = T)

wordnet_list <- lapply(filepaths, read_csv, col_names=F, progress=F, show_col_types=F)
names(wordnet_list) <- filenames

hypernyms <- wordnet_list[['wn_hyp.csv']] %>% 
              dplyr::select(a=X1,b=X2) %>% 
              left_join(wordnet_list[["wn_s.csv"]] %>% dplyr::select(a=X1, a_lemma=X3) ) %>% 
              left_join(wordnet_list[["wn_s.csv"]] %>% dplyr::select(b=X1, b_lemma=X3) )

library(igraph)
g_hypernyms <-  graph_from_data_frame(hypernyms, directed = TRUE)

wordnet <- wordnet_list[["wn_s.csv"]] %>% dplyr::select(wordnetid=X1, lemma=X3, tense=X4) %>% full_join( wordnet_list[['wn_g.csv']] %>% dplyr::select(wordnetid=X1, gloss=X2) )

wordnet_verbs <- wordnet %>% filter(tense=='v')

verbs_sentence_wordnet <- verbs_sentence %>% left_join(wordnet_verbs)
#dim(verbs_sentence_wordnet) #227,426

#I think we use sbert to embed the original sentence and the gloss and then calculate pairwise distances

```

```{r, echo=F, messages=F, warnings=F, results='hide'}

fromsrcatch=F
if(fromsrcatch){
  #py$sentences
  #py$glosses
  sentences_embeddings <- py$sentences_embeddings
  glosses_embeddings <- py$glosses_embeddings

  rownames(sentences_embeddings) <- py$sentences #[1:100]
  rownames(glosses_embeddings) <- py$glosses #[1:100]
  dim(glosses_embeddings)

  condition_glosses_embeddings <- verbs_sentence_wordnet$gloss %in% rownames(glosses_embeddings)
  table(condition_glosses_embeddings) #there are 130 glosses that aren't in there
  gloss_clean <- verbs_sentence_wordnet$gloss
  gloss_clean[!condition_glosses_embeddings] <- rownames(glosses_embeddings)[1] #just a placeholder need to kill after
  glosses_embeddings_expanded <- glosses_embeddings[gloss_clean ,]
  dim(glosses_embeddings_expanded)

  condition_sentences_embeddings <- verbs_sentence_wordnet$sentence %in% rownames(sentences_embeddings)
  table(condition_sentences_embeddings) #all are in it
  sentences_clean <- verbs_sentence_wordnet$sentence
  sentences_clean[!condition_sentences_embeddings] <- rownames(sentences_embeddings)[1] #just a placeholder need to kill after
  sentences_embeddings_expanded <- sentences_embeddings[sentences_clean ,]
  dim(sentences_embeddings_expanded)

  #Isn't it just one minus the other?
  verbs_sentence_wordnet$distances <- rowSums((glosses_embeddings_expanded-sentences_embeddings_expanded)^2)
  verbs_sentence_wordnet <- verbs_sentence_wordnet %>% arrange(crisno, sentence_number_int_aligned, token_id, distances)
  verbs_sentence_wordnet %>% saveRDS(paste0(here::here(), "/data_replication/temp/verbs_sentence_wordnet.Rds"))

} else {

  verbs_sentence_wordnet <- readRDS(paste0(here::here(), "/data_replication/temp/verbs_sentence_wordnet.Rds") )

}

verbs_sentence_wordnet_top <- verbs_sentence_wordnet %>% group_by(crisno, sentence_number_int_aligned, token_id) %>% filter(row_number()==1)
verbs_sentence_wordnet_top_unique <- verbs_sentence_wordnet_top %>% ungroup() %>% dplyr::select(lemma, wordnetid,tense, gloss) %>% group_by(lemma, wordnetid,tense, gloss) %>% count() %>% arrange(desc(n))

#length(verbs_sentence_wordnet_top_unique$wordnetid) #3312 wordnetids but only 2,470 have hyponyms
verbs_sentence_wordnet_top_unique_nodes <- V(g_hypernyms)[name %in% verbs_sentence_wordnet_top_unique$wordnetid]

#g_hypernyms_verbs_used <- subcomponent(g_hypernyms, verbs_sentence_wordnet_top_unique_nodes, mode = c("out"))

g_hypernyms_verbs_used_list <- lapply(verbs_sentence_wordnet_top_unique_nodes,
                                 FUN=function(x) subcomponent(g_hypernyms, x, mode = c("out"))$name
                                 )

g_hypernyms_verbs_used <- NULL
g_hypernyms_verbs_used <- subgraph(g_hypernyms, V(g_hypernyms)[name %in% unlist(g_hypernyms_verbs_used_list )] )
g_hypernyms_verbs_used_df <- as_edgelist(g_hypernyms_verbs_used, names = TRUE)
roots <- setdiff(g_hypernyms_verbs_used_df[,2],g_hypernyms_verbs_used_df[,1])

subject_headings_verbs_toadd <- subject_headings_verbs %>% filter(a %in% roots) #I want to add these edges to the graph
#V(g_hypernyms_verbs_used) #This is only 1600

#Add the new body labels
g_hypernyms_verbs_used <- g_hypernyms_verbs_used %>% add_vertices(nv=length(unique(subject_headings_verbs_toadd$b)), name = unique(subject_headings_verbs_toadd$b))
#V(g_hypernyms_verbs_used)
#Add edges from the degenerate roots to the domains
g_hypernyms_verbs_used <- g_hypernyms_verbs_used %>%  add_edges(subject_headings_verbs_toadd %>% as.matrix() %>% t() %>% as.vector())
#V(g_hypernyms_verbs_used)
#is_simple(g_hypernyms_verbs_used)
#is_dag(g_hypernyms_verbs_used)
g_hypernyms_verbs_used <- g_hypernyms_verbs_used %>% simplify( remove.multiple = TRUE, remove.loops = TRUE)
#is_simple(g_hypernyms_verbs_used)

g_hypernyms_verbs_used_3hops <- ego(g_hypernyms_verbs_used, order = 1, nodes = V(g_hypernyms_verbs_used)[name %in% unique(subject_headings_verbs_toadd$b)], mode = c("all", "out","in"), mindist = 0)
g_hypernyms_verbs_used_3hops_list <- lapply(g_hypernyms_verbs_used_3hops,
                                 FUN=function(x) subcomponent(g_hypernyms_verbs_used, x, mode = c("all"))$name
                                 )  
g_hypernyms_verbs_used_3hops <- subgraph(g_hypernyms_verbs_used, V(g_hypernyms_verbs_used)[name %in% unlist(g_hypernyms_verbs_used_3hops_list )] )
g_hypernyms_verbs_used_3hops_df <- as_edgelist(g_hypernyms_verbs_used_3hops, names = TRUE)

require(tidygraph)
lemma_df <- wordnet_list[["wn_s.csv"]] %>% dplyr::select(name=X1, lemma=X3) %>% mutate(name=name %>% as.character) %>% 
            group_by(name) %>% summarise(lemma=paste0(lemma, collapse=";"))
rownames(lemma_df) <- lemma_df$name

first_two_hops <- g_hypernyms_verbs_used_3hops_df %>% as_tibble() %>% 
                    left_join(lemma_df %>% rename(V1=name, lemma1=lemma)) %>%
                    left_join(lemma_df %>% rename(V2=name, lemma2=lemma)) %>%
                    dplyr::select(wordnetid_hop0=V2, wordnetid_hop1=V1,lemma1_hop0=lemma2, lemma1_hop1=lemma1)

first_two_hops_appendix <- first_two_hops %>% filter(wordnetid_hop1 %in% roots) %>% distinct() %>% arrange(lemma1_hop0,lemma1_hop1) %>% arrange(wordnetid_hop0,lemma1_hop1) %>% dplyr::select(wordnetid_hop0,lemma1_hop1)

first_two_hops_collapsed <- first_two_hops %>% 
                            dplyr::select(lemma1_hop0, lemma1_hop1) %>%
                            group_by(lemma1_hop0) %>%
                            summarise(
                              lemma1_hop1 = paste0(lemma1_hop1, collapse=" | ")
                            ) %>%
                            arrange(lemma1_hop0)

tblg_hypernyms_verbs_used <- g_hypernyms_verbs_used %>% as_tbl_graph()
V(tblg_hypernyms_verbs_used)$lemma <- lemma_df[V(tblg_hypernyms_verbs_used)$name,]$lemma
condition <- V(tblg_hypernyms_verbs_used)$lemma %>% is.na()
V(tblg_hypernyms_verbs_used)$lemma[condition] <- V(tblg_hypernyms_verbs_used)$name[condition]

g_hypernyms_verbs_used_rooted <- g_hypernyms_verbs_used %>% add_vertices(nv=1, name = "verbs")
#V(g_hypernyms_verbs_used_rooted)
root_df <- data.frame(a=unique(subject_headings_verbs_toadd$b)); root_df$b <- "verbs"

#Add edges from the degenerate roots to the domains
g_hypernyms_verbs_used_rooted <- g_hypernyms_verbs_used_rooted %>%  add_edges(root_df[,c('a','b')] %>% as.matrix() %>% t() %>% as.vector())
#as_edgelist(g_hypernyms_verbs_used_rooted, names = TRUE)

tblg_hypernyms_verbs_used_rooted <- g_hypernyms_verbs_used_rooted %>% as_tbl_graph()
V(tblg_hypernyms_verbs_used_rooted)$lemma <- lemma_df[V(tblg_hypernyms_verbs_used_rooted)$name,]$lemma
condition <- V(tblg_hypernyms_verbs_used_rooted)$lemma %>% is.na()
V(tblg_hypernyms_verbs_used_rooted)$lemma[condition] <- V(tblg_hypernyms_verbs_used_rooted)$name[condition]

library(data.tree)
library(networkD3) 
df_hypernyms_verbs_used_rooted <- g_hypernyms_verbs_used_rooted %>% as_edgelist( names = TRUE) %>% as.data.frame()
df_hypernyms_verbs_used_rooted_labeled <- df_hypernyms_verbs_used_rooted %>% as_tibble() %>% 
                                          left_join(lemma_df %>% rename(V1=name, lemma1=lemma)) %>%
                                          left_join(lemma_df %>% rename(V2=name, lemma2=lemma)) %>%
                                          mutate(lemma1=ifelse(is.na(lemma1), V1, lemma1)) %>%
                                          mutate(lemma2=ifelse(is.na(lemma2), V2, lemma2)) #%>%
                                          #dplyr::select(lemma1,lemma2)

library("dendextend")

tree_hypernyms_verbs_used_rooted <- df_hypernyms_verbs_used_rooted_labeled %>% as.Node(mode="network")
tree_hypernyms_verbs_used_rooted_table <- tree_hypernyms_verbs_used_rooted %>% ToDataFrameTypeCol()

tree_hypernyms_verbs_used_rooted_table_ready <- 
                                            tree_hypernyms_verbs_used_rooted_table %>% 
                                                                      left_join(lemma_df %>% dplyr::rename(level_3=name, lemma3=lemma)) %>% 
                                                                      left_join(lemma_df %>% dplyr::rename(level_4=name, lemma4=lemma)) %>%
                                                                      dplyr::select(level1=level_2, level2=lemma3, level3=lemma4) %>%
                                                                      arrange(level1,level2,level3) %>% distinct()

```

```{r, eval=T, echo=F, results='markup', include=T, message=F, cache=F}

domains <- verbs_sentence_wordnet_top %>% ungroup() %>%  
            dplyr::select(wordnetid) %>%
            mutate(wordnetid=wordnetid %>% as.character() ) %>% 
            left_join(subject_headings_verbs %>% dplyr::select(wordnetid=a,domain=b)) %>%
            count(domain) %>% 
            arrange(desc(n)) %>% 
            na.omit() %>%
            mutate(perc= round( n/sum(n) * 100 ) ) %>%
  left_join(
    tree_hypernyms_verbs_used_rooted_table_ready %>%
  dplyr::select(level1,level2) %>% distinct() %>% group_by(level1) %>% summarise(level2=level2 %>% paste0(collapse=" | ") ) %>% dplyr::select(domain=level1,Sysnets=level2)
  )

p_load(flextable)
p_load(ftExtra)
set_flextable_defaults(fonts_ignore=TRUE)

domains_ft <- domains %>%
        ungroup() %>% 
        mutate(Sysnets=Sysnets %>% str_replace_all("[^A-Za-z0-9 ;\\|]","")) %>%
        flextable::as_flextable() %>%

        #bg( j = condition, part = "body", bg = "#EFEFEF") %>%
        #width(j = 1, width=1.25) %>%
        #width(j = 2:ncol(lit_review_clean), width=0.25) %>%
        flextable::fontsize(size = 8, part = "body" ) %>% 
        #rotate(rotation="tbrl",part="header") %>%
        set_header_labels( values = list(crisno = "#", input_crisis = "Name", sentence_count="Sentences", word_count="Words", event_count_mean_sum="Events")) %>%
        #align(align = "center", part = "body") %>%
        #align(align = "center", part = "header") %>%
        set_table_properties(layout="autofit") %>%  
        flextable::padding(padding = 0, part = "all") %>%
        flextable::line_spacing(space = 1, part = "body") #%>%
        #ftExtra::colformat_md(j = 1, part="body")   #Make sure this goes last or it'll get overwritten by the colformat above

      #autofit()
      #fit_to_width(max_width=12) #This is crazy slow
      #flextable::autofit() 
      #set_table_properties(width = 1, layout = "autofit")

#https://cran.r-project.org/web/packages/ftExtra/vignettes/format_columns.html

domains_ft  %>% saveRDS(paste0(here::here(), '/paper/tables/domains_ft.Rds'))
domains_ft

```