### 2022-01-20 replication_paper/rmd/ICBEdataset_paper_figure_prep.Rmd

---
title: 'ICBEdataset_paper_prep_metro_map'
site: bookdown::bookdown_site
bibliography: ../ICBintro.bib
output: 
  rticles::arxiv_article:
    keep_tex: true
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{pifont}
  - \usepackage{newunicodechar}
  - \newunicodechar{✓}{\ding{51}}
  - \newunicodechar{✗}{\ding{55}}
  - \usepackage{array}
  - \usepackage{ctable} # added for demo
  - \usepackage{natbib} #added for latex citation within huxtable
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{titlesec}
  - \usepackage[parfill]{parskip}
  - \usepackage{makecell}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage[capposition=top]{floatrow}
  - \titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
  - \titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{0ex plus .2ex}
  - \DeclareUnicodeCharacter{00A0}{ }
  - \usepackage{setspace}
  - \usepackage{cellspace}
  - \setlength\cellspacetoplimit{0.8ex}
  - \renewcommand{\arraystretch}{0.8}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \usepackage{float}
---

# Intro

## Library Loads

```{r}
library(flextable)
library(tidyverse)
library(googlesheets4)
library(janitor)
library(ftExtra)
library(igraph)

set_flextable_defaults(fonts_ignore=TRUE) #Warning: fonts used in `flextable` are ignored because the `pdflatex` engine is used and not `xelatex` or `lualatex`. You can avoid this warning by using the `set_flextable_defaults(fonts_ignore=TRUE)` command or use a compatible engine by defining `latex_engine: xelatex` in the YAML header of the R Markdown document.
options(tidyverse.quiet = TRUE)
options(gargle_oauth_email = TRUE)

icb_crises <- read.csv(paste0(here::here(), "/inst/extdata/","icb1v14.csv")) 
#icb_crises <-readRDS(paste0(here::here(), "/data/icb1v14.Rds"))

```

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}
# MOVED TO figures/ICBEdataset_figure_semantic_embeddings.Rmd

#child=c('ICBEdataset_paper_PlotInterCoderAgreement.Rmd')
p_semantic_embeddings <- readRDS(file=paste0(here::here(), '/paper/figures/p_semantic_embeddings.Rds'))
ggsave(file=paste0(here::here(), '//paper/figures/p_semantic_embeddings.png'), plot = p_semantic_embeddings, width=8, height=8)
```

#Case  196 Cuban Missile

```{r}

crisis=196
plot_title <- icb_crises %>%
            filter(crisno==crisis) %>%
            #mutate(title="Crisis #" %>% paste0(crisno, " ", crisname, " (", yrtrig, "-",yrterm,")")) %>%
            mutate(title= crisname %>% paste0( " (", yrtrig, "-",yrterm,")")) %>%
            pull(title)

```

## ICBe Crisis Map

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}
#ICBe  196
library(ggplot2)
p_metro_map_196 <- readRDS(file=paste0(here::here(), '/paper/figures/metro_plots/p_metro_plot_196.Rds'))
actor_colors <- p_metro_map_196$actor_colors
ggsave(file=paste0(here::here(), '/paper/figures/p_metro_map_196.png'),
       plot = p_metro_map_196 +
         theme(plot.margin=unit(c(0,0,0,0),"in")) +
         scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
         scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))),
       width=12, height=14)

```

## Phoenix Crisis Map

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}
#ICBe  196
library(ggplot2)
p_phoenix_metro_plot_196 <- readRDS(file=paste0(here::here(), '/paper/figures/metro_plots/p_phoenix_metro_plot_196.Rds'))
actor_colors <- p_phoenix_metro_plot_196$actor_colors

save_file=paste0(here::here(), '/paper/figures/metro_plots/p_phoenix_metro_plot_196.Rds')

subtitle="Phoenix Dyadic Events"
caption=paste0("Data: Cline Center Historical Phoenix Event Data (Althaus et al. 2017)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- readRDS(save_file) +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_phoenix_metro_plot_196.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## Dyadic ICB

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_icb_metro_plot_196.Rds')
p_metro_plot <- readRDS(save_file)
actor_colors <- p_metro_plot$actor_colors
subtitle="ICB Dyadic Events"
caption=paste0("Data: Dyadic International Crisis Behavior (ICB) (Hewitt 2003)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_icb_metro_plot_196.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## MIDs

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_mids_metro_plot_196.Rds')
p_metro_plot <- readRDS(save_file)
actor_colors <- p_metro_plot$actor_colors
subtitle="Dyadic Militarized Interstate Disputes (MIDs)"
caption=paste0("Source: CrisisEvents.org (Douglass et al. 2021)
                Data: Dyadic Militarized Interstate Disputes (MIDs) (Maoz et al. 2019)
             ", p_metro_plot$labels$caption) #tag original to bottom
p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_mids_metro_plot_196.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

# Crisis 426

```{r}

crisis=426
plot_title <- icb_crises %>%
            filter(crisno==crisis) %>%
            #mutate(title="Crisis #" %>% paste0(crisno, " ", crisname, " (", yrtrig, "-",yrterm,")")) %>%
            mutate(title= crisname %>% paste0( " (", yrtrig, "-",yrterm,")")) %>%
            pull(title)

```

## ICBe

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

# #ICBe  426
p_metro_map_426 <- readRDS(file=paste0(here::here(), '/paper/figures/metro_plots/p_metro_plot_426.Rds'))
actor_colors <- p_metro_map_426$actor_colors
ggsave(file=paste0(here::here(), '/paper/figures/p_metro_map_426.png'),
      plot = p_metro_map_426 +
        theme(plot.margin=unit(c(0,0,0,0),"in")) +
        scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
        scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))),
      width=12, height=14)

devtools::install_github('davidgohel/ggiraph')

```

## Phoenix Crisis Map

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}
#ICBe  426
library(ggplot2)
p_metro_plot <- readRDS(file=paste0(here::here(), '/paper/figures/metro_plots/p_phoenix_metro_plot_426.Rds'))
actor_colors <- p_metro_plot$actor_colors

save_file=paste0(here::here(), '/paper/figures/metro_plots/p_phoenix_metro_plot_426.Rds')
subtitle="Phoenix Dyadic Events"
caption=paste0("Data: Cline Center Historical Phoenix Event Data (Althaus et al. 2017)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_phoenix_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## ICEWs Crisis Map

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_icews_metro_plot_426.Rds')
p_metro_plot <- readRDS(file=save_file)
actor_colors <- p_metro_plot$actor_colors

subtitle="ICEWS Dyadic Events"
caption=paste0("Source: CrisisEvents.org (Douglass et al. 2021)
                Data: Integrated Crisis Early Warning System (ICEWS) (Boschee et al. 2018)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_icews_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## Terrier Crisis Map

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_terrier_metro_plot_426.Rds')
p_metro_plot <- readRDS(file=save_file)
actor_colors <- p_metro_plot$actor_colors

subtitle="Terrier Dyadic Events"
caption=paste0("Source: CrisisEvents.org (Douglass et al. 2021)
             Data: TERRIER (Temporally Extended, Regular, Reproducible International Event Records) (Liang et al. 2018)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_terrier_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## Dyadic ICB

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_icb_metro_plot_426.Rds')
p_metro_plot <- readRDS(save_file)
actor_colors <- p_metro_plot$actor_colors
subtitle="ICB Dyadic Events"
caption=paste0("Data: Dyadic International Crisis Behavior (ICB) (Hewitt 2003)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_icb_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## MIDs

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_mids_metro_plot_426.Rds')
p_metro_plot <- readRDS(save_file)
actor_colors <- p_metro_plot$actor_colors
subtitle="Dyadic Militarized Interstate Disputes (MIDs)"
caption=paste0("Source: CrisisEvents.org (Douglass et al. 2021)
                Data: Dyadic Militarized Interstate Disputes (MIDs) (Maoz et al. 2019)
             ", p_metro_plot$labels$caption) #tag original to bottom
p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_mids_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```

## MIDs

```{r, eval=T, echo=F, results='hide', include=T, message=F, cache=F, warning=F,  ft.arraystretch=0.75, fig.width=6, fig.height=6}

library(ggplot2)
save_file=paste0(here::here(), '/paper/figures/metro_plots/p_mids_incidents_metro_plot_426.Rds')
p_metro_plot <- readRDS(save_file)
actor_colors <- p_metro_plot$actor_colors

subtitle="Militerized Interstate Dispute Incidents"
caption=paste0("Source: CrisisEvents.org (Douglass et al. 2021)
                Data: Militerized Interstate Dispute Incidents 5.0 (MIDs) (Palmer et al. 2021)
             ", p_metro_plot$labels$caption) #tag original to bottom

p_metro_plot <- p_metro_plot +
           theme(plot.margin=unit(c(0,0,0,0),"in")) +
  labs(
    title = plot_title ,
    subtitle = subtitle ,
    caption = caption
  ) +
   theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) 
ggsave(file=paste0(here::here(), '/paper/figures/p_mids_incidents_metro_plot_426.png'),
       plot = p_metro_plot ,
       width=12, height=14)

```