### 2022-01-26 replication_paper/rmd/ICBEdataset_figure_semantic_embeddings.Rmd
input datas

---
title: 'Intercoder ICBEdataset_figure_semantic_embeddings'
site: bookdown::bookdown_site
bibliography: ../ICBintro.bib
output: 
  rticles::arxiv_article:
    keep_tex: true
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{pifont}
  - \usepackage{newunicodechar}
  - \newunicodechar{✓}{\ding{51}}
  - \newunicodechar{✗}{\ding{55}}
  - \usepackage{array}
  - \usepackage{ctable} # added for demo
  - \usepackage{natbib} #added for latex citation within huxtable
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{titlesec}
  - \usepackage[parfill]{parskip}
  - \usepackage{makecell}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage[capposition=top]{floatrow}
  - \titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
  - \titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{0ex plus .2ex}
  - \DeclareUnicodeCharacter{00A0}{ }
  - \usepackage{setspace}
  - \usepackage{cellspace}
  - \setlength\cellspacetoplimit{0.8ex}
  - \renewcommand{\arraystretch}{0.8}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
---

# Intro

## Library Loads

```{r}

library(flextable); library(ftExtra)
library(tidyverse)
library(googlesheets4)

```

```{r}

codings_wide_agreed <- readRDS( paste0(here::here(), "/replication_paper/data/in/codings_wide_agreed.Rds")) 
codings_long_agreed <- readRDS( paste0(here::here(), "/replication_paper/data/in/codings_long_agreed.Rds")) 
icb_long_clean <- readRDS(file=paste0(here::here(), "/replication_paper/data/in/icb_long_clean.Rds")) 
codings_long <- readRDS( paste0(here::here(), "/replication_paper/data/in/codings_long.Rds")) 
codings_long_agreement <- readRDS( paste0(here::here(), "/replication_paper/data/in/codings_long_agreement.Rds"))

```

```{r}

sentences_unique <- icb_long_clean %>% dplyr::select(sentence) %>% distinct() %>% pull(sentence) %>% unique() %>% na.omit()
length(sentences_unique )

sentences_unique_sanitized <- sentences_unique %>% str_replace_all("[A-Z][a-z]*","Entity")
#sentences_unique_sanitized %>% as.data.frame() %>% View()

```

```{r, messages=F, results='hide'}
library(reticulate)
#py_discover_config()

use_python("/home/skynet3/anaconda3/bin/python/")

```

```{python}
#Can't get this and spacy to work in the same run. Once reticulate loads once it sticks with the environment that came first and I can't figure out how to switch it.
#pip install -U sentence-transformers

fromscratch=True

if fromscratch:
  from sentence_transformers import SentenceTransformer
  import numpy as np
  sentences=r.sentences_unique_sanitized
  len(sentences) #10383
  model = SentenceTransformer('sentence-transformers/all-mpnet-base-v2')
  sentences_embeddings = model.encode(sentences) #[0:100]
else:
  print("Skipping")

```

```{r}

sentences_embeddings <- py$sentences_embeddings
dim(sentences_embeddings)

#devtools::install_github("ropenscilabs/umapr")
library(umapr)
embedding <- umap(sentences_embeddings  )

sentences_embeded <- embedding[,c('UMAP1','UMAP2')] %>% as.data.frame()
sentences_embeded$sentence <- sentences_unique

codings_long_agreed <- readRDS( paste0(here::here(), "/replication_paper/data/in/codings_long_agreed.Rds")) 
codings_long_agreed_embeded <- codings_long_agreed %>% ungroup() %>% left_join(sentences_embeded)

codings_long_agreed_embeded %>% saveRDS(paste0(here::here(), '/replication_paper/temp/codings_long_agreed_embeded.Rds'))

codings_wide_agreed <- readRDS(paste0(here::here(), '/replication_paper/data/in/codings_wide_agreed.Rds'))
codings_wide_agreed_embeded <- codings_wide_agreed %>% ungroup() %>% left_join(sentences_embeded)

codings_wide_agreed_embeded_stratified_sample  <- codings_wide_agreed_embeded %>% 
                                                  mutate(UMAP1_round=UMAP1 %>%  round(), UMAP2_round=UMAP2 %>%  round()) %>%
                                                  dplyr::group_by(UMAP1_round,UMAP2_round) %>%
                                                  arrange(UMAP1,UMAP2) %>%
                                                  filter(crisno==crisno[1], sentence_number_int_aligned==sentence_number_int_aligned[1] )

sampled <- codings_wide_agreed_embeded_stratified_sample %>% ungroup() %>% mutate(crisno_sent=paste0(crisno,'.',sentence_number_int_aligned)) %>%
  dplyr::select(crisno_sent, UMAP1,UMAP2) %>% distinct()

codings_wide_agreed_embeded %>% saveRDS(paste0(here::here(), '/replication_paper/temp/codings_wide_agreed_embeded.Rds'))

var_keep <- c('Behavior','do_aggression','Do_Armed_Unarmed','leaf_do_original') %>% tolower() #'Event_Any',

value_means <- codings_long_agreed_embeded %>%
      dplyr::filter(variable_normalized %in% var_keep) %>% 
      dplyr::mutate(variable_normalized=variable_normalized %>% as.factor()) %>%
      dplyr::mutate(value_normalized=value_normalized %>% as.factor()) %>%
      dplyr::select(variable_normalized,value_normalized,UMAP1,UMAP2) %>%
      group_by(variable_normalized, value_normalized) %>% 
      summarise_all(median) %>%
      group_by(variable_normalized) %>%
          mutate(variable_normalized_count=  min(n(),4)  ) %>%
      ungroup()

codings_long_agreed_embeded_subset <- codings_long_agreed_embeded %>%
                                      dplyr::filter(variable_normalized %in% var_keep) %>%
                                      dplyr::mutate(value_normalized=value_normalized %>% as.factor())

library(RColorBrewer)
brewer_pal_info_df <- brewer.pal.info %>% add_rownames( var = "rowname") %>% arrange()
brewer_list <- list() 
for(i in c(11, 1:nrow(brewer_pal_info_df))){
  brewer_list[[as.character(i)]] <- brewer.pal(n=brewer_pal_info_df$maxcolors[i], brewer_pal_info_df$rowname[i])
}
posible_colors <- unlist(brewer_list) %>% unique(brewer_list) #%>% sample()

value_normalized=codings_long_agreed_embeded_subset$value_normalized %>% unique() %>% as.factor()
value_normalized_colors = posible_colors[1:length(value_normalized)] #ok so pick a bewer pallet with lots of values and then assign it to specific values
names(value_normalized_colors) = levels(value_normalized)

p_semantic_embeddings <-
      codings_long_agreed_embeded_subset %>%
      ggplot(aes(x=UMAP1,y=UMAP2, color=value_normalized )) +
        geom_point(size=0.5, alpha=0.5) + 
        geom_text(data=sampled, aes(x=UMAP1, y=UMAP2, label=crisno_sent), color='black',size=1.5) +

        facet_wrap( ~ variable_normalized) + 
        theme_bw() +
        theme(legend.position = "none") + 
        geom_label(data=value_means,
                   aes(x=UMAP1,y=UMAP2, fill = value_normalized, label=value_normalized, size=50),inherit.aes=F, # #variable_normalized_count
                   colour = "white", fontface = "bold", label.padding = unit(0.1, "lines")) + 
        labs(
            title = "Variation in Tags by Semantic Embeddings of Source Sentences",
            subtitle = "Sentences Embedding Sentence-BERT, UMAP Projection ",
            caption = ""
        ) +
       scale_color_manual(values = value_normalized_colors) +
       scale_fill_manual(values = value_normalized_colors)

p_semantic_embeddings %>% saveRDS(paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings.Rds'))
p_semantic_embeddings

```

```{r}

codings_wide_agreed_embeded_stratified_sample  <- codings_wide_agreed_embeded %>% 
                                                  mutate(UMAP1_round=UMAP1 %>%  round(), UMAP2_round=UMAP2 %>%  round()) %>%
                                                  dplyr::group_by(UMAP1_round,UMAP2_round) %>%
                                                  arrange(UMAP1,UMAP2) %>%
                                                  filter(crisno==crisno[1], sentence_number_int_aligned==sentence_number_int_aligned[1] )

```

```{r}

codings_wide_agreed_embeded_stratified_sample_table <-       codings_wide_agreed_embeded_stratified_sample %>%
                                                              group_by(crisno, sentence_number_int_aligned, crisno_sent, sentence) %>%
                                                                summarize(codings_sentence = paste(codings_sentence, collapse="\n")) %>%
                                                              ungroup() %>%
                                                              dplyr::select(crisno_sent, sentence, codings_sentence)  %>%
                                                              dplyr::filter(codings_sentence!="No events.") %>%
                                                              dplyr::select(id=crisno_sent, sentence, codings_sentence)

n=nrow(codings_wide_agreed_embeded_stratified_sample_table)
k=ncol(codings_wide_agreed_embeded_stratified_sample_table)
codings_wide_agreed_embeded_stratified_sample_ft <- 
            codings_wide_agreed_embeded_stratified_sample_table %>%
            as_flextable() %>%
            colformat_md(j=3) %>% 
            flextable::fontsize(size = 7, part = "all") %>%
            flextable::bg( i = which( 1:n %% 2 == 1)  , j=1:3, part = "body", bg = "#EFEFEF") %>%
            flextable::width(j = 1, width=0.3, unit = "in") %>%
            flextable::width(j = 2, width=3, unit = "in") %>%
            flextable::width(j = 3, width=3, unit = "in")

codings_wide_agreed_embeded_stratified_sample_ft  %>% saveRDS(paste0(here::here(), '/replication_paper/tables/codings_wide_agreed_embeded_stratified_sample_ft.Rds'))
codings_wide_agreed_embeded_stratified_sample_ft

#Smaller for main paper
temp2 <- codings_wide_agreed_embeded_stratified_sample_table[rep_len(c(T,F,F),length.out=nrow(codings_wide_agreed_embeded_stratified_sample_table)),] 
n=nrow(temp2)
codings_wide_agreed_embeded_stratified_sample_ft_smaller <- 
            temp2 %>%
            as_flextable() %>%
            colformat_md(j=3) %>% 
            flextable::fontsize(size = 7, part = "all") %>%
            flextable::bg( i = which( 1:n %% 2 == 1)  , j=1:3, part = "body", bg = "#EFEFEF") %>%
            flextable::width(j = 1, width=0.3, unit = "in") %>%
            flextable::width(j = 2, width=3, unit = "in") %>%
            flextable::width(j = 3, width=3, unit = "in")

codings_wide_agreed_embeded_stratified_sample_ft_smaller  %>% saveRDS(paste0(here::here(), '/replication_paper/tables/codings_wide_agreed_embeded_stratified_sample_ft_smaller.Rds'))
codings_wide_agreed_embeded_stratified_sample_ft_smaller

```