### 2022-02-07 README.Rmd
paper run1

---
output: 
  github_document:
      toc: true
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
``` 
[![CC BY-NC-SA 4.0][cc-by-nc-sa-shield]][cc-by-nc-sa]

The International Crisis Behavior Events (ICBe)
-----------------------------

This is a github repository for the International Crisis Behavior Events (ICBe) dataset. Submit any issues regarding the dataset, paper, or github repository using [the issues tab](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/issues/new/choose).

## The Paper:

[Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/paper/IntroducingICBe_DouglassEtAl_2021_BetaDraft_bookdown.pdf)

## The Data:

The agreed datasets are the final, dataset used in much of the paper and figures. It includes our best efforts at cleaning the data and reconciling intercoder agreement. The dataset is available in long and wide format.

  - [codings_long_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/codings_long_agreed.Rds)
  - [codings_wide_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/codings_wide_agreed.Rds)

The coding and cleaning process are described in the paper with additional information and details about the variables in the codebook.
  - [ICBEdataset Codebook](https://docs.google.com/document/d/1aJkweohbfIWtNpJw1CmXbeIiK6czbJ5iPyKwiYP1YlU/edit?usp=sharing)

## The Authors:

[Rex W. Douglass](http://www.rexdouglass.com), [Thomas Leo Scherer](http://tlscherer.com/), [J. Andrés Gannon](https://jandresgannon.com/), [Erik Gartzke](http://erikgartzke.com/), [Jon Lindsay](https://www.jonrlindsay.com/), [Shannon Carcelli](https://www.shannoncarcelli.com/), Jonathan Wilkenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge.

## Citation:
For any use of the dataset or paper, please cite:

Douglass, Rex W.,  Thomas Leo Scherer, J. Andrés Gannon, Erik Gartzke, Jon Lindsay, Shannon Carcelli, Jonathan Wiklenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge. 2022. "Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises." Unpublished Manuscript. Retrieved from https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset.

Replication Code and Analysis
-----------------------------

### Self Contained Package

All of the files used to create the datasets, tables, figures.

## RMarkdown files

### Replication Data

 - 01_compile_saves_and_align
    + compiles the original coding files into `./replication_data/in/icb_coder_crisis_sentence_event.Rds`. The original coding files are not on the public repository. Public users will load `icb_coder_crisis_sentence_event.Rds` directly.
    + aligns codings from multiple GUI versions on similar source sentences
    + creates `./replication_data/temp/ICBe_V1_long_unclean.Rds.Rds` (not included in online repo)
 - 02_format_and_clean
    + applies cleaning dictionaries to create `./replication_data/temp/ICBe_V1_long_clean.Rds` and `./replication_data/temp/ICBe_V1_long_clean.Rds` (not included in online repo)
 - 03_aggregation
    + applies aggregation algorithm to create `./replication_data/out/ICBe_V1_long_agreement.Rds`, `./replication_data/out/ICBe_V1_long_agreed.Rds`, `./replication_data/out/ICBe_V1_wide_agreed.Rds`.

### Replication Paper

  - CaseStudy196
      + creates Cuban Missile Crisis comparison table `/tables/CaseStudy196_ft.Rds`. 
  - CaseStudy426
      + creates DRC Civil War comparison table `/tables/CaseStudy426_ft.Rds`. 
  - ICBEdataset_figure_metro_maps
      + creates the metroplot figures in `/figures/metro_plots/.`.
  - ICBEdataset_figure_semantic_embeddings
  - ICBEdataset_paper_appendixICBCrises
  - ICBEdataset_paper_appendixVerbMeanings
  - ICBEdataset_paper_figure_prep
  - ICBEdataset_paper_litreviewtable
  - ICBEdataset_paper_PlotInterCoderAgreement

## ICBe data products

 - Agreed Data, long form (LINK NEEDED)
 - Agreed Data, wide form (LINK NEEDED)
 - [Raw codings](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/data/icb_coder_crisis_sentence_event_aligned.Rds)
    + the initial codings from different GUI versions adjusted so that sentence numbers are aligned.

### Additional data versions
 - [Clean data, wide form](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/data/icb_wide_clean.Rds)
    + each row is a coder-crisis-event-sentence (i.e. each time a coder looking at a certain sentence in a certain crisis identified an event) 
    + values may contain multiple entries separating with a semicolon
 - [Clean data, long form](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/data/icb_long_clean.Rds)
    + each row contains identifying information for a coder-crisis-event-sentence
    + the information about that sentence is distributed across multiple rows in the columns _variable_ and _value_
    + each value contains a single entry

## Data inputs:

### ICBe preparation
  - Cleaning dictionaries: used to clean raw codings for actors, actions, locations, and dates (./data/icb_manual_recording_master_sheet.xlsx)
    + actor_translator to compare actors across different datasets, created by hand with assistance from [07_actor_dictionaries](https://centerforpeaceandsecuritystudies.github.io/ICBEdataset/07_actor_dictionaries.html).
  - [Lit review and tree/leaf codebook](https://docs.google.com/spreadsheets/d/10tZGzjYgmvrbgQTV3oadVEV8m1LxLEDJEq9pzqq90Fc/edit#gid=1604363724)

### External datasets
  - [The ICB project](https://sites.duke.edu/icbdata/) 
    + System-level (./data/icb1v14.Rds) and Actor-level (./data/icb2v14.Rds) datasets
    + Dyadic-Level Crisis Data ([source](https://sites.duke.edu/icbdata/data-collections/))
  - [Militarized Interstate Disputes (MID)](https://correlatesofwar.org/data-sets/MIDs) version 5.01 at the incident level (MIDI 5.01) and incident-participant level (MIDIP 5.01) converted to Rds. 
  - [UCDP Georeferenced Event Dataset (GED) Global version 21.1](https://ucdp.uu.se/downloads/index.html#ged_global)
  - cameo.eventcode.txt adapted from [CAMEO Conflict and Mediation Event Observations Codebook](https://parusanalytics.com/eventdata/cameo.dir/CAMEO.09b6.pdf)
  - [Phoenix Event data](https://databank.illinois.edu/datasets/IDB-2796521)
  - [Terrier event data](https://osf.io/4m2u7/files/) 
    + too large to include in the github repository
    + to replicate, download the folder 'largegeolocatedata' to ICBEdata/replication_paper/data/ignore and decompress
  - [ICEWS data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/28075&version=30.0)
    + too large to include in the github repository
    + to replicate, download folder 'dataverse_files' to ICBEdata/replication_paper/data/ignore/ and decompress

### License

This work is licensed under a
[Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License][cc-by-nc-sa].

[![CC BY-NC-SA 4.0][cc-by-nc-sa-image]][cc-by-nc-sa]

[cc-by-nc-sa]: http://creativecommons.org/licenses/by-nc-sa/4.0/
[cc-by-nc-sa-image]: https://licensebuttons.net/l/by-nc-sa/4.0/88x31.png
[cc-by-nc-sa-shield]: https://img.shields.io/badge/License-CC%20BY--NC--SA%204.0-lightgrey.svg