### 2022-02-08 README.Rmd
readme and plot colors

---
output: 
  github_document:
      toc: true
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
``` 
[![CC BY-NC-SA 4.0][cc-by-nc-sa-shield]][cc-by-nc-sa]

The International Crisis Behavior Events (ICBe)
-----------------------------

This is a github repository for the International Crisis Behavior Events (ICBe) dataset. Submit any issues regarding the dataset, paper, or github repository using [the issues tab](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/issues/new/choose).

## The Paper:

[Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/replication_paper/IntroducingICBe_DouglassEtAl_2021_BetaDraft_bookdown.pdf)

## The Data:

The agreed datasets are the final dataset used in much of the paper and figures. It includes our best efforts at cleaning the data and reconciling intercoder agreement. The dataset is available in long and wide format.

  - [ICBe_V1_long_agreement.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + All coded values and information about how often they were selected by coders.  
  - [ICBe_V1_long_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_long_agreed.Rds)
      + `ICBe_V1_long_agreement.Rds` filtered down only to those codings that were agreed upon (see Algorithm 1 in paper).
  - [ICBe_V1_wide_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + `ICBe_V1_long_agreed.Rds` in wide form where each row is an event.

The coding and cleaning process are described in the paper with additional information and details about the variables in the codebook.
  - [ICBEdataset Codebook](https://docs.google.com/document/d/1aJkweohbfIWtNpJw1CmXbeIiK6czbJ5iPyKwiYP1YlU/edit?usp=sharing)

## The Authors:

[Rex W. Douglass](http://www.rexdouglass.com), [Thomas Leo Scherer](http://tlscherer.com/), [J. Andrés Gannon](https://jandresgannon.com/), [Erik Gartzke](http://erikgartzke.com/), [Jon Lindsay](https://www.jonrlindsay.com/), [Shannon Carcelli](https://www.shannoncarcelli.com/), Jonathan Wilkenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge.

## Citation:
For any use of the dataset or paper, please cite:

Douglass, Rex W.,  Thomas Leo Scherer, J. Andrés Gannon, Erik Gartzke, Jon Lindsay, Shannon Carcelli, Jonathan Wiklenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge. 2022. "Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises." Unpublished Manuscript. Retrieved from https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset.

Replication Code and Analysis
-----------------------------

### Self Contained Package

All of the files used to create the datasets, tables, figures.

## RMarkdown files

### Replication Data

 - 01_compile_saves_and_align
    + compiles the original coding files into `./replication_data/in/icb_coder_crisis_sentence_event.Rds`. The original coding files are not on the public repository. Public users will load `icb_coder_crisis_sentence_event.Rds` directly.
    + aligns codings from multiple GUI versions on similar source sentences
    + creates `./replication_data/temp/ICBe_V1_long_unclean.Rds.Rds` (not included in online repo)
 - 02_format_and_clean
    + applies cleaning dictionaries to create `./replication_data/temp/ICBe_V1_long_clean.Rds` and `./replication_data/temp/ICBe_V1_long_clean.Rds` (not included in online repo)
 - 03_aggregation
    + applies aggregation algorithm to create `./replication_data/out/ICBe_V1_long_agreement.Rds`, `./replication_data/out/ICBe_V1_long_agreed.Rds`, `./replication_data/out/ICBe_V1_wide_agreed.Rds`.

### Replication Paper

  - CaseStudy196
      + creates Cuban Missile Crisis comparison table `/tables/CaseStudy196_ft.Rds`. 
  - CaseStudy426
      + creates DRC Civil War comparison table `/tables/CaseStudy426_ft.Rds`. 
  - ICBEdataset_figure_metro_maps
      + creates the metroplot Rds files in `/figures/metro_plots/.`.
  - ICBEdataset_figure_semantic_embeddings
      + creates the figure `/figures/p_semantic_embeddings.png` and the `/tables/codings_wide_agreed_embeded_stratified_sample_ft.Rds` and `/tables/codings_wide_agreed_embeded_stratified_sample_ft_smaller.Rds`
  - ICBEdataset_paper_appendixICBCrises
      +  creates `/tables/ft_crisis_text_counts.Rds`.
  - ICBEdataset_paper_appendixVerbMeanings
      + creates `/tables/verbs_sentence_wordnet_top_ft.Rds`
  - ICBEdataset_paper_figure_prep
      + creates the metroplot pngs in `/figures/.`.
  - ICBEdataset_paper_litreviewtable
      + creates the table `/tables/litreview_ft.Rds`. 
  - ICBEdataset_paper_PlotInterCoderAgreement
      + creates the figure `figures/p_percent_chose_tag_by_concept.png`

## Data inputs:

### ICBe preparation
  - Cleaning dictionaries: used to clean raw codings for actors, actions, locations, and dates `/replication_paper/data/in/icb_manual_recording_master_sheet.xlsx`
  - Lit review and tree/leaf codebook: `/replication_paper/data/in/icbe_litreview_trees_sentences.xlsx`
  - Case study tables: `/replication_paper/data/in/CaseStudies.xlsx`

### External datasets
  - [The ICB project](https://sites.duke.edu/icbdata/) 
    + System-level (./data/icb1v14.Rds) and Actor-level (./data/icb2v14.Rds) datasets
    + Dyadic-Level Crisis Data ([source](https://sites.duke.edu/icbdata/data-collections/))
  - [Militarized Interstate Disputes (MID)](https://correlatesofwar.org/data-sets/MIDs) version 5.01 at the incident level (MIDI 5.01) and incident-participant level (MIDIP 5.01) converted to Rds. 
  - [UCDP Georeferenced Event Dataset (GED) Global version 21.1](https://ucdp.uu.se/downloads/index.html#ged_global)
  - cameo.eventcode.txt adapted from [CAMEO Conflict and Mediation Event Observations Codebook](https://parusanalytics.com/eventdata/cameo.dir/CAMEO.09b6.pdf)
  - [Phoenix Event data](https://databank.illinois.edu/datasets/IDB-2796521)
  - [Terrier event data](https://osf.io/4m2u7/files/) 
    + too large to include in the github repository
    + to replicate, download the folder 'largegeolocatedata' to ICBEdata/replication_paper/data/ignore and decompress
  - [ICEWS data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/28075&version=30.0)
    + too large to include in the github repository
    + to replicate, download folder 'dataverse_files' to ICBEdata/replication_paper/data/ignore/ and decompress

### License

This work is licensed under a
[Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License][cc-by-nc-sa].

[![CC BY-NC-SA 4.0][cc-by-nc-sa-image]][cc-by-nc-sa]

[cc-by-nc-sa]: http://creativecommons.org/licenses/by-nc-sa/4.0/
[cc-by-nc-sa-image]: https://licensebuttons.net/l/by-nc-sa/4.0/88x31.png
[cc-by-nc-sa-shield]: https://img.shields.io/badge/License-CC%20BY--NC--SA%204.0-lightgrey.svg