### 2022-07-17 replication_data/rmd/04_validation.Rmd

---
title: "04_validation"
output:
  html_document:
    df_print: paged
  toc: yes
editor_options:
  chunk_output_type: inline
---

## R Markdown

```{r, eval=F }

ICBe_events_agreed <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed.Rds")) 
dim(ICBe_events_agreed)

icb_events_long <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_long.Rds")) 
dim(icb_events_long)

country_flags <- "/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/countries.csv" %>% read_csv(col_names=F)
library(WikidataR)# install.packages("WikidataR","WikidataQueryServiceR")
#https://commons.wikimedia.org/wiki/File:Flag_of_the_United_Nations.svg
UN <- get_item("Q1065")[[1]]$claims$P41

library(countrycode)
flag_codes1 <- 
  bind_rows(
    icb_events_long %>%
    filter(varname_normalized %>% str_detect("actor")) %>%
    dplyr::select(value_normalized, value_qcode) %>%
    distinct() ,

    actor_translator %>% mutate(value_normalized=coalesce(QCode_name,ICEWS_Actor.Name , ICEWS_Country, COWname , UCDPname )) %>% dplyr::select(value_normalized, value_qcode=QCode) %>% distinct()
  ) %>% 
  filter(!duplicated(value_qcode)) %>%
  rowwise() %>%
  mutate(qcode_n= (value_qcode %>% str_split(';'))[[1]] %>% length()) %>%
  filter(qcode_n==1) %>%
  left_join(country_flags %>% dplyr::select(value_normalized=X1, country_code2digit1=X3) ) %>% 
  left_join(country_flags %>% dplyr::select(value_normalized=X2, country_code2digit2=X3) ) %>% 
  mutate(country_code2digit3=value_normalized %>% countrycode::countrycode( origin = 'country.name', destination = 'iso2c')) %>%
  mutate(country_code_iso3c=value_normalized %>% countrycode::countrycode( origin = 'country.name', destination = 'iso3c')) %>%
  mutate(unicode_symbol=value_normalized %>% countrycode::countrycode( origin = 'country.name', destination = 'unicode.symbol')) %>%
  mutate(country_code2digit = coalesce(country_code2digit1,country_code2digit2, country_code2digit3))

flag_codes1$wikidata_flag <- NA
flag_codes1$wikidata_iso3c <- NA

for(i in 1:nrow(flag_codes1)){
  if(!is.na(flag_codes1$wikidata_flag[i])){next}
  flag <- NULL
  iso3c <- NULL
  qcode <- str_split(flag_codes1$value_qcode[i] , ";")[[1]] %>% trimws()
  if(length(qcode)>1){next} #multi qcode rows are bad
  print(qcode)
  try({
    item <- get_item(qcode)[[1]]
    try({
      flag <- item$claims$P41$mainsnak$datavalue$value[1]
      if(class(flag)=="character"){
        flag_codes1$wikidata_flag[i]<-flag
      }
    })
    try({
      iso3c <- item$claims$P984$mainsnak$datavalue$value[1]
      if(class(iso3c)=="character"){
        flag_codes1$wikidata_iso3c[i]<-iso3c
      }
    })
  })
}

library(rvest)
library(jsonlite)
flag_codes1$wikidata_flag_url <- "https://commons.wikimedia.org/wiki/File:" %>% paste0(flag_codes1$wikidata_flag)
flag_codes1$wikidata_flag_url[is.na(flag_codes1$wikidata_flag)]<- NA
flag_codes1$wikidata_flag_url_file <- NA

#which(flag_codes1$value_qcode=="Q347")
for(i in 1:nrow(flag_codes1)){
  print(i)
  if(is.na(flag_codes1$wikidata_flag_url[i])){ next }
  destfile_svg <- glue::glue("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags/{flag_codes1$value_qcode[i]}.svg")
  destfile_png <- glue::glue("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags/{flag_codes1$value_qcode[i]}.png")

  if(!file.exists(destfile_svg) & !file.exists(destfile_png)){
    try({
      html <- flag_codes1$wikidata_flag_url[i] %>% str_replace_all(" ","_") %>% read_html()
      links <- html   %>% html_nodes("a") %>% html_attr("href") 
      df = data.frame(link=links) %>% distinct()
      image_url <- df %>% filter(link %>% str_detect("^http.*?svg$")) %>% filter(row_number()==1) %>% pull(link)
      if(length(image_url)!=0){
              flag_codes1$wikidata_flag_url_file[i] <- image_url
              download.file(url=image_url, destfile=destfile_svg)
      } else {
              image_url <- df %>% filter(link %>% str_detect("^http.*?png$")) %>% filter(row_number()==1) %>% pull(link)
              flag_codes1$wikidata_flag_url_file[i] <- image_url
              download.file(url=image_url, destfile=destfile_png)
      }
    })
  } else {
    print("Already exists")
  }
}

flag_folder <- "/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags/"
flag_files <- list.files(path = flag_folder, pattern = NULL, all.files = FALSE, full.names = FALSE, recursive = FALSE, ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)

flag_files_short <- flag_files %>% str_replace("\\..*$","")
flag_codes1$file_exists <- flag_codes1$value_qcode %in% flag_files_short

```

## Resave Small Flags

```{r}

library(tidyverse)
library(magick) #switching to magick which handles transparency correctly
library(imager); #install.packages("imager")
flag_folder <- "/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags/"
flag_folder_small <- "/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small/"
#20x16 pixels
file="Q15180"

flag_files <- list.files(path = flag_folder, pattern = NULL, all.files = FALSE, full.names = FALSE, recursive = FALSE, ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
for( file in flag_files){

  file_short <- NULL
  file_short <- file %>% str_replace("\\..*$","")
  outfile <- NULL
  outfile <- paste0(flag_folder_small, file_short, ".png")
  if(!file.exists(outfile)){
    print(file)
    im <- NULL
    infile <- NULL
    infile <- paste0(flag_folder, file)
    try({ #nevermind apparently if you convert too many files in a row image magick runs out of some internal memory "session: cache resources exhausted "
      im <- image_read(infile) #load.image(infile)
      im_small <- NULL
      im_small <- im %>% image_scale("20x16") #resize(im, 20, 16)
      im_small %>% image_write(path=outfile, format = "png") #save.image(outfile)
      next
    })
    try({
      im <- load.image(infile)
      im_small <- NULL
      im_small <- resize(im, 20, 16)
      im_small %>% save.image(outfile)
    })
  }
}

```

## Including Plots

```{r pressure, echo=FALSE}

ICBe_events_agreed <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed.Rds")) 
dim(ICBe_events_agreed)

icb_events_long <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_long.Rds")) 
dim(icb_events_long)

ICBe_events_agreed$interact_forces %>% unique()
ICBe_events_agreed$interact_fatalities %>% unique()

ICBe_events_agreed$interact_domains %>% unique()
ICBe_events_agreed$interact_territory %>% unique()
ICBe_events_agreed$interact_geoscope  %>% unique()

#![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/cub.png)

paste_vars <- c("think_actor_a","thinkkind",
"say_actor_a", "sayintkind", "say_actor_b",
"condition",
"do_actor_a" ,               
#"do_kind", "do_interact_kind" , #leaving this out because it rarely helps
"interact_increasecoop" , "interact_decreasecoop" ,  "interact_escalate"  , "interact_deescalate" , "act_cooperative","act_deescalate","act_escalate","act_uncooperative"   ,           
"do_actor_b",
"interact_units",
"interact_domains",
"interact_geoscope",
"interact_territory",
"interact_forces",
"interact_fatalities",
"consequence",
"condition_do_actor_a" ,               
#"condition_do_kind", "condition_do_interact_kind" , 
"condition_interact_increasecoop" , "condition_interact_decreasecoop" ,  "condition_interact_escalate"  , "condition_interact_deescalate" , "condition_act_cooperative","condition_act_deescalate","condition_act_escalate","condition_act_uncooperative"   ,           
"condition_do_actor_b",
"condition_interact_units",
"condition_interact_domains",
"condition_interact_geoscope",
"condition_interact_territory",
"condition_interact_forces",
"condition_interact_fatalities"
)

#We are going to exclude from descriptions common and not particularly informative values "country" "no clear change"
values_exclude <- c('country','no clear change',"none")
flag_files_small_short <- list.files(path = flag_folder, pattern = NULL, all.files = FALSE, full.names = FALSE, recursive = FALSE, ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE) %>% str_replace("\\..*$","")

#"individuals", "hundreds","thousands","tens of thousands", "hundreds of thousands"

icb_events_long_markdown <- icb_events_long %>%
                              filter(varname_normalized %in% paste_vars) %>%
                              mutate(varname_normalized = varname_normalized %>% factor(levels=paste_vars)) %>%
                              arrange(crisno, sentence_number_int_aligned, actor_set, varname_normalized, value_normalized, value_qcode) %>%
                              filter(!value_normalized %in% values_exclude) %>%
                              rowwise() %>%
                                mutate( value_markdown =  glue::glue("![]({flag_folder_small}/{value_qcode}.png)") )  %>% 
                                mutate( value_markdown = ifelse(is.na(value_qcode) | !value_qcode %in%  flag_files_small_short ,value_normalized, value_markdown ) ) %>% #if the q code is missing or the image is missing fall back to text
                                mutate(value_markdown= ifelse(value_normalized=="individuals",value_markdown %>% paste0("1s") ,value_markdown )) %>%
                                mutate(value_markdown= ifelse(value_normalized=="tens",value_markdown %>% paste0("10s"),value_markdown)) %>%
                                mutate(value_markdown= ifelse(value_normalized=="hundreds",value_markdown %>% paste0("100s"),value_markdown)) %>%
                                mutate(value_markdown= ifelse(value_normalized=="thousands",value_markdown %>% paste0("1ks"),value_markdown)) %>%
                                mutate(value_markdown= ifelse(value_normalized=="tens of thousands",value_markdown %>% paste0("10ks"),value_markdown)) %>%
                                mutate(value_markdown= ifelse(value_normalized=="hundreds of thousands",value_markdown %>% paste0("100ks"),value_markdown)) %>%
                              ungroup()

#Make a sentence level markdown summary                                      
icb_events_long_narratives_sentences <- icb_events_long_markdown %>%
                              group_by(crisno, sentence_number_int_aligned, sentence_span_text, actor_set, varname_normalized) %>%
                                summarise( 
                                    value_markdown = paste0(value_markdown , collapse=";")
                                  )  %>%
                              mutate(value_markdown = ifelse(varname_normalized=="think_actor_a", paste0("- ", value_markdown), value_markdown ) ) %>%
                              mutate(value_markdown = ifelse(varname_normalized=="say_actor_a", paste0(" - ", value_markdown), value_markdown ) ) %>%
                              mutate(value_markdown = ifelse(varname_normalized=="do_actor_a", paste0(" - ", value_markdown), value_markdown ) ) %>%
                              mutate(value_markdown = ifelse(varname_normalized=="consequence", paste0(" -C- ", value_markdown), value_markdown ) ) %>%
                              group_by(crisno, sentence_number_int_aligned,sentence_span_text,  actor_set) %>%
                                filter(is.na(lag(value_markdown)) | (value_markdown!=lag(value_markdown))) %>% #with the dashes added above it should allow us to remove dupes in the middle
                                summarise( 
                                    value_markdown = paste0(value_markdown, collapse=" ")
                                  ) %>% ungroup()

ICBe_events_agreed_markdown <- ICBe_events_agreed %>% left_join(icb_events_long_narratives_sentences)

ft <- ICBe_events_agreed_markdown %>%
      filter(crisno==196) %>%
      group_by(crisno, sentence_number_int_aligned) %>%
        filter(!duplicated(value_markdown %>% trimws() )) %>% #note there are dupe markdowns for some reason
      ungroup() %>%
      mutate(crisno= ifelse(!is.na(lag(sentence_span_text)) & sentence_span_text==lag(sentence_span_text), "", crisno)) %>%
      mutate(sentence_number_int_aligned= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_number_int_aligned)) %>%
      mutate(sentence_span_text= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_span_text)) %>%
      dplyr::select(C=crisno, S=sentence_number_int_aligned, ICB=sentence_span_text, ICBe=value_markdown) %>%
      #head(100) %>%
      flextable::as_flextable() %>%
      flextable::width( j = 1, width=0.25)  %>% 
      flextable::width( j = 2, width=0.25)  %>% 
      flextable::width( j = 3, width=5.0)  %>% 
      flextable::width( j = 4, width=5.0)  %>% 
      flextable::fontsize(size = 9, part = "all") %>%
      ftExtra::colformat_md(j = 4, part="body")
ft

#Make a narrative level markdown summary
icb_events_long_narratives <- icb_events_long_markdown %>%
                              #filter(crisno==196) %>%
                              group_by(crisno, varname_normalized) %>%
                                summarise( 
                                    value_markdown = paste0(value_markdown %>% sort() %>% unique() %>% trimws(), collapse=";") %>% trimws()
                                  )  %>%
                              #mutate(value_markdown = ifelse(varname_normalized=="think_actor_a", paste0("", value_markdown), value_markdown ) ) %>%
                              #mutate(value_markdown = ifelse(varname_normalized=="say_actor_a", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                              #mutate(value_markdown = ifelse(varname_normalized=="do_actor_a", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                              #mutate(value_markdown = ifelse(varname_normalized=="interact_units", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                              #mutate(value_markdown = ifelse(varname_normalized=="consequence", paste0("BREAK -C- ", value_markdown), value_markdown ) ) %>%
                              group_by(crisno) %>%
                              summarise( 
                                  value_markdown = paste0(value_markdown %>% trimws(), collapse=" ") %>% trimws()
                              ) %>% ungroup() #%>% 
                              #mutate(value_markdown = strsplit(as.character(value_markdown), "BREAK")) %>% 
                              #unnest(value_markdown)

icb_events_long_narratives_brokenup <- icb_events_long_markdown %>%
                                        mutate(var_group=NA) %>%
                                        mutate(var_group = ifelse(varname_normalized %>% str_detect("actor"),"actors",var_group)) %>%
                                        mutate(var_group = ifelse(varname_normalized %>% str_detect("coop|escalate"),"act",var_group)) %>%
                                        mutate(var_group = ifelse(varname_normalized %>% str_detect("actor"),"actors",var_group)) %>%
                                        mutate(var_group = ifelse(varname_normalized %>% str_detect("units|domains"),"units_domains",var_group)) %>%
                                        mutate(var_group = ifelse(varname_normalized %>% str_detect("forces|fatalities"),"forces_fatalities",var_group)) %>%
                                        #filter(crisno==196) %>%
                                        group_by(crisno, var_group) %>%
                                          summarise( 
                                              value_markdown = paste0(value_markdown %>% sort() %>% unique() %>% trimws(), collapse=";") %>% trimws()
                                            )  %>%
                                        #mutate(value_markdown = ifelse(varname_normalized=="think_actor_a", paste0("", value_markdown), value_markdown ) ) %>%
                                        #mutate(value_markdown = ifelse(varname_normalized=="say_actor_a", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                                        #mutate(value_markdown = ifelse(varname_normalized=="do_actor_a", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                                        #mutate(value_markdown = ifelse(varname_normalized=="interact_units", paste0("BREAK", value_markdown), value_markdown ) ) %>%
                                        #mutate(value_markdown = ifelse(varname_normalized=="consequence", paste0("BREAK -C- ", value_markdown), value_markdown ) ) %>%
                                        group_by(crisno, var_group) %>%
                                        summarise( 
                                            value_markdown = paste0(value_markdown %>% trimws(), collapse=" ") %>% trimws()
                                        ) %>% ungroup()

icb_long_clean <- readRDS(file=paste0(here::here(), "/replication_data/temp/ICBe_V1.1_long_clean.Rds")) %>%
                  filter(!varname %in% c("lastviewed","crisis",'sentencenumber','crisno', 'icb_survey_version',  'section' )) %>%
                  select(-sentence, -sentence_clean, -varname, -value, -status) %>% 
                  distinct()

ICBe_long_agreement <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_long_agreement.Rds"))

icb_crises <- read_csv(file=paste0(here::here(), "/replication_paper/data/in/icb1v14.csv")) %>% janitor::clean_names() #somehow break of all things is a broken name

ICBe_crises_markdown <- ICBe_events_agreed %>%
  #filter(crisno==196) %>%
  group_by(crisno, crisis_text) %>%
  summarise(
    n_sentence_number_int_aligned  = sentence_number_int_aligned %>% unique() %>% length(),
    n_events = sum(!is.na(do_actor_a) | !is.na(say_actor_a) | !is.na(think_actor_a))
  )  %>% distinct() %>% 
  left_join(icb_crises %>% dplyr::select(crisno, yrtrig,yrterm) %>% distinct() ) %>%
  left_join( ICBe_long_agreement %>% dplyr::select(crisno,, coders_expert, coders_novice,) %>% distinct() ) %>% 
  left_join(icb_events_long_narratives_brokenup %>%
              filter(!is.na(var_group)) %>%
              pivot_wider(id_cols=crisno, names_from=var_group, values_from=value_markdown)
              ) %>%
  ungroup()

ft <- ICBe_crises_markdown %>% ungroup() %>%
      #filter(crisno %in% c(1:100) ) %>%
      dplyr::select(crisno, crisis_text,  yrtrig , yrterm,  n_sentence_number_int_aligned, n_events,  coders_expert, coders_novice, actors, units_domains, forces_fatalities, act) %>%
      mutate(crisno = crisno %>% as.character() ) %>%
      mutate(yrtrig = yrtrig %>% as.character() ) %>%
      mutate(yrterm = yrterm %>% as.character() ) %>%
      #group_by(crisno, sentence_number_int_aligned) %>%
      #  filter(!duplicated(value_markdown %>% trimws() )) %>% #note there are dupe markdowns for some reason
      #ungroup() %>%
      #mutate(crisno= ifelse(!is.na(lag(sentence_span_text)) & sentence_span_text==lag(sentence_span_text), "", crisno)) %>%
      #mutate(sentence_number_int_aligned= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_number_int_aligned)) %>%
      #mutate(sentence_span_text= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_span_text)) %>%
      #dplyr::select(C=crisno, S=sentence_number_int_aligned, ICB=sentence_span_text, ICBe=value_markdown) %>%
      #head(100) %>%
      flextable::as_flextable() %>%
      flextable::set_header_labels(values = 
          list(crisno = "#",
                crisis_text = "Title",
                yrtrig = "Start",
                yrterm = "End",
                n_sentence_number_int_aligned = "#S",
                n_events = "#E",
                coders_expert = "#Exp",
                coders_novice = "#Nov"#,
                #value_markdown = "Abstract"
               ) ) %>%
      flextable::add_header(values = 
          list(crisno = "Crisis",
                crisis_text = "Crisis",
                yrtrig = "Years",
                yrterm = "Years",
                n_sentence_number_int_aligned = "Counts",
                n_events = "Counts",
                coders_expert = "Coders",
                coders_novice = "Coders"#,
                #value_markdown = "Abstract"
               ) ) %>%
      flextable::merge_h(part = "header" ) %>%
      flextable::width( j = 1, width=0.1)  %>% 
      flextable::width( j = 2, width=0.5)  %>% 
      flextable::width( j = 3, width=0.1)  %>% 
      flextable::width( j = 4, width=0.1)  %>% 
      flextable::width( j = 5, width=0.1)  %>% 
      flextable::width( j = 6, width=0.1)  %>% 
      flextable::width( j = 7, width=0.1)  %>% 
      flextable::width( j = 8, width=0.1)  %>% 
      flextable::width( j = 9, width=2.0)  %>% 
      flextable::width( j = 10, width=1.0)  %>%
      flextable::width( j = 11, width=1.0)  %>%
      flextable::width( j = 12, width=3.0)  %>%

      flextable::fontsize(size = 9, part = "all") %>%
      #flextable::merge_v(j=1:8) %>%
      ftExtra::colformat_md(j = 9:12, part="body")
ft

```

```{r, eval=T}

saveRDS(ICBe_events_agreed_markdown, paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_markdown.Rds")) 
readr::write_tsv(ICBe_events_agreed_markdown, paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_markdown.tsv"))

saveRDS(ICBe_crises_markdown, paste0(here::here(), "/replication_data/out/ICBe_V1.1_crises_markdown.Rds")) 
readr::write_tsv(ICBe_crises_markdown, paste0(here::here(), "/replication_data/out/ICBe_V1.1_crises_markdown.tsv"))

```