### 2022-07-17 replication_paper/pnas_draft/paper.Rmd

---
title: "Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises"

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Rex W. Douglass
    affiliation: a,1,2
  - name: Thomas Leo Scherer
    affiliation: a
  - name: J. Andrés Gannon
    affiliation: b
  - name: Erik Gartzke
    affiliation: a
  - name: Jon Lindsay
    affiliation: c
  - name: Shannon Carcelli
    affiliation: d
  - name: Jonathan Wilkenfeld
    affiliation: d
  - name: David M. Quinn
    affiliation: e
  - name: Catherine Aiken
    affiliation: f
  - name: Jose Miguel Cabezas Navarro
    affiliation: g
  - name: Neil Lund
    affiliation: d
  - name: Egle Murauskaite
    affiliation: h
  - name: Diana Partridge
    affiliation: h

address:
  - code: a
    address: Department of Political Science, University of California, San Diego, CA, USA.
  - code: b
    address: Belfer Center for Science and International Affairs, Harvard Kennedy School, MA, USA.
  - code: c
    address: School of Cybersecurity and Privacy | Sam Nunn School of International Affairs, Georgia Institute of Technology, GA, USA.
  - code: d
    address: Department of Government and Politics, University of Maryland, College Park, MD, USA.
  - code: e
    address: Faculty Specialist, National Consortium for the Study of Terrorism and Responses to Terrorism (START), University of Maryland, College Park, MD, USA.
  - code: f
    address: Edmund A. Walsh School of Foreign Service, Georgetown University, Washington, DC, USA.
  - code: g
    address: Society and Health Research Center, Universidad Mayor, Santiago, Chile.
  - code: h
    address: ICONS Project // START,  University of Maryland, College Park, MD, USA.
  - code: i
    address: ICONS Project // START,  University of Maryland, College Park, MD, USA.

corresponding_author:
  code: 2
  text: "To whom correspondence should be addressed. E-mail: rexdouglass@gmail.com"

# For footer text
lead_author_surname: Douglass

#following https://docs.google.com/document/d/1aJxrQXYHW5U6By3KEAHrx1Iho6ioeh3ohNsRMwsoGPM/edit
author_contributions: |
  Conceptualization: R.W.D., E.G., J.L.; Methodology: R.W.D., T.L.S.; Software: R.W.D.;  Validation: R.W.D., T.L.S.;  Formal Analysis: R.W.D., T.L.S.;  Investigation: S.C., R.W.D., J.A.G., C.K., N.L., E.M., J.M.C.N., D.P., D.Q., J.W.;  Data Curation: R.W.D., D.Q., T.L.S., J.W.;  Writing - Original Draft: R.W.D., T.L.S.;  Writing - Review & Editing: R.W.D., J.A.G., E.G., T.L.S.;
  Visualization: R.W.D., T.L.S.;  Supervision: E.G.;  Project Administration: S.C., R.W.D., J.A.G., D.Q., T.L.S., J.W.;  Funding Acquisition: E.G., J.L.

## Remove this if not required
conflict_of_interest: |
  Please declare any conflict of interest here.

#Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
abstract: |
  How do international crises unfold? We conceptualize of international relations as a strategic chess game between adversaries and develop a systematic way to measure pieces, moves, and gambits accurately and consistently over a hundred years of history. We introduce a new ontology and dataset of international events called ICBe based on very high-quality corpus of narratives from the International Crisis Behavior (ICB) Project. We demonstrate that ICBe has higher coverage, recall, and precision than existing state of the art datasets and conduct two detailed case studies of the Cuban Missile Crisis (1962) and Crimea-Donbas Crisis (2014). We further introduce two new event visualizations (event icongraphy and crisis maps), an automated benchmark for measuring event recall using natural language processing (sythnetic narratives), and an ontology reconstruction task for objectively measuring event precision. We make the data, replication material, and visualizations of every historical episode available at a companion website www.crisisevents.org.

significance: |
  Countries routinely face crises that risk escalating into full scale war but we do not have systematic measurements of the progression of past crises and what moves and counter moves led to or helped avoid war. Instead policy makers typically rely on one or two historical analogies, chosen through ad hoc selection criteria, and described in unsystematic terms. This paper introduces a new scientific approach to measuring the step by step moves of international crises over the last hundred years, combining subject expertise with state of the art natural language processing and machine learning methods. It serves as a guide for constructing and evaluating large scale measurement collection in the social sciences.

acknowledgements: |
  We thank the ICB Project and its directors and contributors for their foundational work and their help with this effort. We make special acknowledgment of Michael Brecher for helping found the ICB project in 1975, creating a resource that continues to spark new insights to this day.  We thank the many undergraduate coders for their patience and dedication. Thanks to the Center for Peace and Security Studies and its membership for comments. Special thanks to Rebecca Cordell, Philip Schrodt, Zachary Steinert-Threlkeld, and Zhanna Terechshenko for generous feedback. Thank you to the cPASS research assistants that contributed to this project: Helen Chung, Daman Heer, Syeda ShahBano Ijaz, Anthony Limon, Erin Ling, Ari Michelson, Prithviraj Pahwa, Gianna Pedro, Tobias Stodiek, Yiyi 'Effie' Sun, Erin Werner, Lisa Yen, and Ruixuan Zhang. This project was supported by a grant from the Office of Naval Research [N00014-19-1-2491] and benefited from the Charles Koch Foundation's support for the Center for Peace and Security Studies.

keywords:
  - Diplomacy
  - War
  - Crises
  - International Affairs
  - Computational Social Science

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: ICBintro.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

#https://bookdown.org/yihui/bookdown/a-single-document.html
#https://stackoverflow.com/questions/52531637/knitr-rmarkdown-latex-how-to-cross-reference-figures-and-tables-in-2-different/52532269#52532269
#https://stackoverflow.com/questions/51595939/bookdown-cross-reference-figure-in-another-file
#https://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output/51521542#51521542
output: 
  bookdown::pdf_book:
    toc: no
    base_format: rticles::pnas_article
    latex_engine: xelatex

#Disabled several the caused errors
header-includes:
  - \usepackage{xr} \externaldocument{appendix} #External references from the appendix
#  - \usepackage{tinytex}
  - \usepackage[utf8]{inputenc}
  - \usepackage{pifont}
  - \usepackage{newunicodechar}
  - \newunicodechar{✓}{\ding{51}}
  - \newunicodechar{✗}{\ding{55}}
  - \usepackage{array}
  - \usepackage{ctable} # added for demo
  - \usepackage{natbib} #added for latex citation within huxtable
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{titlesec}
  - \usepackage[parfill]{parskip}
  - \usepackage{makecell}
  - \usepackage{graphicx}
#  - \usepackage{caption}
#  - \usepackage[capposition=top]{floatrow}
#  - \titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
#  - \titlespacing*{\subsubsection}{0pt}{3.25ex plus 1ex minus .2ex}{0ex plus .2ex}
#  - \DeclareUnicodeCharacter{00A0}{ }
  - \usepackage{setspace}
  - \usepackage{cellspace}
  - \setlength\cellspacetoplimit{0.8ex}
  - \renewcommand{\arraystretch}{0.8}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
#  - \usepackage[ruled,vlined,linesnumbered]{algorithm2e}
#  - \usepackage{subfigure}
  - \usepackage{subcaption}
  - \usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
  - \usepackage{float}
#  - \usepackage{eso-pic,graphicx,transparent}
#  - \usepackage[perpage]{footmisc}
  - \usepackage[algo2e]{algorithm2e}
#https://tex.stackexchange.com/questions/131646/algorithm2e-command-algorithm-already-defined
---

```{r, echo=F, cache=F, messages=F, results='hide', warnings=F, include=FALSE}
#International Crisis Behavior Events (ICBe) Dataset
knitr::opts_chunk$set(echo=F, cache=TRUE, messages=F, results='hide', warnings=F, include=FALSE) #fig.width = 8, fig.height = 8, message=F, warning=F, , results=F

```

```{r, eval=T, echo=F, cache=F}

#library(pacman)
library(googlesheets4)
library(tidyverse)
library(janitor)
library(flextable)
set_flextable_defaults(fonts_ignore=TRUE) #Warning: fonts used in `flextable` are ignored because the `pdflatex` engine is used and not `xelatex` or `lualatex`. You can avoid this warning by using the `set_flextable_defaults(fonts_ignore=TRUE)` command or use a compatible engine by defining `latex_engine: xelatex` in the YAML header of the R Markdown document.
library(ftExtra)
options(tidyverse.quiet = TRUE)
options(gargle_oauth_email = TRUE)

#install.packages('stargazer')
#install.packages('kableExtra')
#! sh: 1: pdflatex: not found
#sudo apt-get install texlive-latex-recommended texlive-fonts-recommended

#! LaTeX Error: File `letltxmacro.sty' not found.
#sudo apt-get install texlive-latex-extra

#install.packages('tinytex')
#tinytex::install_tinytex()

```

```{r, echo=F}
#Notes
#https://tex.stackexchange.com/questions/83440/inputenc-error-unicode-char-u8-not-set-up-for-use-with-latex
#- \DeclareUnicodeCharacter{00A0}{ }

```

```{r, include=FALSE}
# options(tinytex.verbose = TRUE)
```

```{r, echo=FALSE}

`%>%` <- magrittr::`%>%`

```

```{r, echo=FALSE}

icb_long_clean <- readRDS(file=paste0(here::here(), "/replication_data/temp/ICBe_V1.1_long_clean.Rds"))
codings_long_agreement <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_long_agreement.Rds") )

ICBe_events_agreed <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed.Rds")) 
dim(ICBe_events_agreed)
icb_events_long <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_long.Rds")) 
dim(icb_events_long)

#oh is it because break is reserved in R? lol
#icb_crises <- read_csv(file=paste0(here::here(), "/data_other_datasets/icb/icb1v14.csv")) %>% janitor::clean_names()
icb_crises  <- read.csv(paste0(here::here(), "/replication_data/in/icb1v14.csv")) %>% janitor::clean_names()

#somehow break of all things is a b roken name
#icb_actors <- read_csv(file=paste0(here::here(), "/data_other_datasets/icb/icb2v14.csv"))
icb_actors <- read.csv(paste0(here::here(), "/replication_data/in/icb2v14.csv"))

```

<!-- note: please start your introduction without including the word "Introduction" as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs. -->

If we could record every important interaction between countries in all of diplomacy, military conflict, and international political economy, how much unique information would this chronicle amount to, and how surprised would we be to see something new? In other words, what is the entropy of international relations? This record could in principle be unbounded, but the central conceit of social science is that there are structural regularities that limit what actors can do, their best options, and even which actors are likely to survive [@brecherInternationalStudiesTwentieth1999; @reiterShouldWeLeave2015]. If so, then these events can be systematically measured, and accordingly, massive effort is expended in social science attempting to record these regularities.^[See work on crises [@brecherCrisesWorldPolitics1982; @beardsleyInternationalCrisisBehavior2020], militarized disputes [@palmerMID5Dataset20112021; @giblerInternationalConflicts181620102018; @maozDyadicMilitarizedInterstate2019], wars [@sarkeesResortWar181620072010; @reiterRevisedLookInterstate2016], organized violence [@ralphsundbergUCDPGEDCodebook2016; @petterssonOrganizedViolence19892018], political violence [@raleighIntroducingACLEDArmed2010], sanctions [@felbermayrGlobalSanctionsData2020], trade [@barariDemocracyTradePolicy], and international agreements [@kinneDefenseCooperationAgreement2020; @owsiakInternationalBorderAgreements2018; @vabulasCooperationAutonomyBuilding2021], dispute resolution [@vabulasCooperationAutonomyBuilding2021; @frederickIssueCorrelatesWar2017], and diplomacy [@moyerWhatAreDrivers2020; @sechserMilitarizedCompellentThreats2011].] With improvements in natural language processing, more open-ended efforts have sought to capture entire unstructured streams of international events drawn from news reports.^[See @liComprehensiveSurveySchemabased2021; @haltermanExtractingPoliticalEvents2020; @brandtPhoenixRealTimeEvent2018; @boscheeICEWSCodedEvent2015; @hegreIntroducingUCDPCandidate2020; @grantOUEventData2017. On event-extraction from images and social-media see @zhangCASMDeepLearningApproach2019 and @steinert-threlkeldFutureEventData2019.] How close these efforts are to accurately measuring all or even most of what is essential in international relations is an open empirical question, one for which we provide new evidence here.

Our contribution is a high coverage ontology and event dataset for key historical episodes in 20th and 21st-century international relations. We develop a large, flexible ontology of international events with the help of both human coders and natural language processing. We apply it sentence-by-sentence to an unusually high-quality corpus of historical narratives of international crises [@brecherInternationalStudiesTwentieth1999; @brecherCrisisEscalationWar2000; @wilkenfeldInterstateCrisesViolence2000; @jamesWhatWeKnow2019; @iakhnisCrisesWorldPolitics2019]. The result is a new lower bound estimate of how much actually happens between states during pivotal historical episodes. We then develop several methods for objectively gauging how well these event codings reconstruct the information contained in the original narrative. We conclude by benchmarking our event codings against several current state-of-the-art event data collection efforts. We find that existing systems produce sequences of events that do not contain enough information to reconstruct the underlying historical episode. The underlying fine-grained variation in international affairs is unrecognizable through the lens of current quantification efforts.

This is a measurement paper that makes the following argument --- there is a real-world unobserved latent concept known as international relations, we propose a method for systematically measuring it, we successfully apply this method producing a new large scale set of measurements, those measurements exhibit several desirable kinds of internal and external validity, and those measurements out-perform other existing approaches. The article organizes that argument into eight sections: task definition; corpus; priors/existing state of the art; ICBe coding process; internal consistency; case study selection; recall; and precision. A final section concludes.

Task Definition {.unnumbered}
========================================

We consider the measurement task of abstracting discrete events about a historical episode in international relations. The easiest way to convey the task is with an example. Figure 1 shows a narrative account of the Cuban Missile Crisis (1962) alongside a mapping from each natural language sentence to discrete machine readable abstractive events. Formally, a historical episode, $H$, is demarcated by a period of time $[T_{start}, T_{end}] \in T$, a set of Players $p \in P$, and a set of behaviors they undertook during that time $b \in B$. International Relations, $IR$, is the system of regularities that govern the strategic interactions that world actors make during a historical episode, given their available options, preferences, beliefs, and expectations of choices made by others. We observe neither $H$ nor $IR$ directly. Rather the Historical Record, $HR$, produces documents $d \in D$ containing some relevant and true (as well as irrelevant and untrue) information about behaviors that were undertaken recorded in the form of unstructured natural language text. The task is to combine informative priors about $IR$ with an unstructured corpus $D$ to produce a series of structured discrete events, $e \in E$, that have high coverage, precision, and recall over what actually took place in history, $H$.

```{r cuban case study precision, eval=T, results='hide', echo=F, include=T, message=F, cache=F}

ICBe_events_agreed_markdown <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_markdown.Rds"))

temp <- ICBe_events_agreed_markdown %>%
        filter(crisno==196) %>%
        group_by(crisno, sentence_number_int_aligned) %>%
          filter(!duplicated(value_markdown %>% trimws() )) %>% #note there are dupe markdowns for some reason
        ungroup() %>%
        mutate(banding=sentence_number_int_aligned %% 2)
temp$is_subsstring <- NA
for(i in 1:nrow(temp)){
  temp$is_subsstring[i] <-   stringr::str_detect(temp$value_markdown %>% tolower() %>% str_replace_all('[^a-z0-9]',"") ,
                                                                                pattern = fixed(temp$value_markdown[i] %>% tolower() %>% str_replace_all('[^a-z0-9]',"") )
                                                                              ) %>% replace_na(0) %>%  sum()
}
temp <-temp %>%
        mutate(value_markdown = ifelse(is_subsstring<=1,value_markdown,'')) %>%
        group_by(sentence_number_int_aligned) %>%
          filter(value_markdown!='' | length(value_markdown)==1) %>%
        ungroup() %>%
        mutate(crisno= ifelse(!is.na(lag(sentence_span_text)) & sentence_span_text==lag(sentence_span_text), "", crisno)) %>%
        mutate(sentence_number_int_aligned= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_number_int_aligned)) %>%
        mutate(sentence_span_text= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_span_text))   #C=crisno, 
      #head(100) %>%
ft_cuba_precision <- temp %>%
                     dplyr::select(S=sentence_number_int_aligned, ICB=sentence_span_text, ICBe=value_markdown)  %>%
                      flextable::as_flextable() %>%
        flextable::set_header_labels(values =
        list(
              S = "S",
              ICB = "Natural Language Sentences (ICB Corpus)",
              ICBe = "Machine Readable Events (ICBe)"
             ) ) %>%

                      flextable::width( j = 1, width=0.25)  %>% 
                      #flextable::width( j = 2, width=0.25)  %>% 
                      flextable::width( j = 2, width=8.0)  %>% 
                      flextable::width( j = 3, width=8.0)  %>% 
                      flextable::fontsize(size = 9, part = "all") %>%
                      valign(j = 1:3, valign = "top", part = "body") %>%
                      flextable::line_spacing( space = 1.0, part = "all") %>% 
                      padding( padding = 1, part = "all") %>% 
                      bg( i = which( temp$banding==1 ) , part = "body", bg = "#EFEFEF", j=1:3) %>%
                      ftExtra::colformat_md(j = 3, part="body")
#ft_cuba_precision

ft_cuba_precision %>% save_as_image(path="/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/case_study_cuban_precision.png", zoom=1) # webshot = "webshot2"
temp <- NULL

```

<!--- REX Remember to put in a Legend for the Icons ---->

\clearpage
\onecolumn
\begin{figure}[H]
\caption{Case Study 1: Cuban Missile Crisis (1962) - ICB Narrative vs. ICBe Events \label{fig:case_study_cuban_precision}}
\centering{\includegraphics[ height=22cm]{"/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/case_study_cuban_precision.png"}}
\end{figure}
\clearpage
\twocolumn

Corpus  {.unnumbered}
========================================

For our corpus, $D$, we select a set of unusually high-quality historical narratives from the International Crisis Behavior (ICB) project ($n=471$)(SI Appendix, Table A1)[@brecherInternationalCrisisBehavior2017; @brecher_study_1997]. Their domain is 20th and 21st-century crises, defined as a change in the type, or an increase in the intensity, of disruptive interaction with a heightened probability of military hostilities that destabilizes states' relationships or challenges the structure of the international system [@brecherCrisesWorldPolitics1982].^[On near crises see @iakhnisCrisesWorldPolitics2019.] Crises are a significant focus of detailed single case studies and case comparisons because they provide an opportunity to examine behaviors in IR short of, or at least prior to, full conflict  [@holsti1914Case1965; @paigeKoreanDecisionJune1968; @allisonEssenceDecisionExplaining1971; @snyderConflictNationsBargaining1977; @gavinHistorySecurityStudies2014; @georgeDeterrenceAmericanForeign1974; @brecherCrisesWorldPolitics1982; @gaddisExpandingDataBase1987; @brecherPatternsCrisisManagement1988]. Case selection was exhaustive based on a survey of world news archives and region experts, cross-checked against other databases of war and conflict, and non-English sources [@kangUSBiasStudy2019; @brecherInternationalCrisisBehavior2017 p. 59]. Each narrative was written by consensus by a small number of scholars, using a uniform coding scheme, with similar specificity [@hewittEngagingInternationalData2001]. The corpus is unique in IR because it is designed to be used in a downstream quantitative coding project.

Prior Beliefs about IR, Ontological Coverage, and the Existing State of the Art {.unnumbered}
========================================

Next we draw informative prior beliefs about the underlying process of IR that we expect to govern behavior during historical episodes and their conversion to the historical record. We organize our prior beliefs along two overarching axes, summarized in detail by Table 1.

The first axis (rows) represents the types of information we expect to find in IR and forms the basis for our proposed ontology. We employ a metaphor of a chess game, with players (polities, rebel groups, IGOs, etc.), pieces (military platforms, civilians, domains), and behaviors (think, say, do). Precise sequencing is required to capture gambits (sequences of moves) and outcomes (victory, defeat, peace, etc.), while precise geo-coding is required to understand the chessboard (medium of conflict). We find `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("actor")) %>% pull(value_normalized) %>% unique() %>% length()` actors and 117 different behaviors, and provide a full codebook in the online material.^[See the Github Repository [ICBEventData](https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$).]

We base our informed priors primarily on two sources of information. The first is the extensive existing measurement efforts of IR which we provide citations to alongside each concept. Second, we performed preliminary natural language processing of the corpus and identified named entities and behaviors mentioned in the text. Verbs were matched to the most likely definition found in Wordnet [@millerWordNetLexicalDatabase1995], tallied, and then aggregated into a smaller number hypernyms balancing conceptual detail and manageable sparsity for human coding (SI Appendix, Table A2).

The second axis (columns) compares the very high ontological coverage of ICBe to existing state of the art systems in production and with global coverage. They begin with our contribution ICBe, alongside other event-level datasets including CAMEO dictionary lookup based systems (Historical Phoenix [@althausClineCenterHistorical2019]; ICEWS [@boscheeICEWSCodedEvent2015; @hegreIntroducingUCDPCandidate2020]; Terrier [@grantOUEventData2017]), the Militarized Interstate Disputes Incidents dataset, and the UCDP-GED dataset [@ralphsundbergUCDPGEDCodebook2016; @petterssonOrganizedViolence19892018; @sundbergIntroducingUCDPGeoreferenced2013].^[Additional relevant but dated or too small of an overlap in domain include BCOW [@lengMilitarizedInterstateCrises1988], WEIS [@mcclellandWorldEventInteraction1978], CREON [@hermannComparativeResearchEvents1984], CASCON [@bloomfieldCASCONIIIComputeraided1989], SHERFACS [@shermanSHERFACSCrossParadigmHierarchical2000], Real-Time Phoenix [@brandtPhoenixRealTimeEvent2018], and COfEE [@balaliCOfEEComprehensiveOntology2021] (see histories in @merrittMeasuringEventsInternational1994 and @schrodtTwentyYearsKansas2006).] The final set of columns compares episode-level datasets beginning with the original ICB project [@brecherInternationalCrisisBehavior; @brecherCrisesWorldPolitics1982; @beardsleyInternationalCrisisBehavior2020]; the Militarized Interstate Disputes dataset [@palmerMID5Dataset20112021; @giblerInternationalConflicts181620102018;@braithwaiteMIDLOCIntroducingMilitarized2010;@braithwaiteCodebookMilitarizedInterstate2009], and the Correlates of War [@sarkeesResortWar181620072010]. With the exception of large scale CAMEO dictionary based systems, the existing state of the art quantitative datasets ignore the vast majority of the information content found in international relations.^[See [@balaliCOfEEComprehensiveOntology2021] for a recent review of ontological depth and availability of Gold Standard example text.]

\clearpage
\onecolumn

```{r litreview, eval=T, echo=F, results='asis', include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

#target_file <- paste0(here::here(),"/replication_paper/data/in/icbe_litreview_trees_sentences.xlsx")
#lit_review_pull    <- readxl::read_excel(target_file, sheet="DatasetsComparisonLitReview")

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
lit_review_pull <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/14RHRKgQGHj7Zr4_xt4fq2bIE7mDNayLKj89tYmOOfHQ/edit?usp=sharing", 
                                sheet="DatasetsComparisonLitReview") #%>% janitor::clean_names() 
dim(lit_review_pull)

library(flextable)
library(ftExtra)
library(tidyverse)
flextable::set_flextable_defaults(fonts_ignore=TRUE)

#https://davidgohel.github.io/flextable/reference/knit_print.flextable.html#chunk-options
# NOTE: If one of the citations is degenerate this will throw an error when knitting that is annoying to track down. You have to remove cite by cite to see which one it is.
set_flextable_defaults(fonts_ignore=TRUE)

lit_review_alt <- lit_review_pull %>% 
                  mutate(Literature =   paste0(" [",Literature %>% str_replace("; {0,}$","") ,"]")) %>%
                  mutate(Concept= Concept %>% paste0(Literature)) %>%
                  mutate(Concept = Concept %>% str_replace(fixed("[NA]"),"") ) %>%
                  dplyr::select(-Codebook, -`Cameo Codes`, -`ICB Codes`, -Literature, -`ICB Corpus`) 
                  #dplyr::filter(is.na(group))

n=nrow(lit_review_alt)
k=ncol(lit_review_alt)
horizontal = c(1,k)
r_intro= lit_review_pull$...1 %in% "Domain" %>% sum() #changed to domain
r_players= ( lit_review_pull$...1 %in% "Players" %>% sum() ) + r_intro
r_pieces= ( lit_review_pull$...1 %in% "Pieces" %>% sum() ) + r_players
r_think= ( lit_review_pull$...1 %in% "Think" %>% sum() ) + r_pieces
r_say= ( lit_review_pull$...1 %in% "Say" %>% sum() ) + r_think
r_unarmed= ( lit_review_pull$...2 %in% "Unarmed" %>% sum() ) + r_say
r_armed= ( lit_review_pull$...2 %in% "Armed" %>% sum() ) + r_unarmed
ft <- NULL

library(officer)
big_border = fp_border(color="black", width = 2)

bands=which( 1:nrow(lit_review_alt) %% 2 == 1)
litreview_ft_pnas <- lit_review_alt %>%
      flextable::as_flextable() %>%

      #add_header( 
      #        #"ICB Corpus"="Corpus",
      #        "ICBe"="Ours",
      #        "Phoenix"="CAMEO Dictionary\nSystems",
      #        "Terrier"="CAMEO Dictionary\nSystems",
      #        "ICEWs"="CAMEO Dictionary\nSystems",
      #        "MIDs Events"="Disputes",
      #        "UCDP-GED"="Violence",
      #        "ICB"="Crises",
      #        "COW"="Wars",
      #        "MIDs"="Disputes",
      #        top = T ) %>%

      add_header( 
                  #"ICB Corpus"="Corpus",
                  "ICBe"="Events Datasets",
                  "Phoenix"="Events Datasets",
                  "Terrier"="Events Datasets",
                  "ICEWs"="Events Datasets",
                  "MIDs Incidents"="Events Datasets",
                  "UCDP-GED"="Events Datasets",
                  "ICB"="Episodes Datasets",
                  "COW"="Episodes Datasets",
                  "MIDs"="Episodes Datasets",
                  top = T ) %>%

      set_header_labels(
        values = list("...1" = "" , "...2" = "" ) 
        ) %>%

      merge_h(part = "header") %>% #it's this merge that's killing us

      #Global Borders
      flextable::border_remove() %>% 
      #ok so because of our horizontal merging of headers, we have to manually place a right bar at j=10 2 short of the full width
      flextable::border_outer( part="all", border = big_border ) %>% #this will have a gap
      border(j=10:k, i=1 , part="header", border.right = big_border ) %>% #this fills the gap
      #There's another gap on the bottom left
      border(j=1, i=(n-15):n , part="body", border.bottom =  big_border ) %>% #this fills the gap
      border(j=2, i=(n-5):n , part="body", border.bottom =  big_border ) %>% #this fills the gap

      flextable::line_spacing( space = 1.0, part = "all") %>% #i=1:nrow(lit_review_alt),
      flextable::padding(padding = 0, part = "body") %>%
      flextable::fontsize(size = 7, part = "all") %>%
      #Banding
      bg( i = bands  , j=3:k, part = "body", bg = "#EFEFEF") %>%

      flextable::rotate(i=2, , j=3:k, rotation="tbrl",part="header") %>%
      flextable::align(align = "center", part = "body") %>%
      flextable::align(align = "center", part = "header")   %>%

      flextable::colformat_char(
          j = 1:ncol(lit_review_alt),
          na_str = "",
          prefix = "",
          suffix = ""
      ) %>%
      #Domains
      hline(i=c(r_intro), border = NULL, part = "body") %>% 
      merge_at(i = 1:r_intro, j = 1:2, part = "body") %>%  
      merge_at(i = 4, j = 5:7, part = "body") %>%
      merge_at(i = 5, j = 5:7, part = "body") %>%
      merge_at(i = 6, j = 5:7, part = "body") %>%
      merge_at(i = 7, j = 5:7, part = "body") %>%

      #players
      hline(i=c(r_players), border = NULL, part = "body") %>% 
      merge_at(i = (r_intro+1):r_players, j = 1:2, part = "body") %>%  
      #pieces
      hline(i=c(r_pieces), border = NULL, part = "body") %>% 
      merge_at(i = (r_players+1):r_pieces, j = 1:2, part = "body") %>%  
      #think
      hline(i=c(r_think), border = NULL, part = "body") %>% 
      merge_at(i = (r_pieces+1):r_think, j = 1:2, part = "body") %>%  
      #say
      hline(i=c(r_say), border = NULL, part = "body") %>% 
      merge_at(i = (r_think+1):r_say, j = 1:2, part = "body") %>%
      #Do
      merge_at(i = (r_say+1):r_armed, j = 1, part = "body") %>%

      #n_unarmed
      hline(i=c(r_unarmed), border = NULL, part = "body") %>% 
      merge_at(i = (r_say+1):r_unarmed, j = 2, part = "body") %>%  
      #n_armed
      #hline(i=c(r_armed), border = NULL, part = "body") %>% 
      merge_at(i = (r_unarmed+1):r_armed, j = 2, part = "body") %>%

      #set_table_properties(layout="autofit")  %>%
      vline( i = 1:2, j = c(3,9), border = NULL, part = "header") %>%
      vline( i = 1:(n), j = 1:2, border = NULL, part = "body")  %>%
      vline( i = 1:(n), j = c(3,9), border = NULL, part = "body")  %>%

      flextable::width( j = 3, width=3) %>% 
      flextable::width( j = (4):k, width=0.25) %>%
      flextable::line_spacing( space = 0.5, part = "body") %>% #i=1:nrow(lit_review_alt),
      flextable::padding(padding = 0, part = "body") %>%

      flextable::rotate(j=1, rotation="tbrl",part="all") %>%
      flextable::rotate(j=2, rotation="tbrl",part="all") %>%
      flextable::width( j = 1:2, width=0.1)  %>%
      ftExtra::colformat_md(j = 3, part="body")   #Make sure this goes last or it'll get overwritten by the colformat above

#https://cran.r-project.org/web/packages/ftExtra/vignettes/format_columns.html
#litreview_ft_pnas

#litreview_ft_pnas  %>% saveRDS(paste0(here::here(), '/replication_paper/tables/litreview_ft_pnas.Rds'))
litreview_ft_pnas %>% 
  ftExtra::colformat_md(j = 3, part="body") %>%
  flextable::set_caption(caption = "Ontological coverage of ICBe versus existing State of the Art") %>%
  # flextable::set_caption("Prior Beliefs, State of the Art, and Existing Data \\label{tbl:litreview}")  %>% 
  flextable::flextable_to_rmd()

```

\clearpage
\twocolumn

```{r , eval=T, results='hide', echo=F, include=T, message=F, cache=F}

n_crisis <- ICBe_events_agreed %>% dplyr::select(crisno) %>% unique() %>% nrow()
n_sentences <- ICBe_events_agreed %>% dplyr::select(crisno, sentence_number_int_aligned) %>% unique() %>% nrow()
#n_events <- ICBe_events_agreed %>% dplyr::filter(event_number_int %in% c(1,2,3)) %>% nrow()
#n_events_cuba <- ICBe_events_agreed %>% dplyr::filter(crisno == 196 & event_number_int %in% c(1,2,3)) %>% nrow()
n_coders <- icb_long_clean %>% dplyr::select(email_id) %>% unique() %>% nrow()

avg_coders_per_crisis <- icb_long_clean %>% dplyr::filter(crisno != 196) %>% dplyr::mutate(cris_coder = paste(crisno, email_id)) %>% dplyr::select(crisno, email_id) %>% unique() %>% dplyr::count(crisno) %>% dplyr::select(n) %>% lapply(mean)

avg_coders_per_sentence <- icb_long_clean %>% dplyr::filter(crisno != 196) %>% dplyr::mutate(cris_sent = paste(crisno, sentence_number_int_aligned)) %>% dplyr::select(cris_sent, email_id) %>% unique() %>% dplyr::count(cris_sent) %>% dplyr::select(n) %>% lapply(mean)

```

ICBe Coding Process {.unnumbered}
========================================

The ICBe ontology follows a hierarchical design philosophy where a smaller number of significant decisions are made early on and then progressively refined into more specific details [@brustIntegratingDomainKnowledge2020].^[This process quickly focuses the coder on a smaller number of relevant options while also allowing them to apply multiple tags if the sentence explicitly includes more than one or there is insufficient evidence to choose only one tag. The guided coding process also allows for the possibility that earlier coarse decisions have less error than later fine-grained decisions.] Each coder was instructed to first thoroughly read the full crisis narrative and then presented with a custom graphical user interface (SI Appendix, Fig. B1). Coders then proceeded sentence by sentence, choosing the number of events (0-3) that occurred, the highest behavior (thought, speech, or activity), a set of players ($P$), whether the means were primarily armed or unarmed, whether there was an increase or decrease in aggression (uncooperative/escalating or cooperative/de-escalating), and finally one or more non-mutually exclusive specific activities. Some additional details like location and timing information was always collected while other details were only collected if appropriate, e.g. force size, fatalities, domains, units, etc. A unique feature of the ontology is that thought, speech, and do behaviors can be nested into combinations, e.g. an offer for the U.S.S.R. to remove missiles from Cuba in exchange for the U.S. removing missiles from Turkey. Through compounding, the ontology can capture what players were said to have known, learned, or said about other specific fully described actions.

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.^[Expert coders were graduate students or postgraduates who collaboratively developed the ontology and documentation for the codebook. Undergraduate coders were students who engaged in classroom workshops.] For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=`r icb_long_clean %>% nrow() %>% scales::comma()`), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=`r ICBe_events_agreed %>% nrow() %>% scales::comma()`), using a consensus procedure (SI Appendix, Algorithm B2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts.

```{r}
#We calculate means in the SI appendix on agreement
means_group_any <- codings_long_agreement %>% filter(value_normalized!='') %>% 
                   group_by(crisno,sentence_number_int_aligned, varname_normalized) %>% arrange(desc(percent_expert)) %>% filter(row_number()==1) %>% ungroup() %>% 
                   dplyr::select(Experts=percent_expert,Novices=percent_novice) %>% summarise_all(mean)

means_group_keep <- codings_long_agreement %>% filter(value_normalized!='') %>% 
                    filter(algo1==1) %>%
                    group_by(crisno,sentence_number_int_aligned, varname_normalized) %>% arrange(desc(percent_expert)) %>% filter(row_number()==1) %>% ungroup() %>% 
                    dplyr::select(Experts_keep=percent_expert,Novices_keep=percent_novice) %>% summarise_all(mean)
```

Internal Consistency {.unnumbered}
========================================

We evaluate the internal validity of the coding process in several ways. For every tag applied we calculate the observed intercoder agreement as the percent of other coders who also applied that same tag (SI Appendix, Fig. B3). Across all concepts, the Top 1 Tag Agreement was low among novices (`r means_group_any$Novices %>% round(2) %>% scales::percent()`), moderate for experts (`r means_group_any$Experts %>% round(2) %>% scales::percent()`), and high (`r means_group_keep$Experts_keep %>% round(2) %>% scales::percent()`) following the consensus screening procedure.

We attribute the remaining disagreement primarily to three sources. First, we required coders to rate their confidence which was observed to be low for 20% of sentences- half due to a mismatch between the ontology and the text ("survey doesn't fit event"-45%) and half due to a lack of information or confused writing in the source text ("more knowledge needed"-40%, "confusing sentence"-6%). Observed disagreement varied predictably with self reported confidence (SI Appendix, Fig. B4). Second, as intended agreement is higher (75-80%) for questions with fewer options near the root of the ontology compared to agreement for questions near the leafs of the ontology (50%-60%). Third, individual coders exhibiting nontrivial coding styles, e.g. some more expressive applying many tags per concept while others focused on only the single best match. We further observed unintended synonymity, e.g. the same information can be framed as either a threat to do something or a promise not to do something.

Case Study Selection {.unnumbered}
========================================

The remaining two qualities we seek to measure are recall and precision of ICBe events in absolute terms and relative to other existing systems. We provide full ICB narratives, ICBe coding in an easy to read icongraphic form, and a wide range of visualizations for every case on the companion website. In this paper, we focus on two deep case studies. The first is the Cuban Missile Crisis (Figure 1) which took place primarily in the second half of 1962, involved the United States, the Soviet Union, and Cuba, and is widely known for bringing the world to the brink of nuclear war (hereafter Cuban Missiles). The second is the Crimea-Donbas Crisis (SI Appendix Figure D1) which took place primarily in 2014, involved Russia, Ukraine, and NATO, and within a decade spiraled into a full scale invasion (herafter Crimea-Donbas). Both cases involve a superpower in crisis with a neighbor, initiated by a change from a friendly to hostile regime, with implications for economic and military security for the superpower, risked full scale invasion, and eventually invited intervention by opposing superpowers. We choose these cases because they are substantively significant to 20th and 21st century international relations, widely known across scientific disciplines and popular culture, and are sufficiently brief to evaluate in depth.

Recall {.unnumbered}
========================================

Recall measures the share of desired information recovered by a sequence of coded events, $Pr(E|H)$, and is poorly defined for historical episodes. First, there is no genuine ground truth about what occurred, only surviving texts about it. Second, there is no \textit{a priori} guide to what information is necessary detail and what is ignorable trivia. History suffers from what is known as the Coastline Paradox [@mandelbrotFractalGeometryNature1983] --- it has a fractal dimension greater than one such that the more you zoom in the more detail you will find about individual events and in between every two discrete events. The ICBe ontology is a proposal about what information is important, but we need an independent benchmark to evaluate whether that proposal is a good one and that allows for comparing proposals from event projects that had different goals. We need a yardstick for history.

Our strategy for dealing with both problems is a plausibly objective yardstick called a synthetic historical narrative. For both case studies, we collect a large diverse corpus of narratives spanning timelines, encyclopedia entries, journal articles, news reports, websites, and government documents. Using natural language processing (fully described in SI Appendix, Algorithm C1), we identify details that appear across multiple accounts. The more accounts that mention a detail, the more central it is to understanding the true historical episode. The theoretical motivation is that authors face word limits which force them to pick and choose which details to include, and they choose details which serve the specific context of the document they are producing. With a sufficiently large and diverse corpus of documents, we can vary the context while holding the overall episode constant and see which details tend to be invariant to context. Intuitively, a high quality event dataset should have high recall for context invariant details both because of their broader relevance and also because they are easier to find in source material.

Synthetic historical narratives for Cuban Missiles (51 events drawn from 2020 documents) and Crimea-Donbas (30 events drawn from 971 documents) appear in Figure 2. Each row represents a detail which appeared in at least five documents along with an approximate start date, a hand written summary, the number of documents it was mentioned in, and whether it could be identified in the text of our ICB corpus, in our ICBe events, or any of the competing systems.

From them, we draw several stylized facts. First, there is substantial variation in which details any one document will choose to include. Our ground truth ICB narratives included 17/51 and 23/30 of the events from the synthetic narrative, while including other details that are not in the synthetic narrative. Second, mentions of a detail across accounts is exponentially distributed with context invariant details appearing dozens to hundreds of times more than context dependent details. Third, crisis start and stop dates are arbitrary and the historical record points to many precursor events as necessary detail for understanding later events, e.g. the U.S. was in a \textit{de facto} grey scale war with Cuba before it invited Soviet military protection [@cormacGreyNewBlack2018] and Ukraine provided several security guarantees to Russia that were potentially undone, e.g. a long term lease on naval facilities in Crimea. Fourth, we find variation between the two cases. Cuban Missiles has a cleaner canonical end with the Soviets agreeing to withdraw missiles while Crimea-Donbas meekly ends with a second cease fire agreement (Minsk II) but continued fighting. The canonical narrative of Cuban Missile also includes high level previously classified details, while the more recent Crimea-Donbas case reflects primarily public reporting.

We find substantive variation in recall across systems. Recall for each increases in the number of document mentions which is an important sign of validity for both them and our benchmark. The one outlier is Phoenix which is so noisy that it's flat to decreasing in mentions. The two episode level datasets have very low coverage of contextual details. The two other dictionary systems ICEWs and Terrier have high coverage, with ICEWs outperforming Terrier. ICBe strictly dominates all of the systems but ICEWs in recall though we note that the small sample sizes mean these systems should be considered statistically indistinguishable. Importantly our corpus of ICB narratives has very high recall of frequently mentioned details giving us confidence in how those summaries were constructed, and ICBe lags only slightly behind showing that it left very little additional information on the table.

```{r cuban case study recall, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

#
#Lit review
#child=c('ICBEdataset_paper_litreviewtable.Rmd')
library(tidyverse)
library(flextable)
library(ftExtra)
flextable::set_flextable_defaults(fonts_ignore=TRUE)

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
case_study_cuban <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1NuRWFB1HEbQbJu_JoEzJ9WYk7kfwwCryF7wBfKPT6uc/edit?usp=sharing", 
                                sheet="final") %>% janitor::clean_names() 
dim(case_study_cuban)
#case_study_cuban$date <- case_study_cuban$date %>% lapply(FUN=function(x) { ifelse(is.null(x), "", x[[1]] %>% as.character() )   }   ) %>% unlist() %>% as.Date()

#Replace emoticons with images
replace_emoticons <- function(x){
  x %>% str_replace_all('\U0001f1e8\U0001f1fa',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q241.png) ") %>%
        #https://emojiterra.com/flag-for-united-states/
        str_replace_all('\U0001f1fa\U0001f1f8',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q30.png) ") %>%
        #https://emojiterra.com/flag-for-russia/
        str_replace_all('\U0001f1f7\U0001f1fa',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q15180.png) ") %>% 
        #Turkey
        str_replace_all('\U0001f1f9\U0001f1f7',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q43.png) ") %>% 
        #U.N.
        str_replace_all('\U0001f1fa\U0001f1f3',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q1065.png) ")
}

temp1 <- case_study_cuban %>% 
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(documents = documents %>% as.numeric()) %>%
                mutate(Rank=rank(-documents, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank, Event=event, Docs=documents, icb_narrative=icb_narrative, ICBe=ic_be, ICB=icb, MIDs=mids, phoenix=phoenix) %>% 
                mutate(Event= Event %>%  replace_emoticons) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(ICB = ifelse(!is.na(ICB) & !ICB %in% 0, "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0, "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0, "✓", NA)) %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0, "✓", NA)) %>%

                mutate(D=ifelse(!is.na(lag(D)) & D==lag(D) & M==lag(M),'',D )) %>%
                mutate(M=ifelse(!is.na(lag(M)) & M==lag(M),'',M )) %>%
                mutate(Y=ifelse(!is.na(lag(Y)) & Y==lag(Y),'',Y )) %>%
                mutate(Y=ifelse(duplicated(Y), NA, Y)) %>%
                mutate(YMD=paste(Y,
                                 M %>% str_pad(2, pad = "0") ,
                                 D %>% str_pad(2, pad = "0"), 
                                 sep="-") %>% trimws() %>%
                                 str_replace_all("^NA-","") %>%  #There's like a carriage return and tab in there somehow!
                                 str_replace_all("^NA-","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-","") 
                       ) %>%
                dplyr::select(-Y,-M,-D) %>%
                relocate(YMD, .before = Rank) %>%
                dplyr::select(-Rank)
dim(temp1)

#banding <- temp %>% mutate(date=as.Date(paste(Y,M,D, sep="-")) %>% as.numeric() %>% as.factor() %>% as.numeric() ) %>% pull(date) %% 2
banding <- 1:nrow(temp1) %% 2
k=ncol(temp1)
ft <- temp1   %>%
      flextable::as_flextable() %>% 
      flextable::set_header_labels(values =
        list(
              Event = "Ground\nTruth\nEvents",
              icb_narrative = "ICB Corpus",
              ICB = "ICB",
              MIDs = "MIDs",
              phoenix = "Phoenix"
             ) ) %>%
      flextable::width( j = 1, width=0.6)  %>% 
      #flextable::width( j = 2, width=0.025)  %>%
      #flextable::width( j = 3, width=0.025)  %>%
      #flextable::width( j = 5, width=0.025) %>%
      flextable::width( j = 2, width=3.0)  %>%
      flextable::width( j = 3, width=0.03) %>%
      flextable::width( j = 4:k, width=0.15) %>% #icb
      #merge_v( j = 1,  part = "body", combine = T) %>%
      #merge_v( j = 3,  part = "body", combine = F) %>%
      #merge_v( j = 4,  part = "body", combine = F) %>%

      flextable::rotate(i=1, rotation="tbrl", part="header") %>% 
      align(i = 1,  align = "center", part = "header") %>%
      valign(j = 1:2, valign = "center", part = "all") %>%
      align(i = NULL, j = 1, align = "right", part = "body")  %>%
      align(i = NULL, j = 2, align = "left", part = "body")  %>%
      #align(i = NULL, j = 2, align = "center", part = "body")  %>%
      #align(i = NULL, j = 3, align = "left", part = "body")  %>%
      align(i = NULL, j = 3:k, align = "center", part = "body")  %>%
      flextable::line_spacing( space = 1.0, part = "all") %>% #i=1:nrow(lit_review_alt),
      flextable::padding(padding = 0, part = "body") %>%
      flextable::fontsize(size = 7, part = "all") %>%
      #bg( i = which( 1:nrow(temp) %% 2 == 1) , part = "body", bg = "#EFEFEF", j=1:k) %>%
      bg( i = which( banding==1 ) , part = "body", bg = "#EFEFEF", j=1:k) %>%
      #hline(i=c(26,46), j=3:k,  border = NULL, part = "body") %>%
      vline(j=c(1,2,3,4,5,6,7),  border = NULL, part = "all") %>%
      hline(i=c(26,46)) %>%
      ftExtra::colformat_md(j = 2, part="body") 
      #ftExtra::colformat_md(j = 6, part="body")

ft   %>% 
    #flextable::flextable_to_rmd()
    save_as_image(path="/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/case_study_cuban_recall.png", zoom=1, webshot = "webshot2") # webshot = "webshot2"

#Ok rather than try to match up we're going to do this by hand now

```

```{r, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
case_study_crimea <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1YesAx1CkYCgrEi_WVJ9aho60HesWA6p-3XMma9IzBrI/edit?usp=sharing", 
                                sheet="final") %>% janitor::clean_names() 
dim(case_study_crimea)

#Replace emoticons with images
replace_emoticons_crimea <- function(x){
      x %>%
            #https://emojiterra.com/flag-for-united-states/
            str_replace_all('\U0001f1fa\U0001f1f8',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q30.png) ") %>%
            #https://emojiterra.com/flag-for-russia/
            str_replace_all('\U0001f1f7\U0001f1fa',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q159.png) ") %>% 
            #Ukraine
            str_replace_all('\U0001f1fA\U0001f1e6',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q212.png) ") %>% 
            #
            str_replace_all('Crimea',  " ![](/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_data/in/flags_small//Q15966495.png) ")
}

rex_date <- function(x){
  a <- x[1] %>% as.character() %>% unlist()
  if(is.null(a)){return('')}
  return(a)
}
case_study_crimea$date <- sapply(sapply(case_study_crimea$date, as.character), rex_date)
temp2 <- case_study_crimea %>% 
                filter(keep %in% 1) %>%
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(n_sources = n_sources %>% as.numeric()) %>%
                mutate(Rank=rank(-n_sources, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank,  Event=event, Docs=n_sources, 
                              icb_narrative=icb_narrative, ICBe=ic_be, MIDs=mids, phoenix, terrier, icews ) %>%  #, ICB=icb, MIDs=mids, phoenix=phoenix
                mutate(Event= Event %>%  replace_emoticons_crimea) %>%
                arrange(Y %>% as.numeric() ,M %>% as.numeric()  ,D %>% as.numeric() ) %>%
                mutate(D=ifelse(!is.na(lag(D)) & D==lag(D) & M==lag(M),'',D )) %>%
                mutate(M=ifelse(!is.na(lag(M)) & M==lag(M),'',M )) %>%
                mutate(Y=ifelse(!is.na(lag(Y)) & Y==lag(Y),'',Y )) %>%
                mutate(Y=ifelse(duplicated(Y), NA, Y)) %>%
                mutate(YMD=paste(Y,
                                 M %>% str_pad(2, pad = "0") ,
                                 D %>% str_pad(2, pad = "0"), 
                                 sep="-") %>% trimws() %>%
                                 str_replace_all("^NA-","") %>%  #There's like a carriage return and tab in there somehow!
                                 str_replace_all("^NA-","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-","") 
                       ) %>%
                dplyr::select(-Y,-M,-D) %>%
                relocate(YMD, .before = Rank) %>%
                dplyr::select(-Rank) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(icews = ifelse(!is.na(icews) & !icews %in% 0 , "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0  , "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0  , "✓", NA)) %>%
                mutate(terrier = ifelse(!is.na(terrier) & !terrier %in% 0  , "✓", NA))  %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0  , "✓", NA))

dim(temp2)

banding <- 1:nrow(temp2) %% 2
k=ncol(temp2)
ft <- temp2  %>%
      flextable::as_flextable() %>% 
      flextable::set_header_labels(values =
        list(
              Event = "Ground\nTruth\nEvents",
              icb_narrative = "ICB Corpus",
              ICBe = "ICBe",
              MIDs="MIDs",
              icews_hash = "ICEWS"
             ) ) %>%
      flextable::width( j = 1, width=0.6)  %>% 
      flextable::width( j = 2, width=3.0)  %>%
      flextable::width( j = 3:k, width=0.15) %>% #icb
      flextable::rotate(i=1, rotation="tbrl", part="header") %>% 
      align(i = 1,  align = "center", part = "header") %>%
      valign(j = 1:2, valign = "center", part = "all") %>%
      align(i = NULL, j = 1, align = "right", part = "body")  %>%
      align(i = NULL, j = 2, align = "left", part = "body")  %>%
      align(i = NULL, j = 3:k, align = "center", part = "body")  %>%
      flextable::line_spacing( space = 1.0, part = "all") %>% #i=1:nrow(lit_review_alt),
      flextable::padding(padding = 0, part = "body") %>%
      flextable::fontsize(size = 7, part = "all") %>%
      #bg( i = which( 1:nrow(temp) %% 2 == 1) , part = "body", bg = "#EFEFEF", j=1:k) %>%
      bg( i = which( banding==1 ) , part = "body", bg = "#EFEFEF", j=1:k) %>%
      #hline(i=c(26,46), j=3:k,  border = NULL, part = "body") %>%
      vline(j=1:k,  border = NULL, part = "all") %>%
      #hline(i=c(26,46)) %>%
      ftExtra::colformat_md(j = 2, part="body") 
      #ftExtra::colformat_md(j = 6, part="body")

ft   %>% 
    #flextable::flextable_to_rmd()
    save_as_image(path="/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/case_study_crimea_recall.png",
                  zoom=1, webshot = "webshot2") # webshot = "webshot2"

```

```{r, recall combined, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

temp3 <- temp1 %>% mutate(row=row_number()) %>% full_join(temp2 %>% mutate(row=row_number()), by='row') %>% 
         mutate(row="") #we're going to use this as our vertical break

n_left <- (!is.na(temp3$Event.x) ) %>% sum()
n_right <- (!is.na(temp3$Event.y) ) %>% sum()

k=ncol(temp3)
kleft=ncol(temp1)
kright=ncol(temp2)
left_end=kleft
right_start=left_end+2
left=c(1:left_end)
right=c(right_start:k)
horizontal = c(left,right)
dates=c(1,right_start)
events = dates + 1
checkmarks = c((events[1]+1):left_end , (events[2]+1):k)

library(officer)
big_border = fp_border(color="black", width = 2)

banding1 <- 1:nrow(temp1) %% 2
banding2 <- 1:nrow(temp2) %% 2
ft <- temp3  %>%
      flextable::as_flextable() %>%

      flextable::line_spacing( space = 1.0, part = "all") %>% # 
      flextable::padding(padding = 0, part = "body") %>%
      flextable::fontsize(size = 7, part = "all") %>%
      bg( i = which( banding1==1 ) , part = "body", bg = "#EFEFEF", j=left) %>%
      bg( i = which( banding2==1 ) , part = "body", bg = "#EFEFEF", j=right) %>%

      #Headers
      flextable::set_header_labels(values =
        list(
              YMD.x = "YMD",
              Event.x = "Synthetic Narrative",
              Docs.x = "Mentions",
              icb_narrative.x = "ICB Corpus",
              ICBe.x = "ICBe",
              ICB = "ICB",
              MIDs.x = "MIDs",
              phoenix.x = "Phoenix",
              phoenix.y = "Phoenix",
              row="",
              YMD.y = "YMD",
              Event.y = "Synthetic Narrative",
              Docs.y = "Mentions",
              icb_narrative.y = "ICB Corpus",
              ICBe.y = "ICBe",
              icews = "ICEWS",
              MIDs.y = "MIDs",
              terrier="Terrier"
             ) ) %>%

      add_header_row(values = c("Cuban Missile Crisis (1962)", "", "Crimea-Donbass (2014)"), colwidths = c(kleft, 1,kright) ) %>%

      #Global
      flextable::border_remove() %>% 
      #Left border
      border(j=1, i=1:n_left , part="body", border.left =  big_border ) %>% #
      border(j=1, i=1:2 , part="header", border.left =  big_border ) %>% #

      border(j=left_end, i=1:2 , part="header", border.right =  big_border ) %>% #
      border(j=left_end, i=1:n_left , part="body", border.right =  big_border ) %>% #

      border(j=1:left_end, i=n_left , part="body", border.bottom =  big_border ) %>% #
      border(j=1:left_end, i=1 , part="body", border.top =  big_border ) %>% #
      border(j=1:left_end, i=1:2 , part="header", border.top =  big_border ) %>% #

      #Right border
      border(j=right_start, i=1:n_right , part="body", border.left =  big_border ) %>% #
      border(j=k, i=1:n_right , part="body", border.right =  big_border ) %>% #

      border(j=right_start, i=1:2 , part="header", border.left =  big_border ) %>% #
      border(j=k, i=1:2 , part="header", border.right =  big_border ) %>% #
      border(j=1:k, i=1 , part="header", border.right =  big_border ) %>% #

      border(j=right_start:k, i=n_right , part="body", border.bottom =  big_border ) %>% #
      border(j=right_start:k, i=1 , part="body", border.top =  big_border ) %>% #
      border(j=right_start:k, i=1:2 , part="header", border.top =  big_border ) %>% #

      vline(i=1:n_left, j=1:(left_end-1),  border = NULL, part = "body")  %>%
      vline(i=1:n_right,j=right_start:(k-1),  border = NULL, part = "body")  %>%
      vline(i=2, j=1:(left_end-1),  border = NULL, part = "header")  %>%
      vline(i=2,j=right_start:(k-1),  border = NULL, part = "header")  %>%

      #
      hline(i=c(26,46), j=left,  border = NULL, part = "body")  %>%
      hline(i=c(11), j=right,  border = NULL, part = "body")  %>%

      #Header
      flextable::rotate(i=2, , j=3:left_end, rotation="tbrl", part="header") %>% 
      flextable::rotate(i=2, , j=(left_end+4):k, rotation="tbrl", part="header") %>%

      valign(j = 1:2, valign = "center", part = "all") %>%
      align(i = 2, j = NULL, align = "center", part = "header")  %>%

      #break
      flextable::width( j = left_end+1, width=0.25)  %>%

      #Dates
      flextable::width( j = dates, width=0.6)  %>% 
      align(i = NULL, j = dates, align = "right", part = "body")   %>%

      #events
      flextable::width( j = events, width=2.50)  %>%
      align(i = NULL, j = events, align = "left", part = "body")  %>%
      ftExtra::colformat_md(j = events, part="body") %>%

      #Checkmarks
      flextable::width( j = checkmarks, width=0.05) %>%
      align(i = NULL, j = checkmarks, align = "center", part = "body")

ft   %>% 
    #flextable::flextable_to_rmd()
    save_as_image(path="/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/recall_cuban_and_crimea.png", zoom=1, webshot = "webshot2") # webshot = "webshot2"

```

```{r, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

#Recall Summary Statistics
#, icb_narrative_cd=icb_narrative.y, icews_cd=icews_hash
temp1_long <- temp1 %>% dplyr::select(docs=Docs, ICB_Narrative=icb_narrative, ICBe=ICBe, ICB=ICB,MIDs=MIDs,Phoenix=phoenix) %>% mutate(crisis="Cuban Missile Crisis") %>% pivot_longer(cols=-c(crisis, docs)) %>% 
              mutate(value=ifelse(!is.na(value) & !value %in% 0,1,0))

temp2_long <- temp2 %>% dplyr::select(docs=Docs,ICB_Narrative=icb_narrative,ICBe=ICBe,MIDs=MIDs , Phoenix=phoenix, Terrier=terrier, ICEWS= icews) %>% mutate(crisis="Crimea-Donbas Crisis") %>% pivot_longer(cols=-c(crisis, docs)) %>% 
              mutate(value=ifelse(!is.na(value) & !value %in% 0,1,0))

temp3_long <- bind_rows(temp1_long,temp2_long)

Y_hat <- temp3_long %>% dplyr::select(name) %>% distinct()
mylogit <- glm(value ~ name, data = temp3_long, family = "binomial")
Y_hat$fit <- predict(mylogit, temp3_long %>% dplyr::select(name) %>% distinct() , type='response', se.fit=TRUE)$fit
Y_hat$se <- predict(mylogit, temp3_long %>% dplyr::select(name) %>% distinct() , type='response', se.fit=TRUE)$se.fit
Y_hat$name <- Y_hat$name %>% fct_reorder(Y_hat$fit)

#Y_hat %>% ggplot(aes(name, fit)) +
#          geom_point(size=4) +
#          geom_errorbar(aes(ymin = fit-(1.98*se), ymax = fit+(1.98*se)), width = 0.1, linetype="solid") + coord_flip()  + theme_bw() + 
#          ylab("Mean Recall") + xlab("")

#temp3_long_means <- temp3_long %>% 
#                    group_by(name) %>%
#                    summarise(value= value %>%mean() %>% round(2))

#temp3_long %>% ggplot(aes(x=log(docs), y=value, color=name)) + geom_point() +
#  geom_smooth(method = "glm", method.args = list(family = "binomial"), se=F) + 
#  ylab("Recall") + xlab("Document Mentions (Log)") +
#  facet_wrap(~crisis, scales="free_x")

library(geomtextpath)
temp3_long$crisis <- factor(temp3_long$crisis, levels= temp3_long$crisis %>% as.character() %>% unique() %>% sort() %>% rev() )
p_recall <- temp3_long %>% ggplot(aes(x=log(docs), y=value, color=name)) + 
  #geom_point() +
  geom_textsmooth(aes(label = name), size=3, method = "glm", method.args = list(family = "binomial"), se=F) + 
  ylab("Recall") + xlab("Document Mentions (Log)") +
  facet_wrap(~crisis, scales="free_x") +
  theme_bw() +
  theme(legend.position="none")

#p_recall

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_recall.png'),
       plot = p_recall ,
       width=5, height=3.0)

library(magick)
tiger <- image_read("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/recall_cuban_and_crimea.png")
tiger2 <- image_read(paste0(here::here(), '/replication_paper/figures/p_recall.png'))

combined <- image_composite(tiger, image_scale(tiger2, "x290"), offset = "+475+650")
image_write(combined, "/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/recall_cuban_and_crimea_andcounts.png")

```

\clearpage
\onecolumn
\begin{figure}[H]
\caption{Measuring Recall with Synthetic Historical Narratives \label{fig:case_study_cuban_recall}}
\centering{\includegraphics[height=18cm]{/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/recall_cuban_and_crimea_andcounts.png}}
\textit{Notes: Synthetic narratives combine several thousand accounts of each crisis into a single timeline of events, taking only those mentioned in at least 5 or more documents. Checkmarks represent whether that event could be hand matched to any detail in the ICB corpus, ICBe dataset, or any of the other event datasets.}
\end{figure}
\clearpage
\twocolumn

Precision {.unnumbered}
========================================

The other side of event measurement is precision, the degree to which a sequence of events correctly and usefully describes the information in history, $Pr(H|E)$. It does little good to recall a historical event but too vaguely (e.g. MIDs describes the Cuban Missile crisis as a blockade, a show of force, and a stalemate) or with too much error (e.g. ICEWS records 263 "Detonate Nuclear Weapons" events between 1995-2019) to be useful for downstream applications. ICBe's ontology and coding system is designed to strike a balance so that the most important information is recovered accurately but also abstracted to a level that is still useful and interpretable. You should be able to lay out events of a crisis on a timeline, as in Figure 3, and read off the macro structure of an episode from each individual move. We call this visualization a crisis map, a directed graph intersected with a timeline, and provide crisis maps for every event dataset for each case study (SI Appendix, Fig. D3 and D4) and all crises on the companion website.

```{r crisis map function, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}

global_check_overlap=F
library(igraph)
library(RColorBrewer)
library(ggraph)
target_file <- paste0(here::here(),"/replication_paper/data/in/icb_manual_recoding_master_sheet.xlsx")
dictionary_actors_labels    <- readxl::read_excel(target_file, sheet="actors") %>% 
                              dplyr::select(crisno, value_normalized=value_normalized_wikidata_id, value_normalized_label) %>% dplyr::distinct() %>% na.omit()

actor_translator <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", 
                                sheet="actor_translator") #%>% janitor::clean_names() 
#Convert to a local file for pushing
#actor_translator    <- readxl::read_excel(target_file, sheet="actor_translator") #Ya we just add colors to the actor translator and we're good
pallet <- c(brewer.pal(n=20, "Dark2"), brewer.pal(n=20, "Set1"), brewer.pal(n=20, "Set2"), brewer.pal(n=20, "Paired")) %>% unique()
library(pals); #install.packages('pals')
pallet <- c(pals::glasbey(),  pals::alphabet2())
pallet <- setdiff(pallet, c("#666666", "#999999", "#FFFF99","#FFFF33","#B2DF8A","#B3B3B3","#0000FF")) 
library(colorspace)
#cols <- c("#CDE4F3","#E7F3D3","#F7F0C7","#EFCFE5","#D0D1E7")    
pallet <- rgb(hex2RGB(pallet)@coords * 0.7)
actor_translator$actor_color = rep(pallet, length=nrow(actor_translator))

icbe_crisismaps <- function(thiscrisis=471, prune_rare=T){

  icb_events_subset <-  ICBe_events_agreed %>% 
                         dplyr::filter(crisno==thiscrisis) %>%  #417 196
                         dplyr::filter(!is.na(think_actor_a) | !is.na(say_actor_a) | !is.na(do_actor_a))   %>%
                         dplyr::mutate(decade=date_earliest_decade %>% as.numeric() ) %>%
                         dplyr::mutate(year=date_earliest_year %>% as.numeric() ) %>%
                         dplyr::mutate(month= date_earliest_month %>% as.numeric() ) %>%
                         dplyr::mutate(day= date_earliest_day %>% as.numeric()  ) %>%
                         dplyr::arrange(desc(do_timing %in% 'long before crisis'), sentence_number_int_aligned) %>%
                         tidyr::fill(decade, .direction="downup") %>%
                         tidyr::fill(year, .direction="downup") %>%
                         tidyr::fill(month, .direction="downup") %>%
                         tidyr::fill(day, .direction="downup") %>% 
                         dplyr::mutate(point_in_time_year = year %>% as.numeric())  %>% 
                         dplyr::mutate(point_in_time_month = month %>% as.numeric()) %>% 
                         arrange(decade, year,month,day) %>%
                         mutate(sent= cumsum((sentence_number_int_aligned == lag(sentence_number_int_aligned) ) %in% FALSE) ) #this is now our new order

  icb_events_subset_diad <- icb_events_subset %>% 
                                     mutate(think_actor_a = strsplit(as.character(think_actor_a), ";")) %>% unnest(think_actor_a, keep_empty=T) %>%
                                     mutate(say_actor_a = strsplit(as.character(say_actor_a), ";")) %>% unnest(say_actor_a, keep_empty=T) %>%
                                     mutate(say_actor_b = strsplit(as.character(say_actor_b), ";")) %>% unnest(say_actor_b, keep_empty=T) %>%
                                     mutate(do_actor_a = strsplit(as.character(do_actor_a), ";")) %>% unnest(do_actor_a, keep_empty=T) %>%
                                     mutate(do_actor_b = strsplit(as.character(do_actor_b), ";")) %>% unnest(do_actor_b, keep_empty=T) %>%
                                     mutate(dokind=paste(act_deescalate,act_escalate,act_uncooperative,act_cooperative, interact_deescalate,
                                                         interact_escalate, interact_decreasecoop, interact_increasecoop, sep=";")) %>%
                                     mutate(thinkkind = strsplit(as.character(thinkkind), ";")) %>% unnest(thinkkind, keep_empty=T) %>%
                                     mutate(sayintkind = strsplit(as.character(sayintkind), ";")) %>% unnest(sayintkind, keep_empty=T) %>%
                                     mutate(dokind = strsplit(as.character(dokind), ";")) %>% unnest(dokind, keep_empty=T) %>%
                                     mutate(dokind=ifelse(dokind=='',NA,dokind)) %>%
                                     filter(!is.na(thinkkind) | !is.na(sayintkind) | !is.na(dokind)) %>%
                                     dplyr::select(sentence_number_int_aligned, #You have to keep this to remove dupe monads from the same sentence
                                                    sent,
                                                    think_actor_a,say_actor_a,say_actor_b,do_actor_a,do_actor_b,
                                                    thinkkind, sayintkind, dokind,
                                                    point_in_time_year, point_in_time_month
                                                    ) %>% distinct() %>%
                                filter(!duplicated(paste(sent,think_actor_a, say_actor_a, say_actor_b, do_actor_a, do_actor_b, thinkkind, sayintkind, dokind)))

  notes=NULL
  if(prune_rare==T){
    actor_sent_counts <- 
      bind_rows(
          icb_events_subset %>% dplyr::select(sent=sentence_number_int_aligned,actor=think_actor_a),
          icb_events_subset %>% dplyr::select(sent=sentence_number_int_aligned,actor=say_actor_a),
          icb_events_subset %>% dplyr::select(sent=sentence_number_int_aligned,actor=say_actor_b),
          icb_events_subset %>% dplyr::select(sent=sentence_number_int_aligned,actor=do_actor_a),
          icb_events_subset %>% dplyr::select(sent=sentence_number_int_aligned,actor=do_actor_b)
        ) %>% 
          na.omit() %>%
          mutate(actor = strsplit(as.character(actor), ";")) %>% unnest(actor) %>% distinct() %>% count(actor) %>% arrange(n %>% desc() )

      #New pruning rule, either singletons or actors with less than 10% of the max actor
      singleton_actors <- actor_sent_counts %>% mutate(share_of_max=n/max(n)) %>% filter(n==1 | share_of_max<=0.1) %>% pull(actor) 
      left_over_count <- nrow(actor_sent_counts) - length(singleton_actors)
      if(left_over_count<3){
        #Skip pruning if it leaves less than 3 actors
      } else {

        actor_map <- dictionary_actors_labels %>% filter(crisno==thiscrisis) %>% distinct() %>% rename(actor=value_normalized) 
        actor_singleton_labels <- actor_map$value_normalized_label
        names(actor_singleton_labels) <- actor_map$actor
        actor_singleton_labels <- actor_singleton_labels[singleton_actors]

        notes <- paste("Not shown: ", paste( singleton_actors %>% sort(), collapse="; " )) %>% stringi::stri_wrap(width=80) %>% paste(collapse="\n")

        #Easiest thing is to just prune the original data
        icb_events_subset <- icb_events_subset %>%
          filter(!(
                think_actor_a %in% singleton_actors | #This doesn't work for joint ones
                say_actor_a %in% singleton_actors |
                say_actor_b %in% singleton_actors |
                do_actor_a %in% singleton_actors |
                do_actor_b %in% singleton_actors 
          ) )

          #recreate diads again using the new sentence numbers
          icb_events_subset_diad <- icb_events_subset %>% 
                                             mutate(think_actor_a = strsplit(as.character(think_actor_a), ";")) %>% unnest(think_actor_a, keep_empty=T) %>%
                                             mutate(say_actor_a = strsplit(as.character(say_actor_a), ";")) %>% unnest(say_actor_a, keep_empty=T) %>%
                                             mutate(say_actor_b = strsplit(as.character(say_actor_b), ";")) %>% unnest(say_actor_b, keep_empty=T) %>%
                                             mutate(do_actor_a = strsplit(as.character(do_actor_a), ";")) %>% unnest(do_actor_a, keep_empty=T) %>%
                                             mutate(do_actor_b = strsplit(as.character(do_actor_b), ";")) %>% unnest(do_actor_b, keep_empty=T) %>%
                                             mutate(dokind=paste(act_deescalate,act_escalate,act_uncooperative,act_cooperative, interact_deescalate,
                                                                 interact_escalate, interact_decreasecoop, interact_increasecoop, sep=";")) %>%
                                             mutate(thinkkind = strsplit(as.character(thinkkind), ";")) %>% unnest(thinkkind, keep_empty=T) %>%
                                             mutate(sayintkind = strsplit(as.character(sayintkind), ";")) %>% unnest(sayintkind, keep_empty=T) %>%
                                             mutate(dokind = strsplit(as.character(dokind), ";")) %>% unnest(dokind, keep_empty=T) %>%
                                             mutate(dokind=ifelse(dokind=='',NA,dokind)) %>%
                                             filter(!is.na(thinkkind) | !is.na(sayintkind) | !is.na(dokind)) %>%
                                             dplyr::select(sentence_number_int_aligned, #You have to keep this to remove dupe monads from the same sentence
                                                            sent,
                                                            think_actor_a,say_actor_a,say_actor_b,do_actor_a,do_actor_b,
                                                            thinkkind, sayintkind, dokind,
                                                            point_in_time_year, point_in_time_month
                                                            ) %>% distinct() %>%
                                        filter(!duplicated(paste(sent,think_actor_a, say_actor_a, say_actor_b, do_actor_a, do_actor_b, thinkkind, sayintkind, dokind))) %>%
                                        filter(!(
                                              think_actor_a %in% singleton_actors | #This doesn't work for joint ones
                                              say_actor_a %in% singleton_actors |
                                              say_actor_b %in% singleton_actors |
                                              do_actor_a %in% singleton_actors |
                                              do_actor_b %in% singleton_actors 
                                        ) ) %>% 
                                       arrange(sent) %>%
                                       mutate(sent= cumsum((sent == lag(sent) ) %in% FALSE) ) #
      }
  }

  g <- make_empty_graph(n = 0, directed = TRUE)
    #graph_from_data_frame(vertices=nodes) #relations, directed=TRUE,

  #Order actors by how many interactions they have together
  temp <- bind_rows(
    icb_events_subset_diad  %>% 
      dplyr::select(say_actor_a, say_actor_b, sent) %>% na.omit()  %>% dplyr::rename(actor_1=say_actor_a, actor_2=say_actor_b),
    icb_events_subset_diad %>% 
      filter(is.na(sayintkind)) %>%  dplyr::select(do_actor_a, do_actor_b, sent) %>% na.omit() %>% dplyr::rename(actor_1=do_actor_a, actor_2=do_actor_b)
  )  %>% arrange(actor_1, actor_2) %>% distinct()

  #new plan is to minimize the number of crossovers
  temp2 <- bind_rows(
    temp,
    temp %>% setNames(c("actor_2", "actor_1", "sent"))
  ) %>% count(actor_1, actor_2) %>% arrange(actor_1, actor_2) %>% arrange(n) %>%
  rowwise() %>%
  mutate(actor_unordered= ifelse(actor_1>actor_2, paste(actor_1, actor_2),paste(actor_2, actor_1) ) ) %>%
  arrange(desc(n))

  temp2_wide <- temp2 %>% tidyr::pivot_wider(id_cols=c(actor_1), names_from=actor_2, values_from=n) %>% mutate_if(is.numeric, tidyr::replace_na, 0) %>% as.data.frame()
  rownames(temp2_wide) <- temp2_wide$actor_1
  temp2_wide$actor_1 <- NULL
  temp2_wide <- temp2_wide %>% as.matrix()
  temp2_wide <- 1 -   (temp2_wide/max(temp2_wide))
  temp2_wide <- temp2_wide[sort(rownames(temp2_wide)), sort(colnames(temp2_wide))]
  hc <- hclust(temp2_wide %>% as.dist(), method="single")
  diag(temp2_wide) <- 0
  temp2_wide_pca <- prcomp(temp2_wide, center = F,scale. = F, retx=F, rank=1) #just run PCA on the distance matrix and order along the first eigenvalue

  #ideal_ordering <- hc$labels[hc$order] #There's a bug here, put all the isolated nodes by themselves first
  ideal_ordering <- temp2_wide_pca$rotation %>% as.data.frame() %>% arrange(PC1) %>% rownames() #This all works but it's not optimizing what we want which is to minimize the number of cross over events

  #####Nodes
  icb_events_subset_nodes <- 
            bind_rows(
                      icb_events_subset_diad %>% dplyr::select(sent, actor=think_actor_a, point_in_time_year, point_in_time_month) ,
                      icb_events_subset_diad %>% dplyr::select(sent, actor=say_actor_a, point_in_time_year, point_in_time_month) ,
                      icb_events_subset_diad %>% dplyr::select(sent, actor=say_actor_b, point_in_time_year, point_in_time_month) ,
                      icb_events_subset_diad %>% dplyr::select(sent, actor=do_actor_a, point_in_time_year, point_in_time_month) ,
                      icb_events_subset_diad %>% dplyr::select(sent, actor=do_actor_b, point_in_time_year, point_in_time_month) 
            )  %>%
            filter(!is.na(actor)) %>%
           distinct()  %>% arrange(actor,sent) %>% mutate(actor_sent= actor %>% paste0("_", str_pad(sent, 2, pad = "0") )) %>% #don't na omit this one
            mutate(x=actor %>% factor(levels=actor %>% unique() %>% setdiff(ideal_ordering) %>% c(ideal_ordering)) %>% as.numeric() *100 ,
                   y=sent*-1 *100)

  icbe_actor_to_qcode <- icb_events_long %>%
                         dplyr::filter(crisno==thiscrisis  ) %>% 
                         filter(varname_normalized %>% str_detect("actor")) %>% 
                         dplyr::select(value_normalized,value_qcode) %>% 
                         na.omit() %>% distinct()

  actor_colors_df <- icbe_actor_to_qcode %>%
    left_join(actor_translator %>% dplyr::select(value_qcode=QCode, actor_color))
  actor_colors <- actor_colors_df$actor_color
  names(actor_colors) = actor_colors_df$value_normalized

  date_sineposts <- icb_events_subset_nodes  %>%
                      mutate(x=max(x)+50) %>%
                      dplyr::select(sent, point_in_time_year, point_in_time_month,x,y) %>%
                      na.omit() %>% 
                      distinct() %>%  
                      mutate(sent=as.numeric(sent)) %>% arrange(sent) %>%
                      mutate(year_float=point_in_time_year+(point_in_time_month/12)) %>%
                      group_by(sent) %>%
                       filter(year_float==min(year_float)) %>%
                      ungroup() %>% as.data.frame() %>%
                      filter( row_number()==1 |  ( year_float > lag(year_float)  ) )  %>%
                      filter(!duplicated(year_float ) ) %>%
                      mutate(label=paste0(point_in_time_year,"-",month.abb[point_in_time_month]))

  ####Actor levels
  icb_events_subset_nodes_min <- 
    icb_events_subset_nodes %>% group_by(actor) %>% filter(sent==sent %>% min()) %>% ungroup() %>%
    left_join(dictionary_actors_labels %>% filter(crisno==thiscrisis) %>% distinct() %>% rename(actor=value_normalized) ) %>%
    #rename(qcode=actor, actor=value_normalized_label) %>% #we used to normalize to qcode and n
    rowwise() #%>%
    #mutate(actor=actor %>% stringi::stri_wrap(width=15) %>% paste0(collapse='\n')) #do at plot time so I don't break colors

  for(i in 1:nrow(icb_events_subset_nodes)){
    g <- g %>%  add_vertices(1, name=icb_events_subset_nodes$actor_sent[i],
                                actor = icb_events_subset_nodes$actor[i], 
                                sent = icb_events_subset_nodes$sent[i]  ) #, type = nodes$type[i] #, leaf = nodes$leaf[i]
  }

  xy <- cbind(icb_events_subset_nodes$x, icb_events_subset_nodes$y )

  #####Time Edges
  icb_events_subset_nodes_edges_time <- cbind(icb_events_subset_nodes, bind_rows(icb_events_subset_nodes[-1,],icb_events_subset_nodes[1,] ) ) %>% janitor::clean_names() %>% filter(actor==actor_2 & sent_2>sent) %>% distinct()

  g <- g %>% 
       add_edges(icb_events_subset_nodes_edges_time[,c('actor_sent','actor_sent_2')] %>% as.matrix() %>% t() %>% as.vector(),
                 actor=icb_events_subset_nodes_edges_time$actor, timeedge=T, leaf='', eventedge=F, bundle_count=1)

  ######Monads
  icb_events_subset_monad <- 
            bind_rows(
                      icb_events_subset_diad %>% dplyr::select(sent, actor=think_actor_a,leaf=thinkkind) ,
                      icb_events_subset_diad %>% filter(!is.na(say_actor_a) & is.na(say_actor_b)) %>% dplyr::select(sent,actor=say_actor_a,leaf=sayintkind) ,
                      icb_events_subset_diad %>% filter(!is.na(do_actor_a) & is.na(do_actor_b)) %>% dplyr::select( sent,actor=do_actor_a,leaf=dokind) 
            ) %>% 
            na.omit() %>% 
            distinct() %>% 
            filter(!duplicated(paste(sent,actor,leaf))) %>%
            mutate(actor_sent= actor %>% paste0("_", str_pad(sent, 2, pad = "0") )) %>%
            mutate(leaf = strsplit(as.character(leaf), ";")) %>% unnest(leaf) %>%
            distinct() %>% 
            group_by(sent,actor,actor_sent) %>%
              summarise(leaf=paste(leaf, collapse="\n") ) %>%
            left_join( icb_events_subset_nodes %>% dplyr::select(actor_sent,x,y))

  ######Directed
  icb_events_subset_directed <- 
            #We need to enumerate all the possibilities
            bind_rows(
              #Do alone
              icb_events_subset_diad %>% filter(is.na(thinkkind) & is.na(sayintkind) & !is.na(dokind)) %>% dplyr::select(sent,actor_1=do_actor_a, actor_2=do_actor_b, leaf=dokind) ,
              #Say alone
              icb_events_subset_diad %>% filter(is.na(thinkkind) & !is.na(sayintkind) & is.na(dokind)) %>% dplyr::select(sent,actor_1=say_actor_a, actor_2=say_actor_b, leaf=sayintkind) ,
              #Say Do
              icb_events_subset_diad %>% filter(is.na(thinkkind) & !is.na(sayintkind) & !is.na(dokind)) %>% 
                mutate(leaf=paste0(sayintkind,"-",dokind)) %>% dplyr::select(sent,actor_1=say_actor_a, actor_2=say_actor_b, leaf=leaf) 
            ) %>% 
            na.omit() %>% 
            distinct() %>% 
            #mutate(leaf = strsplit(as.character(leaf), ";")) %>% unnest(leaf) %>% distinct() %>% 
            mutate(actor_sent_1= actor_1 %>% paste0("_", str_pad(sent, 2, pad = "0") )) %>%
            mutate(actor_sent_2= actor_2 %>% paste0("_", str_pad(sent, 2, pad = "0") )) %>%
            arrange(sent, actor_sent_1,leaf) %>%
            group_by(sent,actor_1, actor_sent_1,actor_sent_2) %>%
              summarise(leaf=paste(leaf, collapse="\n") ) %>% 
           #It is actually easier just to rejex out some dupes
            mutate(leaf = leaf %>%
                     str_replace("accept\naccept","accept") %>%
                   str_replace("appeal\nappeal","appeal") %>%
                   str_replace("reject\nreject","reject") %>%
                   str_replace("express intent\nexpress intent","express intent") 
                   ) %>%
           arrange(sent, actor_sent_1,leaf) %>%
           filter(actor_sent_1!=actor_sent_2)

  #debug
  icb_events_subset_directed <- icb_events_subset_directed  %>% 
           group_by(sent, leaf) %>%
               mutate(bundle_count=n()) %>%
               mutate(leaf2=ifelse(bundle_count>1 & row_number()>1,'',leaf)) %>%  
            ungroup()

  g <- g %>% add_edges(icb_events_subset_directed[,c('actor_sent_1','actor_sent_2')] %>% as.matrix() %>% t() %>% as.vector(),
                       timeedge=F, 
                       leaf=icb_events_subset_directed$leaf2, 
                       eventedge=T,
                       actor=icb_events_subset_directed$actor_1,
                       bundle_count=icb_events_subset_directed$bundle_count
                       )

  ######Put it all together
  wrap_names <- function(x){sapply(x, FUN=function(x) x %>% stringi::stri_wrap(width=15) %>% paste0(collapse='\n') ) }
  p_metro_map <- g %>%  
          ggraph("manual",x=xy[,1],y=xy[,2]) +  #geom_edge_parallel

          #Event edges
          #geom_path(data = hbundle, aes(x, y, group = group), col = "#9d0191", size = 0.05) +
          geom_edge_arc(aes(  
                            colour=actor,
                            label=leaf,
                            filter=eventedge==T & bundle_count==1 
                            ),
                        label_colour=NA, 
                        vjust = -0.75,
                        lineheight = .75,
                        sep = unit(1, 'mm'), 
                        edge_width=0.25, 
                        show.legend = F,
                        arrow = arrow(length = unit(2, 'mm')),
                        label_size=3,
                        check_overlap=global_check_overlap,
                        start_cap = circle(3, 'mm'),
                        end_cap = circle(3, 'mm'),
                        angle_calc = 'along' #,
                        #label_push=unit(20, units='mm') #This shifts the curved labels toward the target
                       ) +
              #Grouped Edges
              geom_edge_parallel(aes(  
                                  colour=actor,
                                  label=leaf,
                                  filter=eventedge==T & bundle_count>1
                                  ),
                              vjust = -0.75,
                              lineheight = .75,
                              sep = unit(5, 'mm'), 
                              edge_width=0.4,  #0.25
                              show.legend = F,
                              arrow = arrow(length = unit(1, 'mm')),
                              label_size=4,
                              check_overlap=global_check_overlap,
                              start_cap = circle(3, 'mm'),
                              end_cap = circle(3, 'mm'),
                              angle_calc = 'along'
                             ) +

              #Time edges
              geom_edge_link(aes(
                              edge_colour=actor,
                              filter=timeedge==T
                              ),
                              edge_width=0.1, 
                              show.legend = F,
                              check_overlap=global_check_overlap
                             ) +
              geom_node_point(aes(color=actor),size=1,stroke=0.5) +
              #Fatalities
              #geom_text(data=fatalities %>% filter(interact_fatalities>1), aes(x=x+4,y=y, size=interact_fatalities+1), color='black', label='☠', angle=0, check_overlap=global_check_overlap) +

              geom_text(data=icb_events_subset_monad, aes(x=x,y=y, label=leaf), nudge_y=40, angle=0, lineheight = .75, size=3, check_overlap=global_check_overlap) +
              #Add Actor Names
              geom_text(data=icb_events_subset_nodes_min, aes(x=x,y=y, label=actor  %>% wrap_names() , color=actor),
                        nudge_y=150, angle=0, size=4, lineheight = .75, fontface = "bold",
                        check_overlap=global_check_overlap) + #-45
              geom_text(data=date_sineposts, aes(x=x,y=y, label=label), color="grey", nudge_y=0, angle=0, size=4, lineheight = .75, fontface = "bold",check_overlap=global_check_overlap) + #-45
              scale_color_manual(values = actor_colors)  +
              scale_edge_color_manual(values = actor_colors) +
              #theme_bw() + 
              #labs( title = plot_title ) +
              theme_graph() +
              theme(legend.position = "none") #+
              #ggimage::geom_image(data= units, aes(x=x+offeset,y=y-30, image = IMAGE), size = 0.015)

  if(prune_rare==T & !is.null(notes)){
     p_metro_map <- p_metro_map + labs(caption = notes)
  }

  return(p_metro_map)
}

```

```{r eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}

library(ggplot2)
library(cowplot)

p_196_icbe <- icbe_crisismaps(thiscrisis=196) #readRDS(file=paste0(here::here(), '/replication_paper/figures/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

p_471_icbe <- icbe_crisismaps(thiscrisis=471) #readRDS(file=paste0(here::here(), '/replication_paper/figures/crisismaps/p_metro_plot_471.Rds')) +
    labs(
      #title = plot_title ,
      #subtitle = subtitle ,
      caption = NULL
    ) +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.025, 0.025))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01)))

p_precision_combined <- plot_grid(
            p_196_icbe +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) ,
               p_471_icbe +
    theme(plot.margin=unit(c(0,0,0,0),"in")) +
   scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
   scale_y_continuous(expand = expansion(mult = c(0.01, 0.01))) ,  
           labels = c('Cuban Missiles','Crimea-Donbas'), align = "v", ncol = 2 , rel_widths = c(1, 1) )

save_plot(file=paste0(here::here(), '/replication_paper/figures/p_precision_combined.png'),
          p_precision_combined, base_height = 14, base_width = 18)

```

\clearpage
\onecolumn
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps  \label{fig:p_precision_combined}}
\centering{\includegraphics[width=24cm]{../figures/p_precision_combined.png}}
\end{sidewaysfigure}
\clearpage
\twocolumn

We further want to verify individual event codings, which we can do in the case of ICBe because each event is mapped to a specific span of text. We develop the iconography system for presenting event codings as coherent statements that can be compared side by side to the original source narrative as for Cuban Missiles (Figure 1), Crimea-Donbas (SI Appendix Table D1), and for every case on the companion website. We further provide a stratified sample of event codings alongside their source text (SI Appendix Table D2).

We find both the visualizations of macro structure and head-to-head comparisons of ICBe codings to the raw text to strongly support the quality of ICBe, but as with recall we seek a more objective detached universal benchmark. Our proposed measure is a reconstruction task to see whether our intended ontology can be recovered through only unsupervised clustering of sentences they were applied to. Figure 4 shows the location of every sentence from the ICBe corpus in semantic space as embeded using the same large language model as before, and the median location of each ICBe event tag applied to those sentences.^[We preprocess sentences to replace named entities with a generic Entity token.] Labels reflect the individual leaves of the ontology and colors reflect the higher level coerce branch nodes of the ontology. If ICBe has high precision, substantively similar tags ought to have been applied to substantively similar source text, which is what we see both in two dimensions in the main plot and via hierarchical clustering on all dimensions in the dendrogram along the righthand side.^[Hierchcial clustering on cosine similarity and with Ward's method.]

```{r eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}

library(flextable)
library(ftExtra)
library(tidyverse)

#Load it here before the python reticulate business
#devtools::install_github("ropenscilabs/umapr")
#library(umapr)

ICBe_events_agreed_markdown <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_markdown.Rds"))

ICBe_events_agreed_markdown_embedded <- ICBe_events_agreed_markdown %>%
                                        mutate(sentence_span_text_clean = sentence_span_text %>% str_replace_all("\\(.*?\\)|\\[.*?\\]","") ) %>%
                                        mutate(sentences_unique_sanitized = sentence_span_text_clean %>%
                                                 str_replace_all("[A-Z][a-z]*","Entity") %>%
                                                 str_replace_all("[0-9]{4}","Entity") %>% #All years with entity
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("[0-9]","") %>%  #replace all digits with blanks
                                                 str_replace_all(" {1,}"," ")   
                                               )  #

sentences_unique <- ICBe_events_agreed_markdown_embedded %>% 
                    dplyr::select(sentences_unique_sanitized) %>% 
                    distinct() %>%
                    pull(sentences_unique_sanitized) %>% 
                    unique() %>% na.omit()
length(sentences_unique ) #12162
head(sentences_unique)

```

```{r, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}

#To set python, set it in ~/.Renviron. `use_python()` doesn't seem to work as desired. See [stackoverflow](https://stackoverflow.com/questions/50145643/unable-to-change-python-path-in-reticulate/58743111#58743111).

library(reticulate)
reticulate::py_discover_config()

#use_python("/home/skynet3/anaconda3/bin/python/")
#use_python("/home/tlscherer/anaconda3/bin/python3")

```

```{python, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
#Can't get this and spacy to work in the same run. Once reticulate loads once it sticks with the environment that came first and I can't figure out how to switch it.
#pip install -U sentence-transformers

fromscratch=False

if fromscratch:
  from sentence_transformers import SentenceTransformer
  import numpy as np
  sentences=np.array(r.sentences_unique) #sentences_unique_sanitized
  sentence_chars = np.array([len(q) for q in sentences])
  #sentences=sentences[sentence_chars<1000] #cull anything with too many characters
  len(sentences) #12161 12162
  n=len(sentences)
  chunk_length=100
  n_chunks = int(np.ceil(n/chunk_length)) #can only 
  sentences_chunks=[sentences[i:i + chunk_length] for i in range(0, n, chunk_length)]
  len(sentences_chunks)
  len(sentences_chunks[0])  
  #model = SentenceTransformer('sentence-transformers/all-mpnet-base-v2')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xxl')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xl')
  model = SentenceTransformer('paraphrase-mpnet-base-v2')
  #sentences
  #sentences_test=['That is a happy person','That is a happy dog','That is a not happy person','That person is happy','This person is happy']
  #temp = model.encode(sentences_chunks[0])
  sentences_embeddings = np.vstack([model.encode(q) for q in sentences_chunks])  #[0:100] #cpu is possible just painfully slow #, device='0'
  sentences_embeddings.shape #(12103, 768)

else:
  print("Skipping")

```

```{r, results='hide', echo=F, include=T, message=F, cache=F, warning=F}

fromscratch = FALSE
if(fromscratch){
  sentences_embeddings <- py$sentences_embeddings
  dim(sentences_embeddings)
  sentences_embeddings %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
}else{
  sentences_embeddings <- readRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
  dim(sentences_embeddings)
}

```

```{r, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
#The [umapr](https://github.com/ropensci-archive/umapr) package requires python and [umap](https://github.com/lmcinnes/umap#installing). Please check their repos for installation details.  
embedding <- umapr::umap(sentences_embeddings)
#plot(embedding[,c('UMAP1','UMAP2')])

sentences_embeded_mpnet <- as.data.frame(sentences_embeddings)
sentences_embeded_umap <- embedding[,c('UMAP1','UMAP2')] %>% as.data.frame()

sentences_embeded_mpnet$sentences_unique_sanitized <- sentences_unique
sentences_embeded_umap$sentences_unique_sanitized <- sentences_unique

ICBe_events_agreed_markdown_embedded <- ICBe_events_agreed_markdown_embedded %>% ungroup() %>% 
                                        left_join(sentences_embeded_umap) %>% 
                                        left_join(sentences_embeded_mpnet)

ICBe_events_agreed_markdown_embedded %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/ICBe_events_agreed_markdown_embedded.Rds'))

ICBe_events_agreed_markdown_embedded_sample  <- ICBe_events_agreed_markdown_embedded %>% 
                                                mutate(UMAP1_round=UMAP1 %>%  round(), UMAP2_round=UMAP2 %>%  round()) %>%
                                                dplyr::group_by(UMAP1_round,UMAP2_round) %>%
                                                arrange(UMAP1,UMAP2) %>%
                                                filter(crisno==crisno[1], sentence_number_int_aligned==sentence_number_int_aligned[1] ) #only take one per crisis

```

```{r, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

library(cowplot)
ICBe_events_agreed_markdown_embedded <- readRDS(paste0(here::here(), '/replication_paper/data/temp/ICBe_events_agreed_markdown_embedded.Rds'))

value_means <- ICBe_events_agreed_markdown_embedded %>%
               dplyr::select(UMAP1,UMAP2, contains("escalate"),contains("coop"),contains("sayintkind"), contains("thinkkind")) %>%
               pivot_longer(cols=-c(UMAP1,UMAP2)) %>% filter(!is.na(value) & value!='') %>% 
               mutate(value = value %>% str_split(";")) %>% 
               unnest(value) %>% 
               mutate(value = value %>% str_trim()) %>%
               filter(!value %>% str_detect("sentence|^[0-9]")) %>%
               mutate(color=NA) %>%
               mutate(color=ifelse(name %>% str_detect("_escalate"),"red",color)) %>%
               mutate(color=ifelse(name %>% str_detect("_deescalate"),"darkgreen",color)) %>%
               mutate(color=ifelse(name %>% str_detect("uncoop|decreasecoop"),"purple",color)) %>%
               mutate(color=ifelse(name %>% str_detect("_coop|_increasecoop"),"blue",color)) %>%
               mutate(color=ifelse(name %>% str_detect("sayintkind"),"brown",color)) %>%
               mutate(color=ifelse(name %>% str_detect("thinkkind"),"darkorange",color)) %>%
               dplyr::select(-name) %>%
               group_by(color, value) %>%
               mutate(n=n()) %>%
               filter(n>20) %>% #note require at least 20 examples of that tag
               dplyr::summarise_all(median)

value_normalized_colors <- value_means$color 
names(value_normalized_colors) = value_means$value

library(ggplot2)
library(ggrepel)
p_semantic_embeddings <-
      ICBe_events_agreed_markdown_embedded %>%
      ggplot(aes(x=UMAP2,y=UMAP1)) +
      geom_point(size=0.5, alpha=0.25) + 
      #geom_text(data=value_means, aes(x=UMAP1, y=UMAP2, label=value, color=value %>% as.factor()),size=1.5) + 
      theme_bw() +
      theme(legend.position = "none") + 
      geom_label(data=value_means,
                 aes(x=UMAP2,y=UMAP1, fill = value %>% as.factor(), label=value, size=50), size = 2.5, inherit.aes=F, 
                 colour = "white", fontface = "bold", label.padding = unit(0.1, "lines")
                 ) +   
      #geom_label_repel(data=value_means,
      #           aes(x=UMAP1,y=UMAP2, fill = value %>% as.factor(), label=value, size=20), size = 3, inherit.aes=F, 
      #           colour = "white", fontface = "bold", label.padding = unit(0.1, "lines"),
      #           min.segment.length = 0
      #           ) + 
      labs(
          title = "Variation in Tags by Semantic Embeddings of Source Sentences",
          subtitle = "Sentence Embedding Paraphrase-MPNET-base-v2, UMAP Projection ",
          caption = ""
      ) +
     scale_color_manual(values = value_normalized_colors) +
     scale_fill_manual(values = value_normalized_colors) +
     scale_x_continuous(expand = c(0, 0)) +
     scale_y_continuous( expand = c(0, 0)) #+
     #coord_cartesian( xlim=c(0.75,6.5), ylim=c(7,12)) #+  #you can't do this because umap is stochastic and you'd have to pick by hand each time
     #coord_flip()

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings.png'), plot = p_semantic_embeddings, width=12, height=8)

#Now do the dendro
library(ggdendro)
#We actually need the raw embeddings

#Rex Note to take only one per sentence

#This is unreasonably slow because it's large and in tidy. Couldn't do it every time if we wanted to.
value_highdim <- ICBe_events_agreed_markdown_embedded %>%
                 dplyr::select(starts_with("V",  ignore.case = F), contains("escalate"),contains("coop"),contains("sayintkind"), 
                               contains("thinkkind"), crisno, sentence_number_int_aligned) %>% 
                 pivot_longer(cols=-c(starts_with("V",  ignore.case = F), crisno, sentence_number_int_aligned) ) %>%
                 filter(!is.na(value) & value!='') %>% #40,898
                 filter(!value %>% str_detect("sentence|^[0-9]")) %>% 
                 mutate(value = value %>% str_split(";")) %>% 
                 unnest(value) %>% 
                 mutate(value = value %>% str_trim())   %>% #
                 mutate(color=NA) %>%
                 mutate(color=ifelse(name %>% str_detect("_escalate"),"red",color)) %>%
                 mutate(color=ifelse(name %>% str_detect("_deescalate"),"darkgreen",color)) %>%
                 mutate(color=ifelse(name %>% str_detect("uncoop|decreasecoop"),"purple",color)) %>%
                 mutate(color=ifelse(name %>% str_detect("_coop|_increasecoop"),"blue",color)) %>%
                 mutate(color=ifelse(name %>% str_detect("sayintkind"),"brown",color)) %>%
                 mutate(color=ifelse(name %>% str_detect("thinkkind"),"darkorange",color)) %>%
                 dplyr::select(-name) %>%
                 group_by(color, value) %>%
                  mutate(n=n()) %>% 
                 ungroup() %>%
                 arrange(n, value) %>%
                 filter(!duplicated( paste(crisno, sentence_number_int_aligned)) ) %>% #only take one per sentence to break mechanically induced correlations
                 filter(n>20)

dim(value_highdim)

value_highdim_medians <- value_highdim %>% 
                           group_by(color, value) %>%
                           dplyr::summarise_all(median)

value_normalized_colors <- value_highdim_medians$color 
names(value_normalized_colors) = value_highdim_medians$value

m <- value_highdim_medians %>% dplyr::select(starts_with("V",  ignore.case = F)) %>% as.matrix() 
rownames(m) <- value_highdim_medians$value
d <- m %>% dist()
hc <- hclust(d, method="ward.D2")
library(ggplot2)
library(ggdendro)
hcdata <- dendro_data(hc, type = "rectangle")
p_dendro <- ggplot() +
            geom_segment(data = segment(hcdata), 
                         aes(x = x, y = y, xend = xend, yend = yend)
            ) +
            geom_text(data = label(hcdata) %>% mutate(color=value_highdim_medians$color ), 
                      aes(x = x, y = y, label = label, hjust = 0, color=color[hc$order]), 
                      size = 3
            ) +
            coord_flip() +
            scale_x_continuous(expand = c(0.02, 0.00, 0.02, 0.0) ) +
            scale_y_reverse( expand = c(0.1, 0.0, 0.0, 5.0)) +  #limits = c(-10, 10),
            theme_bw() +
            theme(legend.position="none")
#p_dendro

p_combined <- plot_grid(
  p_semantic_embeddings + xlab("") + ylab("") + 
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank()) + 
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank()),
  p_dendro  + xlab("") + ylab("") + 
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank()) + 
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank()),  
  labels = c('',''),
  align = "v", ncol = 2, rel_widths = c(3, 1)  )
#p_combined

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings_dendro.png'), plot = p_combined, width=12, height=10)

```

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ICBe event codings in comparison to Semantic Embeddings from source sentences\label{fig:semantic_embeddings}}
\centering{\includegraphics[width=17cm]{../figures/p_semantic_embeddings_dendro.png}}
\textit{Notes: Dots represent individual ICB narrative sentences, as embeded by the Paraphrase-MPNET-base-v2 large language model and flattened into two dimensions with UMAP. Text labels reflect individual leaves of the ICBe ontology, and colors represent intermediate branches of the ontology. Label placement is the median of all of the sentences that tag was applied to by the coders. The dendrogram shows hiearchical clustering of the tags. If ICBe precision is high, the sentences tags were applied to ought to say similar things, and the intended shape of the ontology ought to visually recognizable. }
\end{figure}
\clearpage
\twocolumn

Finally, how does ICBe's precision compare to the existing state of the art? The crisis-maps reveal the episode level datasets like MIDs or the original ICB are too sparse and vague to reconstruct the structure of the crisis. On the other end of the spectrum, the high recall dictionary based event datasets like Terrier and ICEWs produce so many noisy events (several hundreds thousands) that even with heavy filtering their crisis maps are completely unintelligible. Further, because of copyright issues, none of these datasets directly provide the original text spans making event level precision difficult to verify.

```{r, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggtext)

icews_clean_471 <- read_tsv("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/data/out/icews_clean_471.tsv")
dim(icews_clean_471) #270,642
icews_clean_471_small <- read_tsv("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/data/out/icews_clean_471_small.tsv")

icews_clean_subset_top10_day <- read_tsv("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/data/out/icews_clean_471_lowest.tsv")

keys <- c("sign formal agreement") #"express accord", #,"declare truce, ceasefire"
events1 <- icews_clean_subset_top10_day %>% 
           filter(cameo_label %in% keys) %>% 
           group_by(cameo_label, dyad_markdown) %>% mutate(n_all=sum(n, na.rm=T)) %>% ungroup() %>%
           mutate(dyad_markdown = tidytext::reorder_within(dyad_markdown, n_all, within = cameo_label)) 
p1 <- events1 %>%
      ggplot( aes(x = date, y = dyad_markdown)) + #
      geom_tile(fill="lightblue") + theme_bw() +
      geomtextpath::geom_textvline(xintercept = as.Date('2013-11-21'), label="Reject EU\nTrade Deal" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-03-18'), label="Annex Crimea" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-09-06'), label="MINSK I" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2015-02-21'), label="MINSK II" , linetype="dotted", color = "red", size=3) +
      ylab("") + xlab("") +
      labs(subtitle=keys %>% paste(collapse="; ")) +
      tidytext::scale_y_reordered()

keys <- c("reject plan, agreement to settle dispute") #"express accord", #,"declare truce, ceasefire"
events2 <- icews_clean_subset_top10_day %>% 
           filter(cameo_label %in% keys) %>% 
           group_by(cameo_label, dyad_markdown) %>% mutate(n_all=sum(n, na.rm=T)) %>% ungroup() %>%
           mutate(dyad_markdown = tidytext::reorder_within(dyad_markdown, n_all, within = cameo_label)) 
p2 <- events2 %>%
      ggplot( aes(x = date, y = dyad_markdown)) + #
      geom_tile(fill="lightblue") + theme_bw() +
      geomtextpath::geom_textvline(xintercept = as.Date('2013-11-21'), label="Reject EU\nTrade Deal" , linetype="dotted", color = "red", size=3) +
      ylab("") + xlab("") +
      labs(subtitle=keys %>% paste(collapse="; ")) +
      tidytext::scale_y_reordered()

#,"fight with artillery and tanks","Use unconventional violence", "fight with small arms and light weapons","Mobilize or increase armed forces","Demonstrate military or police power", "Violate ceasefire", "Occupy territory"
keys <- c("use conventional military force")
events3 <- icews_clean_subset_top10_day %>% 
           filter(cameo_label %in% keys) %>% 
           group_by(cameo_label, dyad_markdown) %>% mutate(n_all=sum(n, na.rm=T)) %>% ungroup() %>%
           mutate(dyad_markdown = tidytext::reorder_within(dyad_markdown, n_all, within = cameo_label)) 
p3 <- events3 %>%
      ggplot( aes(x = date, y = dyad_markdown)) + #
      geom_tile(fill="lightblue") + theme_bw() +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-02-27'), label="Seize Gov Building Crimea" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-02-18'), label="Rus Annex Crimea" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-03-16'), label="Fighting Donbas" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-04-06'), label="Fighting Donbas" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-09-06'), label="Continued Fighting Donbas" , linetype="dotted", color = "red", size=3) +
      ylab("") + xlab("") +
      labs(subtitle=keys %>% paste(collapse="; ")) +
      tidytext::scale_y_reordered()

#,"Refuse to ease economic sanctions, boycott, or embargo", "Reduce or stop economic assistance","Reduce or stop material aid","Impose blockade, restrict movement", "Reduce or stop military assistance"
keys <- c("impose embargo, boycott, or sanctions")
events4 <- icews_clean_subset_top10_day %>% 
           filter(cameo_label %in% keys) %>% 
           group_by(cameo_label, dyad_markdown) %>% mutate(n_all=sum(n, na.rm=T)) %>% ungroup() %>%
           mutate(dyad_markdown = tidytext::reorder_within(dyad_markdown, n_all, within = cameo_label)) 
p4 <- events4 %>%
      ggplot( aes(x = date, y = dyad_markdown)) + #
      geom_tile(fill="lightblue") + theme_bw() +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-03-17'), label="US-EU Sanction Rus" , linetype="dotted", color = "red", size=3) +
      geomtextpath::geom_textvline(xintercept = as.Date('2014-07-17'), label="EU Sanctions Rus\nRus Sanctions Ukr" , linetype="dotted", color = "red", size=3) +
      ylab("") + xlab("") +
      labs(subtitle=keys %>% paste(collapse="; ")) +
      tidytext::scale_y_reordered()

library(cowplot)
p_ICEWS <- plot_grid(
           p1 + xlim(as.Date('2013-07-01'), as.Date('2015-07-01')) + 
             theme(legend.position="none")  + theme(plot.margin = unit(c(0,0,0,0), "cm")) + theme( axis.text.y = element_markdown(size = 12, lineheight = 0.1)  ),
           p2 + xlim(as.Date('2013-07-01'), as.Date('2015-07-01')) + 
             theme(legend.position="none")  + theme(plot.margin = unit(c(0,0,0,0), "cm")) + theme( axis.text.y = element_markdown(size = 12, lineheight = 0.1)  ),
           p3 + xlim(as.Date('2013-07-01'), as.Date('2015-07-01')) + 
             theme(legend.position="none")  + theme(plot.margin = unit(c(0,0,0,0), "cm")) + theme( axis.text.y = element_markdown(size = 12, lineheight = 0.1)  ),
           p4 + xlim(as.Date('2013-07-01'), as.Date('2015-07-01')) + 
             theme(legend.position="none")  + theme(plot.margin = unit(c(0,0,0,0), "cm")) + theme( axis.text.y = element_markdown(size = 12, lineheight = 0.1)  ),
           #labels = c('Cuban Missiles','Crimea-Donbas'), 
           align = "v", ncol = 1 , rel_widths = c(1) )

save_plot(file=paste0(here::here(),
                      '/replication_paper/figures/p_precision_icews.png'),
          p_ICEWS, base_height = 10, base_width = 14)

```

However, given their high recall on our task and the global and real-time coverage of dictionary based event systems, we want to take seriously the possibility that some functional transformation could recover the precision of ICBe. For example, [@terechshenkoHotCollarLatent2020] attempts to correct for the mechanically increasing amount of news coverage each year by detrending violent event counts from Phoenix using a human coded baseline. Others have focused on verifying precision for ICEWs on specific subsets of details against known ground truths, e.g. geolocation [@cookLostAggregationImproving2019], protest events (80%) [@wuestExternalValidationProtest2020], anti-government protest networks (46.1%) [@jagerLimitsStudyingNetworks].

We take the same approach here in Figure 5, selecting four specific CAMEO event codings and checking how often they reflect a true real world event. We choose four event types around key moments in the crisis. The start of the crisis revolves around Ukraine backing out of trade deal with the EU in favor of Russia, but "sign formal agreement" events act more like a topic detector with dozens of events generated by discussions of a possible agreement but not the actual agreement which never materialized. The switch is caught by the "reject plan, agreement to settle dispute", but also continues for Victor Yanukovych for even after he was removed from power because of articles retroactively discussing the cause of his of his removal. Events for "use conventional military force" capture a threshold around the start of hostilities and who the participants were but not any particular battles or campaigns. Likewise, "impose embargo, boycott, or sanctions" captures the start of waves of sanctions and from who but are effectively constantly as the news coverage does not distinguish between subtle changes or additions. In sum, dictionary based methods on news corpora tend to have high recall because they parse everything in the news, but for the same reason their specificity for most event types is too low to back out individual chess like sequencing that ICBe aims to record.

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Counts of Events per Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{../figures/p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Event types are grouped by hand along similar themes. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

<!--
286 Sign formal agreement
Reject 99

crisis.
ICEWS Precision bad for networks  for 
See [@althausTotalErrorApproach2022] for 19 types of errors in event data.
Agreement between ICEWS, GED, and CINEP [@stundalHumanRightsViolations2021]
ICEWs and 
ICEWs protest events about 80% precision [@wangGrowingPainsGlobal2016],

Dictionary based NLP systems based on large news corpora produce many thousands events over a short time which are unintelligible from the standpoint of micro or macro precision as analyzed above.

Dictionary methods are effectively topic detectors. Dozens of events about signing a deal with the EU that didn't happen.

Next, we make available visualizations of every datum in ICBe alongside each of its contemporary competitors in full on the companion website. For ICBe, we show the original source text alongside an iconographic representation of coded events as in Figure 1. 
-->

Conclusion {#format .unnumbered}
========================================

We investigated event extraction from narratives describing key historical episodes in international relations. We synthesized a prior belief about the latent unobserved phenomena that drive these events in international relations and proposed a mapping to observable concepts that enter into the observed historical record. We designed an ontology with high coverage over those concepts and developed a training procedure and technical stack for human coding of historical texts. Multiple validity checks find the resulting codings have high internal validity (e.g. intercoder agreement) and external validity (i.e. matching source material in both micro-details at the sentence level and macro-details spanning full historical episodes). Further, these codings perform much better in terms of recall, precision, coverage, and overall coherence in capturing these historical episodes than existing event systems used in international relations.

We release several open-source products along with supporting code and documentation to further advance the study of IR, event extraction, and natural language processing. The first is the International Crisis Behavior Events (ICBe) dataset, an event-level aggregation of what took place during the crises identified by the ICB project. These data are appropriate for statistical analysis of hard questions about the sequencing of events (e.g. escalation and de-escalation of conflicts). Second, we provide a coder-level disaggregation with multiple codings of each sentence by experts and undergrads that allows for the introduction of uncertainty and human interpretation of events. Further, we release a direct mapping from the codings to the source text at the sentence level as a new resource for natural language processing. Finally, we provide a companion website that incorporates detailed visualizations of all of the data introduced here (www.crisisevents.org).

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak