### 2022-07-17 replication_paper/rmd/old/13_embedding.Rmd

---
title: 'Intercoder ICBEdataset_figure_semantic_embeddings'
site: bookdown::bookdown_site
bibliography: ../ICBintro.bib
---

# Intro

## Library Loads

```{r}

library(flextable)
library(ftExtra)
library(tidyverse)

#Load it here before the python reticulate business
#devtools::install_github("ropenscilabs/umapr")
#library(umapr)

```

```{r}

ICBe_events_agreed_markdown <- readRDS( paste0(here::here(), "/replication_data/out/ICBe_V1.1_events_agreed_markdown.Rds"))

```

```{r}

ICBe_events_agreed_markdown_embedded <- ICBe_events_agreed_markdown %>%
                                        mutate(sentence_span_text_clean = sentence_span_text %>% str_replace_all("\\(.*?\\)|\\[.*?\\]","") ) %>%
                                        mutate(sentences_unique_sanitized = sentence_span_text_clean %>%
                                                 str_replace_all("[A-Z][a-z]*","Entity") %>%
                                                 str_replace_all("[0-9]{4}","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all("Entity Entity","Entity") %>%
                                                 str_replace_all(" {1,}"," ") 
                                               )  #

sentences_unique <- ICBe_events_agreed_markdown_embedded %>% 
                    dplyr::select(sentences_unique_sanitized) %>% 
                    distinct() %>%
                    pull(sentences_unique_sanitized) %>% 
                    unique() %>% na.omit()
length(sentences_unique ) #12162
head(sentences_unique)

```

To set python, set it in ~/.Renviron. `use_python()` doesn't seem to work as desired. See [stackoverflow](https://stackoverflow.com/questions/50145643/unable-to-change-python-path-in-reticulate/58743111#58743111).

```{r, messages=F, results='hide'}

library(reticulate)
reticulate::py_discover_config()

#use_python("/home/skynet3/anaconda3/bin/python/")
#use_python("/home/tlscherer/anaconda3/bin/python3")

```

```{python}
#Can't get this and spacy to work in the same run. Once reticulate loads once it sticks with the environment that came first and I can't figure out how to switch it.
#pip install -U sentence-transformers

fromscratch=False

if fromscratch:
  from sentence_transformers import SentenceTransformer
  import numpy as np
  sentences=np.array(r.sentences_unique) #sentences_unique_sanitized
  sentence_chars = np.array([len(q) for q in sentences])
  #sentences=sentences[sentence_chars<1000] #cull anything with too many characters
  len(sentences) #12161 12162
  n=len(sentences)
  chunk_length=100
  n_chunks = int(np.ceil(n/chunk_length)) #can only 
  sentences_chunks=[sentences[i:i + chunk_length] for i in range(0, n, chunk_length)]
  len(sentences_chunks)
  len(sentences_chunks[0])  
  #model = SentenceTransformer('sentence-transformers/all-mpnet-base-v2')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xxl')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xl')
  model = SentenceTransformer('paraphrase-mpnet-base-v2')
  #sentences
  #sentences_test=['That is a happy person','That is a happy dog','That is a not happy person','That person is happy','This person is happy']
  #temp = model.encode(sentences_chunks[0])
  sentences_embeddings = np.vstack([model.encode(q) for q in sentences_chunks])  #[0:100] #cpu is possible just painfully slow #, device='0'
  sentences_embeddings.shape #12161, 768
  #At this point we're done, we don't need to cluster for this step
  #Parallel version
  #pool = model.start_multi_process_pool()
  #emb = model.encode_multi_process(sentences, pool) #Compute the embeddings using the multi-process pool
  #model.stop_multi_process_pool(pool) #Optional: Stop the proccesses in the pool
  #
  #sentences_embeddings.shape
  #from sentence_transformers import SentenceTransformer, util
  #cosine_scores = util.cos_sim(sentences_embeddings, sentences_embeddings)
  #xxl
  #  tensor([[1.0000, 0.7225, 0.8346, 0.9151, 0.9066],
  #        [0.7225, 1.0000, 0.5697, 0.6481, 0.6713],
  #        [0.8346, 0.5697, 1.0000, 0.7935, 0.7862],
  #        [0.9151, 0.6481, 0.7935, 1.0000, 0.9403],
  #        [0.9066, 0.6713, 0.7862, 0.9403, 1.0000]])
  #xl
  #tensor([[1.0000, 0.7127, 0.8432, 0.9232, 0.9157],
  #        [0.7127, 1.0000, 0.5873, 0.6428, 0.6830],
  #        [0.8432, 0.5873, 1.0000, 0.8099, 0.7972],
  #        [0.9232, 0.6428, 0.8099, 1.0000, 0.9516],
  #        [0.9157, 0.6830, 0.7972, 0.9516, 1.0000]])  
  #sentences_embeddings = model.encode(sentences[0:1000]) #
  #
  #https://towardsdatascience.com/billion-scale-semantic-similarity-search-with-faiss-sbert-c845614962e2
  # from numba import cuda 
  # device = cuda.get_current_device()
  # device.reset()
  # import faiss
  # res = faiss.StandardGpuResources()
  # index = faiss.IndexIDMap(faiss.IndexFlatIP(768)) #Ip stands for inner product which is cosine on normalized vectors
  # index.add_with_ids(sentences_embeddings, np.array(range(0, len(sentences_embeddings))))
  # gpu_index = faiss.index_cpu_to_gpu(res, 0, index)
  # #faiss.write_index(index, 'abc_news')
  # #index = faiss.read_index('abc_news')
  # #import time
  # #def search(query):
  # # t=time.time()
  # # query_vector = model.encode([query])
  # # k = 5
  # # top_k = index.search(query_vector, k)
  # # print('totaltime: {}'.format(time.time()-t))
  # # return [sentences[_id] for _id in top_k[1].tolist()[0]]
  # #
  # n=sentences_embeddings.shape[0]
  # ncentroids = 50
  # niter = 20
  # verbose = True
  # d = sentences_embeddings.shape[1]
  # kmeans_125 = faiss.Kmeans(d, 125, niter=niter, verbose=verbose)
  # kmeans_125.train(sentences_embeddings)
  # D, cluster_125 = kmeans_125.index.search(sentences_embeddings, 1)
  # kmeans_250 = faiss.Kmeans(d, 250, niter=niter, verbose=verbose)
  # kmeans_250.train(sentences_embeddings)
  # D, cluster_250 = kmeans_250.index.search(sentences_embeddings, 1)
  # kmeans_500 = faiss.Kmeans(d, 500, niter=niter, verbose=verbose)
  # kmeans_500.train(sentences_embeddings)
  # D, cluster_500 = kmeans_500.index.search(sentences_embeddings, 1)
  # kmeans_1000 = faiss.Kmeans(d, 1000, niter=niter, verbose=verbose)
  # kmeans_1000.train(sentences_embeddings)
  # D, cluster_1000 = kmeans_1000.index.search(sentences_embeddings, 1)
  # kmeans_2000 = faiss.Kmeans(d, 2000, niter=niter, verbose=verbose)
  # kmeans_2000.train(sentences_embeddings)
  # D, cluster_2000 = kmeans_2000.index.search(sentences_embeddings, 1)

else:
  print("Skipping")

```

```{r, eval=F}

df = data.frame(
  sentences=py$sentences,
  cluster_125=py$cluster_125,
  cluster_250=py$cluster_250,
  cluster_500=py$cluster_500,
  cluster_1000=py$cluster_1000,
  cluster_2000=py$cluster_2000
) %>%
  add_count(cluster_1000) %>% arrange(cluster_125, cluster_250, cluster_500, cluster_1000, cluster_2000)
df$cluster_1000 %>% table() %>% sort()

df %>% filter(cluster_1000==257)

```

```{r}
fromscratch = FALSE

if(fromscratch){
  sentences_embeddings <- py$sentences_embeddings
  dim(sentences_embeddings)
  sentences_embeddings %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
}else{
  sentences_embeddings <- readRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
  dim(sentences_embeddings)
}

```
The [umapr](https://github.com/ropensci-archive/umapr) package requires python and [umap](https://github.com/lmcinnes/umap#installing). Please check their repos for installation details.

```{r}

embedding <- umapr::umap(sentences_embeddings)
plot(embedding[,c('UMAP1','UMAP2')])

sentences_embeded_mpnet <- as.data.frame(sentences_embeddings)
sentences_embeded_umap <- embedding[,c('UMAP1','UMAP2')] %>% as.data.frame()

sentences_embeded_mpnet$sentences_unique_sanitized <- sentences_unique
sentences_embeded_umap$sentences_unique_sanitized <- sentences_unique

ICBe_events_agreed_markdown_embedded <- ICBe_events_agreed_markdown_embedded %>% ungroup() %>% 
                                        left_join(sentences_embeded_umap) %>% 
                                        left_join(sentences_embeded_mpnet)

ICBe_events_agreed_markdown_embedded %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/ICBe_events_agreed_markdown_embedded.Rds'))

ICBe_events_agreed_markdown_embedded_sample  <- ICBe_events_agreed_markdown_embedded %>% 
                                                mutate(UMAP1_round=UMAP1 %>%  round(), UMAP2_round=UMAP2 %>%  round()) %>%
                                                dplyr::group_by(UMAP1_round,UMAP2_round) %>%
                                                arrange(UMAP1,UMAP2) %>%
                                                filter(crisno==crisno[1], sentence_number_int_aligned==sentence_number_int_aligned[1] ) #only take one per crisis

```

```{r, eval=F}

#Moved a copy of this to the paper

ICBe_events_agreed_markdown_embedded <- readRDS(paste0(here::here(), '/replication_paper/data/temp/ICBe_events_agreed_markdown_embedded.Rds'))

#act_escalate
#act_deescalate
#act_uncooperative
#act_cooperative
#interact_escalate
#interact_deescalate
#interact_increasecoop
#interact_decreasecoop

value_means <- ICBe_events_agreed_markdown_embedded %>%
               dplyr::select(UMAP1,UMAP2, contains("escalate"),contains("coop"),contains("sayintkind"), contains("thinkkind")) %>%
               pivot_longer(cols=-c(UMAP1,UMAP2)) %>% filter(!is.na(value) & value!='') %>% 
               mutate(value = value %>% str_split(";")) %>% 
               unnest(value) %>% 
               mutate(value = value %>% str_trim()) %>%
               filter(!value %>% str_detect("sentence|^[0-9]")) %>%
               mutate(color=NA) %>%
               mutate(color=ifelse(name %>% str_detect("_escalate"),"red",color)) %>%
               mutate(color=ifelse(name %>% str_detect("_deescalate"),"darkgreen",color)) %>%
               mutate(color=ifelse(name %>% str_detect("uncoop|decreasecoop"),"purple",color)) %>%
               mutate(color=ifelse(name %>% str_detect("_coop|_increasecoop"),"blue",color)) %>%
               mutate(color=ifelse(name %>% str_detect("sayintkind"),"brown",color)) %>%
               mutate(color=ifelse(name %>% str_detect("thinkkind"),"darkorange",color)) %>%
               dplyr::select(-name) %>%
               group_by(color, value) %>%
               mutate(n=n()) %>%
               filter(n>20) %>%
               dplyr::summarise_all(median)

value_normalized_colors <- value_means$color 
names(value_normalized_colors) = value_means$value

library(ggplot2)
library(ggrepel)
p_semantic_embeddings <-
      ICBe_events_agreed_markdown_embedded %>%
      ggplot(aes(x=UMAP1,y=UMAP2)) +
      geom_point(size=0.5, alpha=0.25) + 
      #geom_text(data=value_means, aes(x=UMAP1, y=UMAP2, label=value, color=value %>% as.factor()),size=1.5) + 
      theme_bw() +
      theme(legend.position = "none") + 
      geom_label(data=value_means,
                 aes(x=UMAP1,y=UMAP2, fill = value %>% as.factor(), label=value, size=50), size = 2.5, inherit.aes=F, 
                 colour = "white", fontface = "bold", label.padding = unit(0.1, "lines")
                 ) +   
      #geom_label_repel(data=value_means,
      #           aes(x=UMAP1,y=UMAP2, fill = value %>% as.factor(), label=value, size=20), size = 3, inherit.aes=F, 
      #           colour = "white", fontface = "bold", label.padding = unit(0.1, "lines"),
      #           min.segment.length = 0
      #           ) + 
      labs(
          title = "Variation in Tags by Semantic Embeddings of Source Sentences",
          subtitle = "Sentence Embedding Paraphrase-MPNET-base-v2, UMAP Projection ",
          caption = ""
      ) +
     scale_color_manual(values = value_normalized_colors) +
     scale_fill_manual(values = value_normalized_colors) +
     scale_x_continuous(expand = c(0, 0)) +
     scale_y_continuous( expand = c(0, 0)) + 
       coord_flip()

ggsave(file=paste0(here::here(), '/replication_paper/figures/p_semantic_embeddings.png'), plot = p_semantic_embeddings, width=12, height=8)

```

```{r cuban case study precision, eval=T, results='hide', echo=F, include=T, message=F, cache=F}

temp <- ICBe_events_agreed_markdown_embedded_sample %>%
        #filter(crisno==196) %>%
        group_by(crisno, sentence_number_int_aligned) %>%
          filter(!duplicated(value_markdown %>% trimws() )) %>% #note there are dupe markdowns for some reason
        ungroup() %>%
        mutate(banding=sentence_number_int_aligned %% 2)
temp$is_subsstring <- NA
for(i in 1:nrow(temp)){
  temp$is_subsstring[i] <-   stringr::str_detect(temp$value_markdown %>% tolower() %>% str_replace_all('[^a-z0-9]',"") ,
                                                                                pattern = fixed(temp$value_markdown[i] %>% tolower() %>% str_replace_all('[^a-z0-9]',"") )
                                                                              ) %>% replace_na(0) %>%  sum()
}
temp <-temp %>%
        mutate(value_markdown = ifelse(is_subsstring<=1,value_markdown,'')) %>%
        group_by(sentence_number_int_aligned) %>%
          filter(value_markdown!='' | length(value_markdown)==1) %>%
        ungroup() %>%
        mutate(crisno= ifelse(!is.na(lag(sentence_span_text)) & sentence_span_text==lag(sentence_span_text), "", crisno)) %>%
        mutate(sentence_number_int_aligned= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_number_int_aligned)) %>%
        mutate(sentence_span_text= ifelse(!is.na(lag(sentence_span_text)) &  sentence_span_text==lag(sentence_span_text), "", sentence_span_text))   #C=crisno, 
      #head(100) %>%
ft_sample_precision <- temp %>%
                     dplyr::select(C=crisno,S=sentence_number_int_aligned, ICB=sentence_span_text, ICBe=value_markdown)  %>%
                      flextable::as_flextable() %>%
                      flextable::width( j = 1, width=0.25)  %>%
                      flextable::width( j = 2, width=0.25)  %>% 
                      flextable::width( j = 3, width=8.0)  %>% 
                      flextable::width( j = 4, width=8.0)  %>% 
                      flextable::fontsize(size = 9, part = "all") %>%
                      valign(j = 1:4, valign = "top", part = "body") %>%
                      flextable::line_spacing( space = 1.0, part = "all") %>% 
                      padding( padding = 1, part = "all") %>% 
                      bg( i = which( temp$banding==1 ) , part = "body", bg = "#EFEFEF", j=1:4) %>%
                      ftExtra::colformat_md(j = 4, part="body")
#ft_sample_precision

ft_sample_precision %>% save_as_image(path="/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/pnas_draft/ft_sample_precision_precision.png", zoom=1) # webshot = "webshot2"
temp <- NULL

```