### 2022-07-22 replication_corpus/rmd/download_and_clean.Rmd

---
title: "corpus download and clean"
output: html_document
date: '2022-05-16'
---

```{r setup, include=FALSE, eval=F}
knitr::opts_chunk$set(echo = TRUE)

# library(rvest)
# urls = paste0("http://www.icb.umd.edu/dataviewer/?crisno=",1:487)
# loose_list <- list()
# for(i in 1:length(urls)){
#   try({
#     df <- data.frame(text=urls[i] %>% read_html %>% html_text )  %>% 
#       mutate(url=urls[i]) %>%
#       mutate(crisno=i)
#     loose_list[[file]] <- df
#   })
# }
# cuban_missile_loose <- bind_rows(loose_list) %>% 
#   group_by(url) %>%
#   summarise(text=paste(text, collapse="\n"))  %>% screen_rex()
# dim(cuban_missile_loose) #958
# cuban_missile_loose$short <- NA

fromscratch=F
if(fromscratch){
  urls <- readLines(here::here("replication_data", "in", "sheet1.xml.rels"))[[2]] %>%
    str_replace_all("Target=","\n") %>%
    str_replace_all("TargetMode=","\n") %>%  
    str_replace_all("Type=","\n") %>%
    str_split("\n") %>% unique() %>%
    as.data.frame() %>% setNames("V1") %>% 
    filter(!V1 %>% str_detect("hyperlink|External|Relationships")) %>%
    mutate(V1 = V1 %>% str_replace_all('\"','') %>% trimws() )
  urls$V1 %>% sapply(FUN= function(x) try({ drive_download(x) })  )
  #Google drive download
  library("googledrive")
  #https://drive.google.com/open?id=186JCTWqxQCCRBbLl0JDzeFEpJxijoyoB 
  setwd(here::here("replication_data", "in", "crisis_texts"))
  getwd()
  temp <- drive_download(file="https://drive.google.com/open?id=186JCTWqxQCCRBbLl0JDzeFEpJxijoyoB" ) # type = "csv",
}
#Nope the original csv files are broken. I'm going to have to reparse them from the website
#icb_long  %>%
#  dplyr::select(sentence_hash) %>% distinct() %>% View()
#There are 4 not in the icb long #397 403 451 459 474
setdiff(1:476, icb_long %>% dplyr::filter(varname=="crisno") %>% dplyr::pull(value) %>% unique() %>% as.numeric() %>% sort())
icb_crisis_summaries_20150604 <- read_csv(here::here("replication_data", "in", "icb_crisis_summaries_20150604.csv"))

icb_crisis_summaries_20150604_handadditions <- read_csv(here::here("replication_data", "in", "icb_crisis_summaries_20150604_handadditions.csv")) #this gets us 1 to 455

icb_crisis_summaries <- bind_rows(icb_crisis_summaries_20150604,icb_crisis_summaries_20150604_handadditions) %>%
  mutate(background = background %>% stringi::stri_enc_toutf8(validate=T)) %>% #there are non utf-8 characters
  mutate(text = text %>% stringi::stri_enc_toutf8(validate=T))

#459 is in ours but not the original
#397 403 451 459 there are 4 crises is in the text that aren't in our codings
setdiff(icb_crisis_summaries$crisno %>% unique(), icb_long %>% dplyr::filter(varname=="crisno") %>% dplyr::pull(value) %>% unique() %>% as.numeric() %>% sort()) #
#And none that our in our codings that aren't in the text
setdiff(icb_long %>% dplyr::filter(varname=="crisno") %>% dplyr::pull(value) %>% unique() %>% as.numeric() %>% sort(), icb_crisis_summaries$crisno %>% unique() ) #

```

```{r}

#Note I had to manually recreate the 001 crisis because it was just a dump of a large text file containing first 200ish crises

files <- list.files(path = here::here("replication_corpus", "data", "in", "crisis_texts/"), pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE, ignore.case = FALSE, include.dirs = FALSE)

filed_df <- data.frame(file=files) %>%
            mutate(filetype=NA) %>%
            mutate(filetype=ifelse(file %>% str_detect("pdf$"), "pdf", filetype)) %>%
            mutate(filetype=ifelse(file %>% str_detect("txt$"), "txt", filetype)) %>%
            mutate(filetype=ifelse(file %>% str_detect("docx$"), "docx", filetype)) %>%
            mutate(text=NA)
n=nrow(filed_df)
path <- here::here("replication_corpus", "data", "in", "crisis_texts/")
library(pdftools)
library(textreadr)
for(i in 1:n){
  if(filed_df$filetype[i]=="txt"){
    filed_df$text[i] <- path %>% paste0(filed_df$file[i]) %>% readLines() %>% paste0(collapse="\n") %>% str_replace_all("\n",' ')
  }
  if(filed_df$filetype[i]=="docx"){
    filed_df$text[i] <- path %>% paste0(filed_df$file[i])  %>% read_docx() %>% paste0(collapse="\n") %>% str_replace_all("\n",' ')
  }
  if(filed_df$filetype[i]=="pdf"){
    filed_df$text[i] <- path %>% paste0(filed_df$file[i])  %>% pdftools::pdf_text() %>% str_replace_all("1\n|2\n|3\n|4\n",' ') %>% paste0(collapse="\n") %>% str_replace_all("\n",' ')
  }
}

#Some had both a background and pre-crisis
filed_df_clean <- filed_df %>%
                  mutate(text = text %>%
                    str_replace_all("\\(see Background and Pre-crisis, Case #7\\)","(see Case #7\\)")  %>%  #replace one parenthetical that breaks things
                    str_replace_all("see Background to","")    %>%
                    str_replace_all("Summary of World Broadcasts","SWB")  %>%
                    str_replace_all(
  "BACKGROUND|PRE-CRISIS|Background and Pre-Crisis:|Background:|Crisis:|Sources:|References:|Summary:|Summary|Background and Pre-crisis|Background |Background |Pre-crisis |Sources |Summary |Precrisis: |Crisis: |Bibliography","\n")  %>% 
                    str_replace_all(" Crisis On the morning of","\nOn the morning of")  %>% 
                    str_replace_all("Crisis A crisis for Pakistan","\nA crisis for Pakistan") 
                  ) %>% 
                  mutate(text = strsplit(as.character(text), "\n")) %>% 
                  unnest(text) %>%
                  mutate(text_nchar=text %>% nchar() )  %>%
                  group_by(file) %>%
                    mutate(i=row_number()) %>%
                    mutate(crisis_title=text[1] %>% trimws() ) %>%
                    mutate(references=text[max(i)] %>% trimws() ) %>%
                    filter(i!=1) %>% #remove the title
                    filter(i!=max(i)) %>%  #remove the references
                  ungroup()  %>%
                  add_count(file) %>%
                  mutate(text=text %>% trimws() %>%
                    str_replace_all("^ {0,}: {0,}","") %>% 
                    trimws()
                    ) %>%
                  filter(text!='') %>%
                  mutate(crisno=file %>% str_replace_all("[^0-9]","") %>% substring(1,3) %>% as.numeric())

filed_df_clean$file %>% table() %>% table()

icb_corpus <- filed_df_clean %>%
  arrange(crisno, i) %>%
  group_by(crisno) %>%
  summarise(
    crisis_title=crisis_title[1], 
    text=text %>% paste(collapse=" "),
    references=references[1],
    file=file[1]
  )

icb_corpus %>% saveRDS(here::here("replication_corpus", "data", "out", "icb_corpus_V1.0_May_16_2022.Rds"))

#icb_corpus$text %>% writeLines("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_corpus/data/in/test.txt")

```