### 2022-08-10 README.Rmd
readme update

---
output: 
  github_document:
      toc: true
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
``` 
[![CC BY-NC-SA 4.0][cc-by-nc-sa-shield]][cc-by-nc-sa]

The International Crisis Behavior Events (ICBe)
-----------------------------

This is a github repository for the International Crisis Behavior Events (ICBe) dataset. Submit any issues regarding the dataset, paper, or github repository using [the issues tab](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/issues/new/choose).

## The Paper:

[Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/replication_paper/arxiv_draft/paper.pdf) (also available on [ArXiv](https://arxiv.org/abs/2202.07081)).

The v1.1 Online Appendix can be downloaded [here](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/replication_paper/arxiv_draft/appendix.pdf).

Version 1.1 is the most recent version and was posted on [ArXiv](https://arxiv.org/abs/2202.07081) on July 26, 2022. Version 1.0 was posted on February 15, 2022.

## The Data:

The agreed datasets are the final dataset used in much of the paper and figures. It includes our best efforts at cleaning the data and reconciling intercoder agreement. The dataset is available in long and wide format. The data is also available in .tsv format in the same folders.

  - [ICBe_V1_long.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_long_agreed.Rds)
  - [ICBe_V1.1_long_agreement.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
      + All coded values individually along with information about how often they were selected by coders.  
      + `ICBe_V1_long.Rds` filtered down only to those codings that were agreed upon (see Algorithm 1 in paper).
  - [ICBe_V1.1_events_agreed_long.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds)
    - [ICBe_V1.1_events_agreed.Rds](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed_long.Rds)

      + `ICBe_V1.1_events_agreed.Rds` in wide form where each row is an event.

The coding and cleaning process are described in the paper with additional information and details about the variables in the codebook.
  - [ICBEdataset Codebook](https://docs.google.com/document/d/1aJkweohbfIWtNpJw1CmXbeIiK6czbJ5iPyKwiYP1YlU/edit?usp=sharing)

## The Authors:

[Rex W. Douglass](http://www.rexdouglass.com), [Thomas Leo Scherer](http://tlscherer.com/), [J. Andrés Gannon](https://jandresgannon.com/), [Erik Gartzke](http://erikgartzke.com/), [Jon Lindsay](https://www.jonrlindsay.com/), [Shannon Carcelli](https://www.shannoncarcelli.com/), Jonathan Wilkenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge.

## Citation:
For any use of the dataset or paper, please cite:

Douglass, Rex W., Thomas Leo Scherer, J. Andrés Gannon, Erik Gartzke, Jon Lindsay, Shannon Carcelli, Jonathan Wiklenfeld, David M. Quinn, Catherine Aiken, Jose Miguel Cabezas Navarro, Neil Lund, Egle Murauskaite, and Diana Partridge. 2022. "Introducing the ICBe Dataset: Very High Recall and Precision Event Extraction from Narratives about International Crises." arXiv:2202.07081 [cs, stat]. http://arxiv.org/abs/2202.07081.

Replication Code and Analysis
-----------------------------

A description of the file and folders in the repository used to create the datasets, tables, figures.

## data preparation

### replication_corpus

 - download_and_clean
    + creates a succinct rds of the crisis narratives: `.replication_corpus/data/out/icb_corpus_V1.0_May_16_2022`

### replication_data

 - 01_compile_saves_and_align
    + compiles the original coding files into `./replication_data/in/icb_long_spans.Rds`. The original coding files are not on the public repository. Public users will load `icb_long_spans.Rds` directly.
    + aligns codings from multiple GUI versions on similar source sentences
 - 02_format_and_clean
    + applies cleaning dictionaries to create `./replication_data/out/ICBe_V1.1_long_clean.Rds`.
 - 03_aggregation
    + applies aggregation algorithm to create `./replication_data/out/ICBe_V1.1_long_agreement.Rds`, `ICBe_V1.1_long.Rds`,  `./replication_data/out/ICBe_V1.1_events_agreed.Rds`, `./replication_data/out/ICBe_V1.1_events_agreed_long.Rds`.
 - 04_validation
    + applies iconography to crises and events to create `ICBe_V1.1_crises_markdown.Rds` and `ICBe_V1.1_events_agreed_markdown.Rds`

### replication_paper

The figures are created in Rmd file for the paper (./replication_paper/pnas_draft/ICBe_pnas_submission_rmd.Rmd). In some cases they have been transformed to other formats via GNU Image Manipulation Program.

  - `case_study_cuban_precision.png`
  - `recall_cuban_and_crimea_andcounts.png`
    + Draws from the [Cuban Missile Automated Case Study googlesheet](https://docs.google.com/spreadsheets/d/1NuRWFB1HEbQbJu_JoEzJ9WYk7kfwwCryF7wBfKPT6uc/edit?usp=sharing) and the [Crimea-Donbas Automated Case Study (redone) googlesheet](https://docs.google.com/spreadsheets/d/1YesAx1CkYCgrEi_WVJ9aho60HesWA6p-3XMma9IzBrI/edit?usp=sharing).
    + uses the iconographry in `./replication_data/in/flags_small/`
  - `p_precision_combined.png`
    + combines metro maps of the two case studies using ICBe_V1.1_events_agreed.Rds
  - `p_semantic_embeddings_dendro.png`
    + plot of semantic embeddings of ICBe_V1.1_events_agreed_markdown.Rds 
  - `p_precision_icews.png`
    + mapping of icews using `./replication_paper/data/out/icews_clean_471_lowest.tsv` (created in `./replication_paper/pnas_draft/appendix.Rmd`)

## Data inputs:

### ICBe preparation
  - Cleaning dictionaries: used to clean raw codings for actors, actions, locations, and dates `/replication_paper/data/in/icb_manual_recording_master_sheet.xlsx`
  - Lit review and tree/leaf codebook: `replication_data/in/icbe_litreview_trees_sentences.xlsx` and `/replication_paper/data/in/icbe_litreview_trees_sentences.xlsx`
  - Case study tables: `/replication_paper/data/in/CaseStudies.xlsx`

### External datasets
  - [The ICB project](https://sites.duke.edu/icbdata/) 
    + System-level (icb1v14.csv) and Actor-level (icb2v14.csv) datasets
    + Dyadic-Level Crisis Data ([source](https://sites.duke.edu/icbdata/data-collections/))
  - [Militarized Interstate Disputes (MID)](https://correlatesofwar.org/data-sets/MIDs) version 5.01 at the incident level (MIDI_5.01.Rds) and incident-participant level (MIDIP_5.01.Rds) converted to Rds. 
  - [UCDP Georeferenced Event Dataset (GED) Global version 21.1](https://ucdp.uu.se/downloads/index.html#ged_global) (GEDEvent_v23_1.RData)
  - Cameo Event Codes adapted from [CAMEO Conflict and Mediation Event Observations Codebook](https://parusanalytics.com/eventdata/cameo.dir/CAMEO.09b6.pdf) (cameo.eventcode.txt)
  - [Phoenix Event data](https://databank.illinois.edu/datasets/IDB-2796521)
  - [Terrier event data](https://osf.io/4m2u7/files/) 
    + too large to include in the github repository
    + to replicate, download the folder 'largegeolocatedata' to ICBEdata/replication_paper/data/ignore and decompress
  - [ICEWS data](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/28075&version=30.0)
    + too large to include in the github repository
    + to replicate, download folder 'dataverse_files' to ICBEdata/replication_paper/data/ignore/ and decompress

### Appendix

The v1.1 Online Appendix can be downloaded [here](https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/raw/master/replication_paper/arxiv_draft/appendix.pdf).

### License

This work is licensed under a
[Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License][cc-by-nc-sa].

[![CC BY-NC-SA 4.0][cc-by-nc-sa-image]][cc-by-nc-sa]

[cc-by-nc-sa]: http://creativecommons.org/licenses/by-nc-sa/4.0/
[cc-by-nc-sa-image]: https://licensebuttons.net/l/by-nc-sa/4.0/88x31.png
[cc-by-nc-sa-shield]: https://img.shields.io/badge/License-CC%20BY--NC--SA%204.0-lightgrey.svg