### 2022-09-12 replication_paper/political_analysis_draft/appendix.tex
supplemental authors

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Introducing ICBe: Supplementary Information}

\author{
    Rex W. Douglass
    \thanks{Correspondence should be addressed to Rex W. Douglass at
\href{mailto:rexdouglass@gmail.com}{\nolinkurl{rexdouglass@gmail.com}}.}
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Thomas Leo Scherer
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    J. Andrés Gannon
   \\
    Vanderbilt University \\
   \\
  \texttt{} \\
   \And
    Erik Gartzke
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Jon Lindsay
   \\
    Georgia Institute of Technology \\
   \\
  \texttt{} \\
   \And
    Shannon Carcelli
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Jonathan Wilkenfeld
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    David M. Quinn
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Catherine Aiken
   \\
    Georgetown University \\
   \\
  \texttt{} \\
   \And
    Jose Miguel Cabezas Navarro
   \\
    Universidad Mayor \\
   \\
  \texttt{} \\
   \And
    Neil Lund
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Egle Murauskaite
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Diana Partridge
   \\
    University of Maryland \\
   \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

\newpage 
\tableofcontents 
\newpage

\clearpage

\hypertarget{appendix-a-corpus-and-cases}{%
\section{Appendix A Corpus and
Cases}\label{appendix-a-corpus-and-cases}}

\hypertarget{app:crises}{%
\subsection{Table A1: ICBe Crises and Summary
Information}\label{app:crises}}

\begin{figure}[H]
\caption{ICBe Crises Summary Statistics}
\centering{\includegraphics[ height=19cm]{"./crisis_text_counts_ft1.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft2.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft3.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft4.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft5.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft6.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft7.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft8.png"}}
\end{figure}
\clearpage

\hypertarget{app:verbs}{%
\subsection{Table A2: Verb Meanings (Glosses) found in the ICB Corpus
(top 200)}\label{app:verbs}}

\hphantom{em}

\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{0.75}

\begin{longtable}[c]{|p{0.30in}|p{0.75in}|p{6.00in}}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{uses}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Lemma}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Gloss}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}

\endfirsthead

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{uses}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Lemma}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{Gloss}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}\endhead

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{381}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{end}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ to\ an\ end\ or\ halt\ "She\ ended\ their\ friendship\ when\ she\ found\ out\ that\ he\ had\ once\ been\ convicted\ of\ a\ crime"\ "The\ attack\ on\ Poland\ terminated\ the\ relatively\ peaceful\ period\ after\ WW\ I"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{365}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{trigger}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{release\ or\ pull\ the\ trigger\ on\ "Trigger\ a\ gun"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{323}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{involve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{engage\ as\ a\ participant\ "Dont\ involve\ me\ in\ your\ family\ affairs"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{309}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{begin}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have\ a\ beginning,\ of\ a\ temporal\ event\ "WW\ II\ began\ in\ 1939\ when\ Hitler\ marched\ into\ Poland"\ "The\ companys\ Asia\ tour\ begins\ next\ month"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{278}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{trigger}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{put\ in\ motion\ or\ move\ to\ act\ "trigger\ a\ reaction"\ "actuate\ the\ circuits"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{260}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{announce}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{announce\ publicly\ or\ officially\ "The\ President\ declared\ war"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{236}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{respond}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{respond\ favorably\ or\ as\ hoped\ "The\ cancer\ responded\ to\ the\ aggressive\ therapy"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{212}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{call}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{call\ a\ meeting\ invite\ or\ command\ to\ meet\ "The\ Wannsee\ Conference\ was\ called\ to\ discuss\ the\ Final\ Solution"\ "The\ new\ dean\ calls\ meetings\ every\ week"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{194}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{occur}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{come\ to\ pass\ "What\ is\ happening"\ "The\ meeting\ took\ place\ off\ without\ an\ incidence"\ "Nothing\ occurred\ that\ seemed\ important"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{194}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{sign}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{approve\ and\ express\ assent,\ responsibility,\ or\ obligation\ "All\ parties\ ratified\ the\ peace\ treaty"\ "Have\ you\ signed\ your\ contract\ yet"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{185}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{terminate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ to\ an\ end\ or\ halt\ "She\ ended\ their\ friendship\ when\ she\ found\ out\ that\ he\ had\ once\ been\ convicted\ of\ a\ crime"\ "The\ attack\ on\ Poland\ terminated\ the\ relatively\ peaceful\ period\ after\ WW\ I"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{178}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{begin}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{set\ in\ motion,\ cause\ to\ start\ "The\ U.S.\ started\ a\ war\ in\ the\ Middle\ East"\ "The\ Iraqis\ began\ hostilities"\ "begin\ a\ new\ chapter\ in\ your\ life"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{160}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{meet}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{come\ together\ "Ill\ probably\ see\ you\ at\ the\ meeting"\ "How\ nice\ to\ see\ you\ again"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{146}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{send}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ cause\ or\ order\ to\ be\ taken,\ directed,\ or\ transmitted\ to\ another\ place\ "He\ had\ sent\ the\ dispatches\ downtown\ to\ the\ proper\ people\ and\ had\ slept"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{145}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{occupy}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{march\ aggressively\ into\ anothers\ territory\ by\ military\ force\ for\ the\ purposes\ of\ conquest\ and\ occupation\ "Hitler\ invaded\ Poland\ on\ September\ 1,\ 1939"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{135}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{happen,\ occur,\ take\ place\ "I\ lost\ my\ wallet\ this\ was\ during\ the\ visit\ to\ my\ parents\ house"\ "There\ were\ two\ hundred\ people\ at\ his\ funeral"\ "There\ was\ a\ lot\ of\ noise\ in\ the\ kitchen"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{128}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{perceive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{become\ conscious\ of\ "She\ finally\ perceived\ the\ futility\ of\ her\ protest"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{126}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{become}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{undergo\ a\ change\ or\ development\ "The\ water\ turned\ into\ ice"\ "Her\ former\ friend\ became\ her\ worst\ enemy"\ "He\ turned\ traitor"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{125}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{condemn}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{express\ strong\ disapproval\ of\ "We\ condemn\ the\ racism\ in\ South\ Africa"\ "These\ ideas\ were\ reprobated"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{125}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ by\ force\ "Hitler\ took\ the\ Baltic\ Republics"\ "The\ army\ took\ the\ fort\ on\ the\ hill"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{124}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{escalate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{increase\ in\ extent\ or\ intensity\ "The\ Allies\ escalated\ the\ bombing"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{119}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reject}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{refuse\ to\ accept\ or\ acknowledge\ "I\ reject\ the\ idea\ of\ starting\ a\ war"\ "The\ journal\ rejected\ the\ students\ paper"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{117}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{respond}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{react\ verbally\ "She\ didnt\ want\ to\ answer"\ "answer\ the\ question"\ "We\ answered\ that\ we\ would\ accept\ the\ invitation"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{116}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{demand}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{request\ urgently\ and\ forcefully\ "The\ victims\ family\ is\ demanding\ compensation"\ "The\ boss\ demanded\ that\ he\ be\ fired\ immediately"\ "She\ demanded\ to\ see\ the\ manager"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{112}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ remain\ unmolested,\ undisturbed,\ or\ uninterrupted\ \ used\ only\ in\ infinitive\ form\ "let\ her\ be"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{111}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{agree}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ in\ accord\ be\ in\ agreement\ "We\ agreed\ on\ the\ terms\ of\ the\ settlement"\ "I\ cant\ agree\ with\ you"\ "I\ hold\ with\ those\ who\ say\ life\ is\ sacred"\ "Both\ philosophers\ concord\ on\ this\ point"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{110}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reach}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ in\ or\ establish\ communication\ with\ "Our\ advertisements\ reach\ millions"\ "He\ never\ contacted\ his\ children\ after\ he\ emigrated\ to\ Australia"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{106}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{withdraw}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{pull\ back\ or\ move\ away\ or\ backward\ "The\ enemy\ withdrew"\ "The\ limo\ pulled\ away\ from\ the\ curb"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{105}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accuse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{blame\ for,\ make\ a\ claim\ of\ wrongdoing\ or\ misbehavior\ against\ "he\ charged\ the\ director\ with\ indifference"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{105}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{dispute}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have\ a\ disagreement\ over\ something\ "We\ quarreled\ over\ the\ question\ as\ to\ who\ discovered\ America"\ "These\ two\ fellows\ are\ always\ scrapping\ over\ something"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{105}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fight}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ engaged\ in\ a\ fight\ carry\ on\ a\ fight\ "the\ tribesmen\ fought\ each\ other"\ "Siblings\ are\ always\ fighting"\ "Militant\ groups\ are\ contending\ for\ control\ of\ the\ country"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{102}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue\ talking\ "I\ know\ its\ hard,\ he\ continued,\ but\ there\ is\ no\ choice"\ "carry\ onpretend\ we\ are\ not\ in\ the\ room"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{100}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{include}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{consider\ as\ part\ of\ something\ "I\ include\ you\ in\ the\ list\ of\ culprits"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{99}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{remain}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ left\ of\ persons,\ questions,\ problems,\ results,\ evidence,\ etc.\ "There\ remains\ the\ question\ of\ who\ pulled\ the\ trigger"\ "Carter\ remains\ the\ only\ President\ in\ recent\ history\ under\ whose\ Presidency\ the\ U.S.\ did\ not\ fight\ a\ war"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{98}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{experience}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{undergo\ or\ live\ through\ a\ difficult\ experience\ "We\ had\ many\ trials\ to\ go\ through"\ "he\ saw\ action\ in\ Viet\ Nam"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{97}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{attempt}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ an\ effort\ or\ attempt\ "He\ tried\ to\ shake\ off\ his\ fears"\ "The\ infant\ had\ essayed\ a\ few\ wobbly\ steps"\ "The\ police\ attempted\ to\ stop\ the\ thief"\ "He\ sought\ to\ improve\ himself"\ "She\ always\ seeks\ to\ do\ good\ in\ the\ world"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{94}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{increase}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ bigger\ or\ more\ "The\ boss\ finally\ increased\ her\ salary"\ "The\ university\ increased\ the\ number\ of\ students\ it\ admitted"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{94}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{mediate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{act\ between\ parties\ with\ a\ view\ to\ reconciling\ differences\ "He\ interceded\ in\ the\ family\ dispute"\ "He\ mediated\ a\ settlement"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{94}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{threaten}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ utter\ intentions\ of\ injury\ or\ punishment\ against\ "He\ threatened\ me\ when\ I\ tried\ to\ call\ the\ police"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{93}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{request}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{inquire\ for\ information\ "I\ requested\ information\ from\ the\ secretary"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{92}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{withdraw}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ a\ retreat\ from\ an\ earlier\ commitment\ or\ activity\ "Well\ have\ to\ crawfish\ out\ from\ meeting\ with\ him"\ "He\ backed\ out\ of\ his\ earlier\ promise"\ "The\ aggressive\ investment\ company\ pulled\ in\ its\ horns"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{91}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{hold}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stop\ dealing\ with\ "hold\ all\ calls\ to\ the\ Presidents\ office\ while\ he\ is\ in\ a\ meeting"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{90}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{invade}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{march\ aggressively\ into\ anothers\ territory\ by\ military\ force\ for\ the\ purposes\ of\ conquest\ and\ occupation\ "Hitler\ invaded\ Poland\ on\ September\ 1,\ 1939"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{87}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{send}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause\ to\ go\ somewhere\ "The\ explosion\ sent\ the\ car\ flying\ in\ the\ air"\ "She\ sent\ her\ children\ to\ camp"\ "He\ directed\ all\ his\ energies\ into\ his\ dissertation"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{86}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{launch}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{begin\ with\ vigor\ "He\ launched\ into\ a\ long\ diatribe"\ "She\ plunged\ into\ a\ dangerous\ adventure"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{84}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{report}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{announce\ as\ the\ result\ of\ an\ investigation\ or\ experience\ or\ finding\ "Dozens\ of\ incidents\ of\ wife\ beatings\ are\ reported\ daily\ in\ this\ city"\ "The\ team\ reported\ significant\ advances\ in\ their\ research"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{84}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{see}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{perceive\ or\ be\ contemporaneous\ with\ "We\ found\ Republicans\ winning\ the\ offices"\ "Youll\ see\ a\ lot\ of\ cheating\ in\ this\ school"\ "The\ 1960s\ saw\ the\ rebellion\ of\ the\ younger\ generation\ against\ established\ traditions"\ "I\ want\ to\ see\ results"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{81}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{perceive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ become\ aware\ of\ through\ the\ senses\ "I\ could\ perceive\ the\ ship\ coming\ over\ the\ horizon"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{80}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accuse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ an\ accusation\ against\ level\ a\ charge\ against\ "The\ neighbors\ accused\ the\ man\ of\ spousal\ abuse"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{77}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{attack}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{launch\ an\ attack\ or\ assault\ on\ begin\ hostilities\ or\ start\ warfare\ with\ "Hitler\ attacked\ Poland\ on\ September\ 1,\ 1939\ and\ started\ World\ War\ II"\ "Serbian\ forces\ assailed\ Bosnian\ towns\ all\ week"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{77}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{dispatch}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{kill\ without\ delay\ "the\ traitor\ was\ dispatched\ by\ the\ conspirators"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{77}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{end}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{put\ an\ end\ to\ "The\ terrible\ news\ ended\ our\ hopes\ that\ he\ had\ survived"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{74}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{declare}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{announce\ publicly\ or\ officially\ "The\ President\ declared\ war"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{74}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{discuss}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{speak\ with\ others\ about\ something\ talk\ something\ over\ in\ detail\ have\ a\ discussion\ "We\ discussed\ our\ household\ budget"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{74}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{follow}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{keep\ under\ surveillance\ "The\ police\ had\ been\ following\ him\ for\ weeks\ but\ they\ could\ not\ prove\ his\ involvement\ in\ the\ bombing"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{73}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{lead}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{tend\ to\ or\ result\ in\ "This\ remark\ lead\ to\ further\ arguments\ among\ the\ guests"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{73}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{remain}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stay\ behind\ "The\ smell\ stayed\ in\ the\ room"\ "The\ hostility\ remained\ long\ after\ they\ made\ up"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{72}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{propose}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{present\ for\ consideration,\ examination,\ criticism,\ etc.\ "He\ proposed\ a\ new\ plan\ for\ dealing\ with\ terrorism"\ "She\ proposed\ a\ new\ theory\ of\ relativity"\ "The\ candidate\ projects\ himself\ as\ a\ moderate\ and\ a\ reformer"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{72}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{recognize}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accept\ someone\ to\ be\ what\ is\ claimed\ or\ accept\ his\ power\ and\ authority\ "The\ Crown\ Prince\ was\ acknowledged\ as\ the\ true\ heir\ to\ the\ throne"\ "We\ do\ not\ recognize\ your\ gods"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{70}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{capture}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ possession\ of\ by\ force,\ as\ after\ an\ invasion\ "the\ invaders\ seized\ the\ land\ and\ property\ of\ the\ inhabitants"\ "The\ army\ seized\ the\ town"\ "The\ militia\ captured\ the\ castle"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{70}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{deny}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{deny\ formally\ an\ allegation\ of\ fact\ by\ the\ opposing\ party\ in\ a\ legal\ suit}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{69}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{exist\ over\ a\ prolonged\ period\ of\ time\ "The\ bad\ weather\ continued\ for\ two\ more\ weeks"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{69}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{issue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ out\ an\ official\ document\ such\ as\ a\ warrant}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{67}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{declare}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{proclaim\ ones\ support,\ sympathy,\ or\ opinion\ for\ or\ against\ "His\ wife\ declared\ at\ once\ for\ moving\ to\ the\ West\ Coast"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{66}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{lead}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stretch\ out\ over\ a\ distance,\ space,\ time,\ or\ scope\ run\ or\ extend\ between\ two\ points\ or\ beyond\ a\ certain\ point\ "Service\ runs\ all\ the\ way\ to\ Cranbury"\ "His\ knowledge\ doesnt\ go\ very\ far"\ "My\ memory\ extends\ back\ to\ my\ fourth\ year\ of\ life"\ "The\ facts\ extend\ beyond\ a\ consideration\ of\ her\ personal\ assets"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{65}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{renew}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reestablish\ on\ a\ new,\ usually\ improved,\ basis\ or\ make\ new\ or\ like\ new\ "We\ renewed\ our\ friendship\ after\ a\ hiatus\ of\ twenty\ years"\ "They\ renewed\ their\ membership"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{64}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{achieve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ gain\ with\ effort\ "she\ achieved\ her\ goal\ despite\ setbacks"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{63}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{provide}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{determine\ what\ is\ to\ happen\ in\ certain\ contingencies,\ especially\ by\ including\ a\ proviso\ condition\ or\ stipulation\ "The\ will\ provides\ that\ each\ child\ should\ receive\ half\ of\ the\ money"\ "The\ Constitution\ provides\ for\ the\ right\ to\ free\ speech"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{61}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{appeal}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ a\ court\ case\ to\ a\ higher\ court\ for\ review\ "He\ was\ found\ guilty\ but\ appealed\ immediately"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{61}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{seek}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ an\ effort\ or\ attempt\ "He\ tried\ to\ shake\ off\ his\ fears"\ "The\ infant\ had\ essayed\ a\ few\ wobbly\ steps"\ "The\ police\ attempted\ to\ stop\ the\ thief"\ "He\ sought\ to\ improve\ himself"\ "She\ always\ seeks\ to\ do\ good\ in\ the\ world"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{60}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{attack}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ the\ initiative\ and\ go\ on\ the\ offensive\ "The\ Serbs\ attacked\ the\ village\ at\ night"\ "The\ visiting\ team\ started\ to\ attack"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{60}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{pass}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{go\ unchallenged\ be\ approved\ "The\ bill\ cleared\ the\ House"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{59}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{agree}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{achieve\ harmony\ of\ opinion,\ feeling,\ or\ purpose\ "No\ two\ of\ my\ colleagues\ would\ agree\ on\ whom\ to\ elect\ chairman"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{59}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{last}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{persist\ for\ a\ specified\ period\ of\ time\ "The\ bad\ weather\ lasted\ for\ three\ days"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{58}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{claim}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{demand\ as\ being\ ones\ due\ or\ property\ assert\ ones\ right\ or\ title\ to\ "He\ claimed\ his\ suitcases\ at\ the\ airline\ counter"\ "Mr.\ Smith\ claims\ special\ tax\ exemptions\ because\ he\ is\ a\ foreign\ resident"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{57}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{establish}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ about\ "The\ trompe\ loeilillusion\ establishes\ depth"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{57}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{inform}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{impart\ knowledge\ of\ some\ fact,\ state\ of\ affairs,\ or\ event\ to\ "I\ informed\ him\ of\ his\ rights"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{57}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{warn}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{notify,\ usually\ in\ advance\ "I\ warned\ you\ that\ I\ would\ ask\ some\ difficult\ questions"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{56}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{offer}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{threaten\ to\ do\ something\ "I\ offered\ to\ leave\ the\ committee\ if\ they\ did\ not\ accept\ my\ proposal"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{56}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{restore}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ back\ into\ original\ existence,\ use,\ function,\ or\ position\ "restore\ law\ and\ order"\ "reestablish\ peace\ in\ the\ region"\ "restore\ the\ emperor\ to\ the\ throne"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{55}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{create}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ or\ cause\ to\ be\ or\ to\ become\ "make\ a\ mess\ in\ ones\ office"\ "create\ a\ furor"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{55}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{issue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{circulate\ or\ distribute\ or\ equip\ with\ "issue\ a\ new\ uniform\ to\ the\ children"\ "supply\ blankets\ for\ the\ beds"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{55}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{prevent}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{keep\ from\ happening\ or\ arising\ make\ impossible\ "My\ sense\ of\ tact\ forbids\ an\ honest\ answer"\ "Your\ role\ in\ the\ projects\ precludes\ your\ involvement\ in\ the\ competitive\ project"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{55}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{support}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{argue\ or\ speak\ in\ defense\ of\ "She\ supported\ the\ motion\ to\ strike"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{54}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cross}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{travel\ across\ or\ pass\ over\ "The\ caravan\ covered\ almost\ 100\ miles\ each\ day"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{54}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{include}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{allow\ participation\ in\ or\ the\ right\ to\ be\ part\ of\ permit\ to\ exercise\ the\ rights,\ functions,\ and\ responsibilities\ of\ "admit\ someone\ to\ the\ profession"\ "She\ was\ admitted\ to\ the\ New\ Jersey\ Bar"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{54}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{urge}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{push\ for\ something\ "The\ travel\ agent\ recommended\ strongly\ that\ we\ not\ travel\ on\ Thanksgiving\ Day"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{53}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{agree}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{consent\ or\ assent\ to\ a\ condition,\ or\ agree\ to\ do\ something\ "She\ agreed\ to\ all\ my\ conditions"\ "He\ agreed\ to\ leave\ her\ alone"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{53}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{do\ something\ repeatedly\ and\ showing\ no\ intention\ to\ stop\ "We\ continued\ our\ research\ into\ the\ cause\ of\ the\ illness"\ "The\ landlord\ persists\ in\ asking\ us\ to\ move"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{53}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{send}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{transfer\ "The\ spy\ sent\ the\ classified\ information\ off\ to\ Russia"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{53}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{support}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ behind\ approve\ of\ "He\ plumped\ for\ the\ Labor\ Party"\ "I\ backed\ Kennedy\ in\ 1960"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{52}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{erupt}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{erupt\ or\ intensify\ suddenly\ "Unrest\ erupted\ in\ the\ country"\ "Tempers\ flared\ at\ the\ meeting"\ "The\ crowd\ irrupted\ into\ a\ burst\ of\ patriotism"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{52}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{resolve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reach\ a\ conclusion\ after\ a\ discussion\ or\ deliberation}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{52}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{view}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{deem\ to\ be\ "She\ views\ this\ quite\ differently\ from\ me"\ "I\ consider\ her\ to\ be\ shallow"\ "I\ dont\ see\ the\ situation\ quite\ as\ negatively\ as\ you\ do"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{50}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{occur}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ be\ found\ to\ exist\ "sexism\ occurs\ in\ many\ workplaces"\ "precious\ stones\ occur\ in\ a\ large\ area\ in\ Brazil"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{50}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{admit\ into\ a\ group\ or\ community\ "accept\ students\ for\ graduate\ study"\ "Well\ have\ to\ vote\ on\ whether\ or\ not\ to\ admit\ a\ new\ member"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{claim}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ as\ an\ undesirable\ consequence\ of\ some\ event\ or\ state\ of\ affairs\ "the\ accident\ claimed\ three\ lives"\ "The\ hard\ work\ took\ its\ toll\ on\ her"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fade}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{disappear\ gradually\ "The\ pain\ eventually\ passed\ off"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{intervene}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{get\ involved,\ so\ as\ to\ alter\ or\ hinder\ an\ action,\ or\ through\ force\ or\ threat\ of\ force\ "Why\ did\ the\ U.S.\ not\ intervene\ earlier\ in\ WW\ II"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{kill}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{deprive\ of\ life\ "AIDS\ has\ killed\ thousands\ in\ Africa"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{note}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ mention\ of\ "She\ observed\ that\ his\ presentation\ took\ up\ too\ much\ time"\ "They\ noted\ that\ it\ was\ a\ fine\ day\ to\ go\ sailing"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{49}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{suspend}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stop\ a\ process\ or\ a\ habit\ by\ imposing\ a\ freeze\ on\ it\ "Suspend\ the\ aid\ to\ the\ wartorn\ country"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{48}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{deploy}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ distribute\ systematically\ or\ strategically\ "The\ U.S.\ deploys\ its\ weapons\ in\ the\ Middle\ East"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{48}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{overthrow}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause\ the\ downfall\ of\ of\ rulers\ "The\ Czar\ was\ overthrown"\ "subvert\ the\ ruling\ class"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{48}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{resolve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ to\ an\ end\ settle\ conclusively\ "The\ case\ was\ decided"\ "The\ judge\ decided\ the\ case\ in\ favor\ of\ the\ plaintiff"\ "The\ father\ adjudicated\ when\ the\ sons\ were\ quarreling\ over\ their\ inheritance"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{48}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{state}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{express\ in\ words\ "He\ said\ that\ he\ wanted\ to\ marry\ her"\ "tell\ me\ what\ is\ bothering\ you"\ "state\ your\ opinion"\ "state\ your\ name"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{47}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accept}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{give\ an\ affirmative\ reply\ to\ respond\ favorably\ to\ "I\ cannot\ accept\ your\ invitation"\ "I\ go\ for\ this\ resolution"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{47}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{form}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{develop\ into\ a\ distinctive\ entity\ "our\ plans\ began\ to\ take\ shape"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{47}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{use}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{avail\ oneself\ to\ "apply\ a\ principle"\ "practice\ a\ religion"\ "use\ care\ when\ going\ down\ the\ stairs"\ "use\ your\ common\ sense"\ "practice\ nonviolent\ resistance"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{46}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{allege}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{report\ or\ maintain\ "He\ alleged\ that\ he\ was\ the\ victim\ of\ a\ crime"\ "He\ said\ it\ was\ too\ late\ to\ intervene\ in\ the\ war"\ "The\ registrar\ says\ that\ I\ owe\ the\ school\ money"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{46}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{approve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{give\ sanction\ to\ "I\ approve\ of\ his\ educational\ policies"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{46}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{arrive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reach\ a\ destination\ arrive\ by\ movement\ or\ progress\ "She\ arrived\ home\ at\ 7\ oclock"\ "She\ didnt\ get\ to\ Chicago\ until\ after\ midnight"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{46}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{break}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ a\ rupture\ in\ the\ ranks\ of\ the\ enemy\ or\ ones\ own\ by\ quitting\ or\ fleeing\ "The\ ranks\ broke"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{46}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{resume}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ up\ or\ begin\ anew\ "We\ resumed\ the\ negotiations"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{45}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{negotiate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{discuss\ the\ terms\ of\ an\ arrangement\ "They\ negotiated\ the\ sale\ of\ the\ house"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{45}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{protect}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{shield\ from\ danger,\ injury,\ destruction,\ or\ damage\ "Weatherbeater\ protects\ your\ roof\ from\ the\ rain"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{45}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{protract}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{lengthen\ in\ time\ cause\ to\ be\ or\ last\ longer\ "We\ prolonged\ our\ stay"\ "She\ extended\ her\ visit\ by\ another\ day"\ "The\ meeting\ was\ drawn\ out\ until\ midnight"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{44}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{follow}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ bring\ something\ about\ at\ a\ later\ time\ than\ "She\ followed\ dinner\ with\ a\ brandy"\ "He\ followed\ his\ lecture\ with\ a\ question\ and\ answer\ period"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{44}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{leave}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{remove\ oneself\ from\ an\ association\ with\ or\ participation\ in\ "She\ wants\ to\ leave"\ "The\ teenager\ left\ home"\ "She\ left\ her\ position\ with\ the\ Red\ Cross"\ "He\ left\ the\ Senate\ after\ two\ terms"\ "after\ 20\ years\ with\ the\ same\ company,\ she\ pulled\ up\ stakes"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accept}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{react\ favorably\ to\ consider\ right\ and\ proper\ "People\ did\ not\ accept\ atonal\ music\ at\ that\ time"\ "We\ accept\ the\ idea\ of\ universal\ health\ care"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{comply}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{act\ in\ accordance\ with\ someones\ rules,\ commands,\ or\ wishes\ "He\ complied\ with\ my\ instructions"\ "You\ must\ comply\ or\ else"\ "Follow\ these\ simple\ rules"\ "abide\ by\ the\ rules"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{generate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ into\ existence\ "The\ new\ manager\ generated\ a\ lot\ of\ problems"\ "The\ computer\ bug\ generated\ chaos\ in\ the\ office"\ "The\ computer\ generated\ this\ image"\ "The\ earthquake\ generated\ a\ tsunami"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{increase}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{become\ bigger\ or\ greater\ in\ amount\ "The\ amount\ of\ work\ increased"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{initiate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{set\ in\ motion,\ start\ an\ event\ or\ prepare\ the\ way\ for\ "Hitlers\ attack\ on\ Poland\ led\ up\ to\ World\ War\ II"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{meet}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{experience\ as\ a\ reaction\ "My\ proposal\ met\ with\ much\ opposition"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{refuse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{show\ unwillingness\ towards\ "he\ declined\ to\ join\ the\ group\ on\ a\ hike"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{43}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{succeed}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{attain\ success\ or\ reach\ a\ desired\ goal\ "The\ enterprise\ succeeded"\ "We\ succeeded\ in\ getting\ tickets\ to\ the\ show"\ "she\ struggled\ to\ overcome\ her\ handicap\ and\ won"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{42}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{come}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ received\ "News\ came\ in\ of\ the\ massacre\ in\ Rwanda"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{42}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{launch}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{propel\ with\ force\ "launch\ the\ space\ shuttle"\ "Launch\ a\ ship"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{42}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{move}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{follow\ a\ procedure\ or\ take\ a\ course\ "We\ should\ go\ farther\ in\ this\ matter"\ "She\ went\ through\ a\ lot\ of\ trouble"\ "go\ about\ the\ world\ in\ a\ certain\ manner"\ "Messages\ must\ go\ through\ diplomatic\ channels"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{42}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reaffirm}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{affirm\ once\ again\ "He\ reaffirmed\ his\ faith\ in\ the\ church"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{42}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ into\ consideration\ for\ exemplifying\ purposes\ "Take\ the\ case\ of\ China"\ "Consider\ the\ following\ case"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{41}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{adopt}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ up\ the\ cause,\ ideology,\ practice,\ method,\ of\ someone\ and\ use\ it\ as\ ones\ own\ "She\ embraced\ Catholicism"\ "They\ adopted\ the\ Jewish\ faith"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{41}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{control}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{exercise\ authoritative\ control\ or\ power\ over\ "control\ the\ budget"\ "Command\ the\ military\ forces"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{41}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{proclaim}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{state\ or\ announce\ "I\ am\ not\ a\ Communist,\ he\ exclaimed"\ "The\ King\ will\ proclaim\ an\ amnesty"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{41}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{threaten}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{pose\ a\ threat\ to\ present\ a\ danger\ to\ "The\ pollution\ is\ endangering\ the\ crops"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{40}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bomb}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{throw\ bombs\ at\ or\ attack\ with\ bombs\ "The\ Americans\ bombed\ Dresden"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{40}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{last}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue\ to\ live\ and\ avoid\ dying\ "We\ went\ without\ water\ and\ food\ for\ 3\ days"\ "These\ superstitions\ survive\ in\ the\ backwaters\ of\ America"\ "The\ race\ car\ driver\ lived\ through\ several\ very\ serious\ accidents"\ "how\ long\ can\ a\ person\ last\ without\ food\ and\ water"\ "One\ crash\ victim\ died,\ the\ other\ lived"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{39}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{implement}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{pursue\ to\ a\ conclusion\ or\ bring\ to\ a\ successful\ issue\ "Did\ he\ go\ through\ with\ the\ treatment"\ "He\ implemented\ a\ new\ economic\ plan"\ "She\ followed\ up\ his\ recommendations\ with\ a\ written\ proposal"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{38}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{strengthen}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ strong\ or\ stronger\ "This\ exercise\ will\ strengthen\ your\ upper\ body"\ "strengthen\ the\ relations\ between\ the\ two\ countries"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{consider}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{think\ about\ carefully\ weigh\ "They\ considered\ the\ possibility\ of\ a\ strike"\ "Turn\ the\ proposal\ over\ in\ your\ mind"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{decide}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reach,\ make,\ or\ come\ to\ a\ decision\ about\ something\ "We\ finally\ decided\ after\ lengthy\ deliberations"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{do}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{carry\ on\ or\ function\ "We\ could\ do\ with\ a\ little\ more\ help\ around\ here"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause\ to\ move\ cause\ to\ be\ in\ a\ certain\ position\ or\ condition\ "He\ got\ his\ squad\ on\ the\ ball"\ "This\ let\ me\ in\ for\ a\ big\ surprise"\ "He\ got\ a\ girl\ into\ trouble"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{help}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{contribute\ to\ the\ furtherance\ of\ "This\ money\ will\ help\ the\ development\ of\ literacy\ in\ developing\ countries"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{hold}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ and\ maintain\ control\ over,\ often\ by\ violent\ means\ "The\ dissatisfied\ students\ held\ the\ Presidents\ office\ for\ almost\ a\ week"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{impose}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{impose\ and\ collect\ "levy\ a\ fine"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{indicate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ a\ signal\ for\ or\ a\ symptom\ of\ "These\ symptoms\ indicate\ a\ serious\ illness"\ "Her\ behavior\ points\ to\ a\ severe\ neurosis"\ "The\ economic\ indicators\ signal\ that\ the\ euro\ is\ undervalued"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{37}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{order}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{issue\ commands\ or\ orders\ for}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{36}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{become}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{enter\ or\ assume\ a\ certain\ state\ or\ condition\ "He\ became\ annoyed\ when\ he\ heard\ the\ bad\ news"\ "It\ must\ be\ getting\ more\ serious"\ "her\ face\ went\ red\ with\ anger"\ "She\ went\ into\ ecstasy"\ "Get\ going"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{36}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{participate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{become\ a\ participant\ be\ involved\ in\ "enter\ a\ race"\ "enter\ an\ agreement"\ "enter\ a\ drug\ treatment\ program"\ "enter\ negotiations"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{35}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{call}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{order,\ summon,\ or\ request\ for\ a\ specific\ duty\ or\ activity,\ work,\ role\ "He\ was\ already\ called\ 4\ times\ for\ jury\ duty"\ "They\ called\ him\ to\ active\ military\ duty"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{35}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause\ to\ do\ cause\ to\ act\ in\ a\ specified\ manner\ "The\ ads\ induced\ me\ to\ buy\ a\ VCR"\ "My\ children\ finally\ got\ me\ to\ buy\ a\ computer"\ "My\ wife\ made\ me\ buy\ a\ new\ sofa"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{35}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{enmesh}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{entangle\ or\ catch\ in\ or\ as\ if\ in\ a\ mesh}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{35}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{be\ confronted\ with\ "What\ do\ we\ have\ here"\ "Now\ we\ have\ a\ fine\ mess"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{35}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{give\ certain\ properties\ to\ something\ "get\ someone\ mad"\ "She\ made\ us\ look\ silly"\ "He\ made\ a\ fool\ of\ himself\ at\ the\ meeting"\ "Dont\ make\ this\ into\ a\ big\ deal"\ "This\ invention\ will\ make\ you\ a\ millionaire"\ "Make\ yourself\ clear"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{evacuate}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{move\ out\ of\ an\ unsafe\ location\ into\ safety\ "After\ the\ earthquake,\ residents\ were\ evacuated"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fire}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{drive\ out\ or\ away\ by\ or\ as\ if\ by\ fire\ "The\ soldiers\ were\ fired"\ "Surrender\ fires\ the\ cold\ skepticism"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{intend}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have\ in\ mind\ as\ a\ purpose\ "I\ mean\ no\ harm"\ "I\ only\ meant\ to\ help\ you"\ "She\ didnt\ think\ to\ harm\ me"\ "We\ thought\ to\ return\ early\ that\ night"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{lead}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{preside\ over\ "John\ moderated\ the\ discussion"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{refuse}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{resist\ immunologically\ the\ introduction\ of\ some\ foreign\ tissue\ or\ organ\ "His\ body\ rejected\ the\ liver\ of\ the\ donor"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{warn}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{notify\ of\ danger,\ potential\ harm,\ or\ risk\ "The\ director\ warned\ him\ that\ he\ might\ be\ fired"\ "The\ doctor\ warned\ me\ about\ the\ dangers\ of\ smoking"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{bring\ into\ a\ different\ state\ "this\ may\ land\ you\ in\ jail"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{provide}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{make\ a\ possibility\ or\ provide\ opportunity\ for\ permit\ to\ be\ attainable\ or\ cause\ to\ remain\ "This\ leaves\ no\ room\ for\ improvement"\ "The\ evidence\ allows\ only\ one\ conclusion"\ "allow\ for\ mistakes"\ "leave\ lots\ of\ time\ for\ the\ trip"\ "This\ procedure\ provides\ for\ lots\ of\ leeway"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{respond}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{show\ a\ response\ or\ a\ reaction\ to\ something}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{shoot}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{kill\ by\ firing\ a\ missile}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stop}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{hold\ back,\ as\ of\ a\ danger\ or\ an\ enemy\ check\ the\ expansion\ or\ influence\ of\ "Arrest\ the\ downward\ trend"\ "Check\ the\ growth\ of\ communism\ in\ South\ East\ Asia"\ "Contain\ the\ rebel\ movement"\ "Turn\ back\ the\ tide\ of\ communism"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{33}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{use}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{seek\ or\ achieve\ an\ end\ by\ using\ to\ ones\ advantage\ "She\ uses\ her\ influential\ friends\ to\ get\ jobs"\ "The\ presidents\ wife\ used\ her\ good\ connections"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accord}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{go\ together\ "The\ colors\ dont\ harmonize"\ "Their\ ideas\ concorded"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{broker}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{act\ as\ a\ broker}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{continue}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{span\ an\ interval\ of\ distance,\ space\ or\ time\ "The\ war\ extended\ over\ five\ years"\ "The\ period\ covered\ the\ turn\ of\ the\ century"\ "My\ land\ extends\ over\ the\ hills\ on\ the\ horizon"\ "This\ farm\ covers\ some\ 200\ acres"\ "The\ Archipelago\ continues\ for\ another\ 500\ miles"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fail}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fall\ short\ in\ what\ is\ expected\ "She\ failed\ in\ her\ obligations\ as\ a\ good\ daughterinlaw"\ "We\ must\ not\ fail\ his\ obligation\ to\ the\ victims\ of\ the\ Holocaust"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{flee}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{run\ away\ quickly\ "He\ threw\ down\ his\ gun\ and\ fled"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{involve}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{connect\ closely\ and\ often\ incriminatingly\ "This\ new\ ruling\ affects\ your\ business"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{32}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{maintain}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{keep\ in\ safety\ and\ protect\ from\ harm,\ decay,\ loss,\ or\ destruction\ "We\ preserve\ these\ archeological\ findings"\ "The\ old\ lady\ could\ not\ keep\ up\ the\ building"\ "children\ must\ be\ taught\ to\ conserve\ our\ national\ heritage"\ "The\ museum\ curator\ conserved\ the\ ancient\ manuscripts"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{adopt}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{choose\ and\ follow\ as\ of\ theories,\ ideas,\ policies,\ strategies\ or\ plans\ "She\ followed\ the\ feminist\ movement"\ "The\ candidate\ espouses\ Republican\ ideals"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{assist}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{give\ help\ or\ assistance\ be\ of\ service\ "Everyone\ helped\ out\ during\ the\ earthquake"\ "Can\ you\ help\ me\ carry\ this\ table"\ "She\ never\ helps\ around\ the\ house"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{fail}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{disappoint,\ prove\ undependable\ to\ abandon,\ forsake\ "His\ sense\ of\ smell\ failed\ him\ this\ time"\ "His\ strength\ finally\ failed\ him"\ "His\ children\ failed\ him\ in\ the\ crisis"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{follow}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{accept\ and\ follow\ the\ leadership\ or\ command\ or\ guidance\ of\ "Lets\ follow\ our\ great\ helmsman"\ "She\ followed\ a\ guru\ for\ years"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{force}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{take\ by\ force\ "Storm\ the\ fort"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{hold}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{to\ close\ within\ bounds,\ or\ otherwise\ limit\ or\ deprive\ of\ free\ movement\ "This\ holds\ the\ local\ until\ the\ express\ passengers\ change\ trains"\ "About\ a\ dozen\ animals\ were\ held\ inside\ the\ stockade"\ "The\ illegal\ immigrants\ were\ held\ at\ a\ detention\ center"\ "The\ terrorists\ held\ the\ journalists\ for\ ransom"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{monitor}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{keep\ tabs\ on\ keep\ an\ eye\ on\ keep\ under\ surveillance\ "we\ are\ monitoring\ the\ air\ quality"\ "the\ police\ monitor\ the\ suspects\ moves"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{move}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{change\ location\ move,\ travel,\ or\ proceed,\ also\ metaphorically\ "How\ fast\ does\ your\ new\ car\ go"\ "We\ travelled\ from\ Rome\ to\ Naples\ by\ bus"\ "The\ policemen\ went\ from\ door\ to\ door\ looking\ for\ the\ suspect"\ "The\ soldiers\ moved\ towards\ the\ city\ in\ an\ attempt\ to\ take\ it\ before\ night\ fell"\ "news\ travelled\ fast"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{order}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{give\ instructions\ to\ or\ direct\ somebody\ to\ do\ something\ with\ authority\ "I\ said\ to\ him\ to\ go\ home"\ "She\ ordered\ him\ to\ do\ the\ shopping"\ "The\ mother\ told\ the\ child\ to\ get\ dressed"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{31}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{receive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{receive\ a\ specified\ treatment\ abstract\ "These\ aspects\ of\ civilization\ do\ not\ find\ expression\ or\ receive\ an\ interpretation"\ "His\ movie\ received\ a\ good\ review"\ "I\ got\ nothing\ but\ trouble\ for\ my\ good\ intentions"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{30}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{engage}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{carry\ on\ wars,\ battles,\ or\ campaigns\ "Napoleon\ and\ Hitler\ waged\ war\ against\ all\ of\ Europe"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{30}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{express}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{send\ by\ rapid\ transport\ or\ special\ messenger\ service\ "She\ expressed\ the\ letter\ to\ Florida"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{30}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{impose}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{impose\ something\ unpleasant\ "The\ principal\ visited\ his\ rage\ on\ the\ students"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{30}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{kill}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{mark\ for\ deletion,\ rub\ off,\ or\ erase\ "kill\ these\ lines\ in\ the\ Presidents\ speech"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{30}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{seek}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{try\ to\ locate\ or\ discover,\ or\ try\ to\ establish\ the\ existence\ of\ "The\ police\ are\ searching\ for\ clues"\ "They\ are\ searching\ for\ the\ missing\ man\ in\ the\ entire\ county"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{conclude}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{reach\ agreement\ on\ "They\ concluded\ an\ economic\ agreement"\ "We\ concluded\ a\ ceasefire"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{decide}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{cause\ to\ decide\ "This\ new\ development\ finally\ decided\ me"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{include}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{add\ as\ part\ of\ something\ else\ put\ in\ as\ part\ of\ a\ set,\ group,\ or\ category\ "We\ must\ include\ this\ chemical\ element\ in\ the\ group"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{plan}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have\ the\ will\ and\ intention\ to\ carry\ out\ some\ action\ "He\ plans\ to\ be\ in\ graduate\ school\ next\ year"\ "The\ rebels\ had\ planned\ turmoil\ and\ confusion"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{protest}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{express\ opposition\ through\ action\ or\ words\ "dissent\ to\ the\ laws\ of\ the\ country"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{receive}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{express\ willingness\ to\ have\ in\ ones\ home\ or\ environs\ "The\ community\ warmly\ received\ the\ refugees"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{regard}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{connect\ closely\ and\ often\ incriminatingly\ "This\ new\ ruling\ affects\ your\ business"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{say}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{report\ or\ maintain\ "He\ alleged\ that\ he\ was\ the\ victim\ of\ a\ crime"\ "He\ said\ it\ was\ too\ late\ to\ intervene\ in\ the\ war"\ "The\ registrar\ says\ that\ I\ owe\ the\ school\ money"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{stand}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\cellcolor[HTML]{EFEFEF}\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{have\ or\ maintain\ a\ position\ or\ stand\ on\ an\ issue\ "Where\ do\ you\ stand\ on\ the\ War"}}} \\

\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedleft}p{\dimexpr 0.3in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{29}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep+0\arrayrulewidth}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{undermine}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 6in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\fontsize{6}{6}\selectfont{\textcolor[HTML]{000000}{destroy\ property\ or\ hinder\ normal\ operations\ "The\ Resistance\ sabotaged\ railroad\ operations\ during\ the\ war"}}} \\

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=2pt}-}

\end{longtable}

\clearpage

\hypertarget{appendix-b-coding-process}{%
\section{Appendix B: Coding Process}\label{appendix-b-coding-process}}

\hphantom{em}

\hypertarget{app:gui}{%
\subsection{Figure B1: Graphical User Interface}\label{app:gui}}

\hphantom{em}

\begin{figure}[H]
\caption{Graphical User Interface (GUI) for coding ICBe\label{fig:coding_gui}}
\centering{\includegraphics[width=17cm]{./gui_example}}
\end{figure}
\clearpage

\hypertarget{app:algo1}{%
\subsection{Algorithm B2: Coding Aggregation and
Filtering}\label{app:algo1}}

\hphantom{em}

\begin{algorithm}[H]
\SetAlgoLined
\KwData{Coder-Sentence-Tokens}
\KwResult{Sentence-Events}
\ForEach{sentence in sentences}{
  \ForEach{concept in concepts}{
    \ForEach{token in tokens}{
      \lIf{expert votes==0}{reject; \tcc*[f]{Require at least one expert vote}}
      \uElseIf{expert votes >= expert voters/2}{accept; \tcc*[f]{And a majority of expert votes}}
      \uElseIf{novice votes >= novice voters/2}{accept; \tcc*[f]{Or a majority of novice votes}}
    }
    \If{sum(accepted) == 0}{
      sort tags by votes;
      head 1;
      accept; \tcc*[f]{If still no tag, take one with most votes}
    }
  }
  \ForEach{ActorSet in unique(ActorSets)}{
    emit event;
  }
}
\caption{Aggregating Coder-Tokens into Sentence-Events}
\end{algorithm}
\clearpage

\hypertarget{app:intercoderagreement}{%
\subsection{Figure B3 : Intercoder
Agreement}\label{app:intercoderagreement}}

\hphantom{em}

\begin{figure}[H]
\caption{Intercoder Agreement by Concept/Type \label{fig:intercoderagreement}}
\centering{\includegraphics[width=15cm]{./p_percent_chose_tag_by_concept.png}}
\end{figure}
\clearpage

\hypertarget{app:agreement_by_confidence}{%
\subsection{Figure B4 : Intercoder Agreement by
Confidence}\label{app:agreement_by_confidence}}

\hphantom{em}

\begin{figure}[H]
\caption{Intercoder Agreement by Concept/Type \label{fig:intercoderagreement_confidence}}
\centering{\includegraphics[width=15cm]{./p_percent_chose_tag_by_confidence.png}}
\end{figure}
\clearpage

\hypertarget{appendix-c-recall}{%
\section{Appendix C: Recall}\label{appendix-c-recall}}

\hphantom{em}

\hypertarget{algorithm-c1-synthetic-narratives}{%
\subsection{Algorithm C1: Synthetic
Narratives}\label{algorithm-c1-synthetic-narratives}}

\hphantom{em}

A synthetic narrative is constructed as follows. First, compile a large
corpus, \(D\), of timelines, encyclopedia entries, journal articles,
news reports, websites, and government documents. Next, segment into
sentences, \(s \in S\), clean, filter, and break into simpler concise
sentences, \(s^* \in S^*\), using a large natural language model, BART
(Lewis et al. 2019), fine-tuned on summarization tasks,
distilbart-cnn-12-6 (Shleifer and Rush 2020). Then, calculate semantic
similarity between each \(S*\) by embedding them with a different large
language model, MPNet (Song et al., n.d.), fine tuned to identify
paraphrasing (Reimers and Gurevych 2019, 2020). Next, constructed an
undirected weighted graph of pair above 95th percentile cosine
similarity and hierarchically cluster them into potential events
\(e^* \in E^*\) (Clauset, Newman, and Moore 2004). The resulting set of
\(E^*\) are small enough to check by hand, summarize, and place in
chronological order. The relevant formulas and algorithm are shown
below.

Cosine Similarity \begin{equation}
cosine\;similarity = \frac{a \cdot b}{ || a  ||\;|| b  || }
\end{equation}

Adjacency Matrix \begin{equation} \label{eq:sclrec2}
A_{vw}= \begin{cases} 1 \text{ if vertices v and w are connected} \\ 0 \text{ otherwise } \end{cases}
\end{equation}

Identity Function \begin{equation}
\delta(i,j)= \begin{cases} 1 \text{ if i=j} \\ 0 \text{ otherwise } \end{cases}
\end{equation}

Degree of a vertex \begin{equation}
k_v = \sum_{w}A_{vw}
\end{equation}

Modularity - Is the difference between observed within-community edges
and average expected edges on a random graph
\begin{equation} \label{eq:sclrec2}
Q = \frac{1}{2m}\sum_{vw}[A_{vw}-\frac{k_vk_w}{2m}]\delta(C_w,C_w)
\end{equation}

\begin{algorithm}[H]
\SetAlgoLined
\KwData{Document-Sentences}
\KwResult{Event-Mentions}
\tcc{ Embed text as a 768 dimensional semantic vector with a large language model}
\ForEach{document in Document}{
  \ForEach{sentence in sentence}{
    embedding=distilbart-cnn-6-6(sentence);
  }
}

similarity = cosine similarity(embeddings,embeddings);

\tcc{ Threshold similarity matrix and create weighted undirected graph }

\tcc{ Frame as adding edges so efficient nearest neighbor lookup can be used }

g=EmptyUndirectedWeightedGraph();\tcc*[f]{Sparse undirected graph to store edges}

\ForEach{v in 1:rows(similarity)}{
  \ForEach{w in 1:cols(similarity)}{ 
    \lIf{similarity[v,w]>0.96 and doc(V)!=doc(W)}{append edge v<->w to g}\tcc*[f]{If similar append as undirected edge to graph}
  }
}

\tcc{ Community detection with greedy hierarchically clustering (Clauset et al. 2004)}

gClustered=EmptyUndirectedGraph(vertices(g)); \tcc*[f]{Sparse undirected graph to store edges}
dendrogram=list()
\While{length(edges(g))>1}{

  \ForEach{edge in edges(g)}{
    Qpossible[edge]=Q(gClustered + edge)
  }

  \ForEach{ edge in edges(g) } {
    \tcc{ Add the edge with the maximum increase in modularity }
    \tcc{ Remove from original graph as a housekeeping trick }
    \lIf{ QPossible[edge]=max(QPossible) }{ append edge to gClustered and remove edge from g; append QPossible to dendrogram }
  }
}
\tcc{ Cut the dendrogram to maximize modularity and return event clusters}
\Return dendrogram where Q(dendrogram)==max(dendrogram) \tcc*[f]{}
\caption{Clustering and counting sentences about the same event}
\end{algorithm}
\clearpage

\hypertarget{table-c2-synthetic-narrative-matches-cuban-missile}{%
\subsection{Table C2: Synthetic Narrative Matches Cuban
Missile}\label{table-c2-synthetic-narrative-matches-cuban-missile}}

\hphantom{em}

\begin{figure}[H]
\caption{ Cuban Missile Crisis Synthetic Narrative Recall Matches \label{fig:ft_cuba_recall}}
\centering{\includegraphics[height=21cm]{./ft_cuba_recall.png}}
\end{figure}
\clearpage

\hypertarget{table-c3-synthetic-narrative-matches-crimea-donbas}{%
\subsection{Table C3: Synthetic Narrative Matches
Crimea-Donbas}\label{table-c3-synthetic-narrative-matches-crimea-donbas}}

\hphantom{em}

\begin{figure}[H]
\caption{ Crimea-Donbas Synthetic Narrative Recall Matches \label{fig:ft_crimea_recall}}
\centering{\includegraphics[height=21cm]{./ft_crimea_recall.png}}
\end{figure}
\clearpage

\hypertarget{appendix-d-precision}{%
\section{Appendix D: Precision}\label{appendix-d-precision}}

\hypertarget{figure-d1-crimea-donbos}{%
\subsection{Figure D1 Crimea-Donbos}\label{figure-d1-crimea-donbos}}

\hphantom{em}

\begin{figure}[H]
\caption{ Annexation of Crimea ICB Narrative and ICBe codings \label{fig:case_study_crimea_precision}}
\centering{\includegraphics[height=19cm]{./case_study_crimea_precision_part1.png}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[height=21cm]{./case_study_crimea_precision_part2.png}}
\end{figure}
\clearpage

\hypertarget{table-d2-stratified-sentence-examples}{%
\subsection{Table D2: Stratified Sentence
Examples}\label{table-d2-stratified-sentence-examples}}

\hphantom{em}
\begin{figure}[H]
\caption{ Stratified Sentence Examples  \label{fig:ft_sample_precision_precision1}}
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision1.png"}}
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision2.png"}}
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision3.png"}}
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision4.png"}}
\end{figure}
\clearpage

\hypertarget{figure-d3-crisis-maps-cuban-missiles}{%
\subsection{Figure D3 Crisis Maps Cuban
Missiles}\label{figure-d3-crisis-maps-cuban-missiles}}

\hphantom{em}

\begin{sideways}%[htbp]
\centering{\includegraphics[width=20cm]{./p_196_combined.png}}
\end{sideways}

\hphantom{em}
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps - Cuban Missiles - Phoenix Disaggregated \label{fig:p_phoenix_plot_lv0_196}}
\centering{\includegraphics[width=20cm]{./p_phoenix_plot_lv0_196.png}}
\end{sidewaysfigure}
\clearpage

\hypertarget{figure-d4-crisis-maps-crimea-donbas}{%
\subsection{Figure D4 Crisis Maps
Crimea-Donbas}\label{figure-d4-crisis-maps-crimea-donbas}}

\hphantom{em}
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps - Crimea-Donbas \label{fig:p_471_combined}}
\centering{\includegraphics[width=19cm]{./p_471_combined.png}}
\end{sidewaysfigure}

\hphantom{em}
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps - Crimea-Donbas Phoenix Disaggregated \label{fig:p_phoenix_plot_lv0_471}}
\centering{\includegraphics[width=20cm]{./p_phoenix_plot_lv0_471.png}}
\end{sidewaysfigure}

\hphantom{em}
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps - Crimea-Donbas Terrier Disaggregated \label{fig:p_terrier_plot_lv0_471}}
\centering{\includegraphics[width=20cm]{./p_terrier_plot_lv0_471.png}}
\end{sidewaysfigure}

\hphantom{em}
\begin{sidewaysfigure}[ht]
\caption{Crisis Maps - Crimea-Donbas ICEWs Disaggregated \label{fig:p_icews_plot_lv0_471}}
\centering{\includegraphics[width=20cm]{./p_icews_plot_lv0_471.png}}
\end{sidewaysfigure}
\clearpage

\hypertarget{references}{%
\section{References}\label{references}}

\hphantom{em}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-clausetFindingCommunityStructure2004}{}}%
Clauset, Aaron, M. E. J. Newman, and Cristopher Moore. 2004. {``Finding
Community Structure in Very Large Networks.''} \emph{Physical Review E}
70 (6): 066111. \url{https://doi.org/10.1103/PhysRevE.70.066111}.

\leavevmode\vadjust pre{\hypertarget{ref-lewisBARTDenoisingSequencetoSequence2019}{}}%
Lewis, Mike, Yinhan Liu, Naman Goyal, Marjan Ghazvininejad, Abdelrahman
Mohamed, Omer Levy, Ves Stoyanov, and Luke Zettlemoyer. 2019. {``{BART}:
{Denoising Sequence-to-Sequence Pre-training} for {Natural Language
Generation}, {Translation}, and {Comprehension}.''} {arXiv}.
\url{https://doi.org/10.48550/arXiv.1910.13461}.

\leavevmode\vadjust pre{\hypertarget{ref-reimersSentenceBERTSentenceEmbeddings2019}{}}%
Reimers, Nils, and Iryna Gurevych. 2019. {``Sentence-{BERT}: {Sentence
Embeddings} Using {Siamese BERT-Networks}.''}
\url{http://arxiv.org/abs/1908.10084}.

\leavevmode\vadjust pre{\hypertarget{ref-reimersMakingMonolingualSentence2020}{}}%
---------. 2020. {``Making {Monolingual Sentence Embeddings
Multilingual} Using {Knowledge Distillation}.''}
\url{http://arxiv.org/abs/2004.09813}.

\leavevmode\vadjust pre{\hypertarget{ref-shleiferPretrainedSummarizationDistillation2020}{}}%
Shleifer, Sam, and Alexander M. Rush. 2020. {``Pre-Trained
{Summarization Distillation}.''} \url{http://arxiv.org/abs/2010.13002}.

\leavevmode\vadjust pre{\hypertarget{ref-songMPNetMaskedPermuted}{}}%
Song, Kaitao, Xu Tan, Tao Qin, Jianfeng Lu, and Tie-Yan Liu. n.d.
{``{MPNet}: {Masked} and {Permuted Pre-training} for {Language
Understanding},''} 14.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{paper.bib}

\end{document}