### 2023-08-19 replication_paper/psrm_draft/ICBe_supplemental_information.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Introducing ICBe: Supplementary Information}

\author{
    Rex W. Douglass
    \thanks{Correspondence should be addressed to Rex W. Douglass at
\href{mailto:rexdouglass@gmail.com}{\nolinkurl{rexdouglass@gmail.com}}.}
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Thomas Leo Scherer
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    J. Andrés Gannon
   \\
    Vanderbilt University \\
   \\
  \texttt{} \\
   \And
    Erik Gartzke
   \\
    University of California, San Diego \\
   \\
  \texttt{} \\
   \And
    Jon Lindsay
   \\
    Georgia Institute of Technology \\
   \\
  \texttt{} \\
   \And
    Shannon Carcelli
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Jonathan Wilkenfeld
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    David M. Quinn
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Catherine Aiken
   \\
    Georgetown University \\
   \\
  \texttt{} \\
   \And
    Jose Miguel Cabezas Navarro
   \\
    Universidad Mayor \\
   \\
  \texttt{} \\
   \And
    Neil Lund
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Egle Murauskaite
   \\
    University of Maryland \\
   \\
  \texttt{} \\
   \And
    Diana Partridge
   \\
    University of Maryland \\
   \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[utf8]{inputenc}
\usepackage{pifont}
\usepackage{newunicodechar}
\newunicodechar{✓}{\ding{51}}
\newunicodechar{✗}{\ding{55}}
\usepackage{array}
\usepackage{ctable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{float}
\usepackage{setspace}
\usepackage{cellspace}
\setlength\cellspacetoplimit{0.8ex}
\renewcommand{\arraystretch}{0.8}
\AtBeginEnvironment{lltable}{\singlespacing}
\usepackage{subcaption}
\captionsetup[algorithm]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{atbegshi}% http://ctan.org/pkg/atbegshi
\usepackage{float}
\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{changepage}
\renewenvironment{abstract}{}{}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

\newpage
\tableofcontents 
\newpage

\clearpage

\hypertarget{corpus-and-cases}{%
\section{Corpus and Cases}\label{corpus-and-cases}}

\hypertarget{app:crises}{%
\subsection{ICBe Crises and Summary Information}\label{app:crises}}

\begin{figure}[H]
\caption{ICBe Crises Summary Statistics}
\centering{\includegraphics[ height=19cm]{"./crisis_text_counts_ft1.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft2.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft3.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft4.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft5.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft6.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft7.png"}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[ height=21cm]{"./crisis_text_counts_ft8.png"}}
\end{figure}
\clearpage

\hypertarget{app:verbs}{%
\subsection{Verb Meanings (Glosses) found in the ICB Corpus (top
200)}\label{app:verbs}}

\hphantom{em}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{0.75}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{|p{0.30in}|p{0.75in}|p{6.00in}}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{uses}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Lemma}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Gloss}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}\endfirsthead

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{uses}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Lemma}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{Gloss}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}\endhead

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{381}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{end}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ to\ an\ end\ or\ halt\ "She\ ended\ their\ friendship\ when\ she\ found\ out\ that\ he\ had\ once\ been\ convicted\ of\ a\ crime"\ "The\ attack\ on\ Poland\ terminated\ the\ relatively\ peaceful\ period\ after\ WW\ I"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{365}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{trigger}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{release\ or\ pull\ the\ trigger\ on\ "Trigger\ a\ gun"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{323}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{involve}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{engage\ as\ a\ participant\ "Dont\ involve\ me\ in\ your\ family\ affairs"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{309}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{begin}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have\ a\ beginning,\ of\ a\ temporal\ event\ "WW\ II\ began\ in\ 1939\ when\ Hitler\ marched\ into\ Poland"\ "The\ companys\ Asia\ tour\ begins\ next\ month"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{278}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{trigger}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{put\ in\ motion\ or\ move\ to\ act\ "trigger\ a\ reaction"\ "actuate\ the\ circuits"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{260}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{announce}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{announce\ publicly\ or\ officially\ "The\ President\ declared\ war"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{236}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{respond}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{respond\ favorably\ or\ as\ hoped\ "The\ cancer\ responded\ to\ the\ aggressive\ therapy"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{212}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{call}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{call\ a\ meeting\ invite\ or\ command\ to\ meet\ "The\ Wannsee\ Conference\ was\ called\ to\ discuss\ the\ Final\ Solution"\ "The\ new\ dean\ calls\ meetings\ every\ week"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{194}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{occur}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{come\ to\ pass\ "What\ is\ happening"\ "The\ meeting\ took\ place\ off\ without\ an\ incidence"\ "Nothing\ occurred\ that\ seemed\ important"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{194}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{sign}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{approve\ and\ express\ assent,\ responsibility,\ or\ obligation\ "All\ parties\ ratified\ the\ peace\ treaty"\ "Have\ you\ signed\ your\ contract\ yet"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{185}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{terminate}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ to\ an\ end\ or\ halt\ "She\ ended\ their\ friendship\ when\ she\ found\ out\ that\ he\ had\ once\ been\ convicted\ of\ a\ crime"\ "The\ attack\ on\ Poland\ terminated\ the\ relatively\ peaceful\ period\ after\ WW\ I"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{178}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{begin}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{set\ in\ motion,\ cause\ to\ start\ "The\ U.S.\ started\ a\ war\ in\ the\ Middle\ East"\ "The\ Iraqis\ began\ hostilities"\ "begin\ a\ new\ chapter\ in\ your\ life"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{160}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{meet}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{come\ together\ "Ill\ probably\ see\ you\ at\ the\ meeting"\ "How\ nice\ to\ see\ you\ again"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{146}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{send}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ cause\ or\ order\ to\ be\ taken,\ directed,\ or\ transmitted\ to\ another\ place\ "He\ had\ sent\ the\ dispatches\ downtown\ to\ the\ proper\ people\ and\ had\ slept"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{145}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{occupy}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{march\ aggressively\ into\ anothers\ territory\ by\ military\ force\ for\ the\ purposes\ of\ conquest\ and\ occupation\ "Hitler\ invaded\ Poland\ on\ September\ 1,\ 1939"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{135}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{happen,\ occur,\ take\ place\ "I\ lost\ my\ wallet\ this\ was\ during\ the\ visit\ to\ my\ parents\ house"\ "There\ were\ two\ hundred\ people\ at\ his\ funeral"\ "There\ was\ a\ lot\ of\ noise\ in\ the\ kitchen"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{128}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{perceive}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{become\ conscious\ of\ "She\ finally\ perceived\ the\ futility\ of\ her\ protest"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{126}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{become}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{undergo\ a\ change\ or\ development\ "The\ water\ turned\ into\ ice"\ "Her\ former\ friend\ became\ her\ worst\ enemy"\ "He\ turned\ traitor"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{125}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{condemn}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{express\ strong\ disapproval\ of\ "We\ condemn\ the\ racism\ in\ South\ Africa"\ "These\ ideas\ were\ reprobated"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{125}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ by\ force\ "Hitler\ took\ the\ Baltic\ Republics"\ "The\ army\ took\ the\ fort\ on\ the\ hill"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{124}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{escalate}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{increase\ in\ extent\ or\ intensity\ "The\ Allies\ escalated\ the\ bombing"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{119}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reject}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{refuse\ to\ accept\ or\ acknowledge\ "I\ reject\ the\ idea\ of\ starting\ a\ war"\ "The\ journal\ rejected\ the\ students\ paper"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{117}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{respond}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{react\ verbally\ "She\ didnt\ want\ to\ answer"\ "answer\ the\ question"\ "We\ answered\ that\ we\ would\ accept\ the\ invitation"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{116}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{demand}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{request\ urgently\ and\ forcefully\ "The\ victims\ family\ is\ demanding\ compensation"\ "The\ boss\ demanded\ that\ he\ be\ fired\ immediately"\ "She\ demanded\ to\ see\ the\ manager"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{112}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ remain\ unmolested,\ undisturbed,\ or\ uninterrupted\ \ used\ only\ in\ infinitive\ form\ "let\ her\ be"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{111}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{agree}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ in\ accord\ be\ in\ agreement\ "We\ agreed\ on\ the\ terms\ of\ the\ settlement"\ "I\ cant\ agree\ with\ you"\ "I\ hold\ with\ those\ who\ say\ life\ is\ sacred"\ "Both\ philosophers\ concord\ on\ this\ point"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{110}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reach}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ in\ or\ establish\ communication\ with\ "Our\ advertisements\ reach\ millions"\ "He\ never\ contacted\ his\ children\ after\ he\ emigrated\ to\ Australia"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{106}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{withdraw}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{pull\ back\ or\ move\ away\ or\ backward\ "The\ enemy\ withdrew"\ "The\ limo\ pulled\ away\ from\ the\ curb"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{105}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accuse}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{blame\ for,\ make\ a\ claim\ of\ wrongdoing\ or\ misbehavior\ against\ "he\ charged\ the\ director\ with\ indifference"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{105}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{dispute}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have\ a\ disagreement\ over\ something\ "We\ quarreled\ over\ the\ question\ as\ to\ who\ discovered\ America"\ "These\ two\ fellows\ are\ always\ scrapping\ over\ something"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{105}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fight}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ engaged\ in\ a\ fight\ carry\ on\ a\ fight\ "the\ tribesmen\ fought\ each\ other"\ "Siblings\ are\ always\ fighting"\ "Militant\ groups\ are\ contending\ for\ control\ of\ the\ country"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{102}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue\ talking\ "I\ know\ its\ hard,\ he\ continued,\ but\ there\ is\ no\ choice"\ "carry\ onpretend\ we\ are\ not\ in\ the\ room"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{100}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{include}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{consider\ as\ part\ of\ something\ "I\ include\ you\ in\ the\ list\ of\ culprits"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{99}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{remain}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ left\ of\ persons,\ questions,\ problems,\ results,\ evidence,\ etc.\ "There\ remains\ the\ question\ of\ who\ pulled\ the\ trigger"\ "Carter\ remains\ the\ only\ President\ in\ recent\ history\ under\ whose\ Presidency\ the\ U.S.\ did\ not\ fight\ a\ war"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{98}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{experience}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{undergo\ or\ live\ through\ a\ difficult\ experience\ "We\ had\ many\ trials\ to\ go\ through"\ "he\ saw\ action\ in\ Viet\ Nam"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{97}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{attempt}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ an\ effort\ or\ attempt\ "He\ tried\ to\ shake\ off\ his\ fears"\ "The\ infant\ had\ essayed\ a\ few\ wobbly\ steps"\ "The\ police\ attempted\ to\ stop\ the\ thief"\ "He\ sought\ to\ improve\ himself"\ "She\ always\ seeks\ to\ do\ good\ in\ the\ world"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{94}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{increase}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ bigger\ or\ more\ "The\ boss\ finally\ increased\ her\ salary"\ "The\ university\ increased\ the\ number\ of\ students\ it\ admitted"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{94}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{mediate}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{act\ between\ parties\ with\ a\ view\ to\ reconciling\ differences\ "He\ interceded\ in\ the\ family\ dispute"\ "He\ mediated\ a\ settlement"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{94}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{threaten}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ utter\ intentions\ of\ injury\ or\ punishment\ against\ "He\ threatened\ me\ when\ I\ tried\ to\ call\ the\ police"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{93}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{request}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{inquire\ for\ information\ "I\ requested\ information\ from\ the\ secretary"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{92}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{withdraw}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ a\ retreat\ from\ an\ earlier\ commitment\ or\ activity\ "Well\ have\ to\ crawfish\ out\ from\ meeting\ with\ him"\ "He\ backed\ out\ of\ his\ earlier\ promise"\ "The\ aggressive\ investment\ company\ pulled\ in\ its\ horns"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{91}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{hold}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stop\ dealing\ with\ "hold\ all\ calls\ to\ the\ Presidents\ office\ while\ he\ is\ in\ a\ meeting"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{90}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{invade}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{march\ aggressively\ into\ anothers\ territory\ by\ military\ force\ for\ the\ purposes\ of\ conquest\ and\ occupation\ "Hitler\ invaded\ Poland\ on\ September\ 1,\ 1939"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{87}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{send}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause\ to\ go\ somewhere\ "The\ explosion\ sent\ the\ car\ flying\ in\ the\ air"\ "She\ sent\ her\ children\ to\ camp"\ "He\ directed\ all\ his\ energies\ into\ his\ dissertation"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{86}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{launch}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{begin\ with\ vigor\ "He\ launched\ into\ a\ long\ diatribe"\ "She\ plunged\ into\ a\ dangerous\ adventure"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{84}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{report}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{announce\ as\ the\ result\ of\ an\ investigation\ or\ experience\ or\ finding\ "Dozens\ of\ incidents\ of\ wife\ beatings\ are\ reported\ daily\ in\ this\ city"\ "The\ team\ reported\ significant\ advances\ in\ their\ research"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{84}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{see}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{perceive\ or\ be\ contemporaneous\ with\ "We\ found\ Republicans\ winning\ the\ offices"\ "Youll\ see\ a\ lot\ of\ cheating\ in\ this\ school"\ "The\ 1960s\ saw\ the\ rebellion\ of\ the\ younger\ generation\ against\ established\ traditions"\ "I\ want\ to\ see\ results"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{81}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{perceive}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ become\ aware\ of\ through\ the\ senses\ "I\ could\ perceive\ the\ ship\ coming\ over\ the\ horizon"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{80}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accuse}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ an\ accusation\ against\ level\ a\ charge\ against\ "The\ neighbors\ accused\ the\ man\ of\ spousal\ abuse"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{77}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{attack}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{launch\ an\ attack\ or\ assault\ on\ begin\ hostilities\ or\ start\ warfare\ with\ "Hitler\ attacked\ Poland\ on\ September\ 1,\ 1939\ and\ started\ World\ War\ II"\ "Serbian\ forces\ assailed\ Bosnian\ towns\ all\ week"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{77}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{dispatch}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{kill\ without\ delay\ "the\ traitor\ was\ dispatched\ by\ the\ conspirators"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{77}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{end}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{put\ an\ end\ to\ "The\ terrible\ news\ ended\ our\ hopes\ that\ he\ had\ survived"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{74}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{declare}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{announce\ publicly\ or\ officially\ "The\ President\ declared\ war"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{74}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{discuss}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{speak\ with\ others\ about\ something\ talk\ something\ over\ in\ detail\ have\ a\ discussion\ "We\ discussed\ our\ household\ budget"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{74}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{follow}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{keep\ under\ surveillance\ "The\ police\ had\ been\ following\ him\ for\ weeks\ but\ they\ could\ not\ prove\ his\ involvement\ in\ the\ bombing"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{73}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{lead}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{tend\ to\ or\ result\ in\ "This\ remark\ lead\ to\ further\ arguments\ among\ the\ guests"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{73}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{remain}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stay\ behind\ "The\ smell\ stayed\ in\ the\ room"\ "The\ hostility\ remained\ long\ after\ they\ made\ up"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{72}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{propose}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{present\ for\ consideration,\ examination,\ criticism,\ etc.\ "He\ proposed\ a\ new\ plan\ for\ dealing\ with\ terrorism"\ "She\ proposed\ a\ new\ theory\ of\ relativity"\ "The\ candidate\ projects\ himself\ as\ a\ moderate\ and\ a\ reformer"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{72}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{recognize}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accept\ someone\ to\ be\ what\ is\ claimed\ or\ accept\ his\ power\ and\ authority\ "The\ Crown\ Prince\ was\ acknowledged\ as\ the\ true\ heir\ to\ the\ throne"\ "We\ do\ not\ recognize\ your\ gods"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{70}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{capture}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ possession\ of\ by\ force,\ as\ after\ an\ invasion\ "the\ invaders\ seized\ the\ land\ and\ property\ of\ the\ inhabitants"\ "The\ army\ seized\ the\ town"\ "The\ militia\ captured\ the\ castle"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{70}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{deny}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{deny\ formally\ an\ allegation\ of\ fact\ by\ the\ opposing\ party\ in\ a\ legal\ suit}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{69}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{exist\ over\ a\ prolonged\ period\ of\ time\ "The\ bad\ weather\ continued\ for\ two\ more\ weeks"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{69}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{issue}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ out\ an\ official\ document\ such\ as\ a\ warrant}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{67}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{declare}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{proclaim\ ones\ support,\ sympathy,\ or\ opinion\ for\ or\ against\ "His\ wife\ declared\ at\ once\ for\ moving\ to\ the\ West\ Coast"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{66}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{lead}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stretch\ out\ over\ a\ distance,\ space,\ time,\ or\ scope\ run\ or\ extend\ between\ two\ points\ or\ beyond\ a\ certain\ point\ "Service\ runs\ all\ the\ way\ to\ Cranbury"\ "His\ knowledge\ doesnt\ go\ very\ far"\ "My\ memory\ extends\ back\ to\ my\ fourth\ year\ of\ life"\ "The\ facts\ extend\ beyond\ a\ consideration\ of\ her\ personal\ assets"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{65}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{renew}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reestablish\ on\ a\ new,\ usually\ improved,\ basis\ or\ make\ new\ or\ like\ new\ "We\ renewed\ our\ friendship\ after\ a\ hiatus\ of\ twenty\ years"\ "They\ renewed\ their\ membership"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{64}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{achieve}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ gain\ with\ effort\ "she\ achieved\ her\ goal\ despite\ setbacks"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{63}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{provide}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{determine\ what\ is\ to\ happen\ in\ certain\ contingencies,\ especially\ by\ including\ a\ proviso\ condition\ or\ stipulation\ "The\ will\ provides\ that\ each\ child\ should\ receive\ half\ of\ the\ money"\ "The\ Constitution\ provides\ for\ the\ right\ to\ free\ speech"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{61}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{appeal}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ a\ court\ case\ to\ a\ higher\ court\ for\ review\ "He\ was\ found\ guilty\ but\ appealed\ immediately"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{61}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{seek}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ an\ effort\ or\ attempt\ "He\ tried\ to\ shake\ off\ his\ fears"\ "The\ infant\ had\ essayed\ a\ few\ wobbly\ steps"\ "The\ police\ attempted\ to\ stop\ the\ thief"\ "He\ sought\ to\ improve\ himself"\ "She\ always\ seeks\ to\ do\ good\ in\ the\ world"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{60}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{attack}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ the\ initiative\ and\ go\ on\ the\ offensive\ "The\ Serbs\ attacked\ the\ village\ at\ night"\ "The\ visiting\ team\ started\ to\ attack"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{60}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{pass}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{go\ unchallenged\ be\ approved\ "The\ bill\ cleared\ the\ House"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{59}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{agree}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{achieve\ harmony\ of\ opinion,\ feeling,\ or\ purpose\ "No\ two\ of\ my\ colleagues\ would\ agree\ on\ whom\ to\ elect\ chairman"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{59}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{last}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{persist\ for\ a\ specified\ period\ of\ time\ "The\ bad\ weather\ lasted\ for\ three\ days"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{58}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{claim}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{demand\ as\ being\ ones\ due\ or\ property\ assert\ ones\ right\ or\ title\ to\ "He\ claimed\ his\ suitcases\ at\ the\ airline\ counter"\ "Mr.\ Smith\ claims\ special\ tax\ exemptions\ because\ he\ is\ a\ foreign\ resident"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{57}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{establish}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ about\ "The\ trompe\ loeilillusion\ establishes\ depth"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{57}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{inform}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{impart\ knowledge\ of\ some\ fact,\ state\ of\ affairs,\ or\ event\ to\ "I\ informed\ him\ of\ his\ rights"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{57}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{warn}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{notify,\ usually\ in\ advance\ "I\ warned\ you\ that\ I\ would\ ask\ some\ difficult\ questions"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{56}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{offer}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{threaten\ to\ do\ something\ "I\ offered\ to\ leave\ the\ committee\ if\ they\ did\ not\ accept\ my\ proposal"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{56}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{restore}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ back\ into\ original\ existence,\ use,\ function,\ or\ position\ "restore\ law\ and\ order"\ "reestablish\ peace\ in\ the\ region"\ "restore\ the\ emperor\ to\ the\ throne"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{55}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{create}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ or\ cause\ to\ be\ or\ to\ become\ "make\ a\ mess\ in\ ones\ office"\ "create\ a\ furor"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{55}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{issue}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{circulate\ or\ distribute\ or\ equip\ with\ "issue\ a\ new\ uniform\ to\ the\ children"\ "supply\ blankets\ for\ the\ beds"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{55}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{prevent}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{keep\ from\ happening\ or\ arising\ make\ impossible\ "My\ sense\ of\ tact\ forbids\ an\ honest\ answer"\ "Your\ role\ in\ the\ projects\ precludes\ your\ involvement\ in\ the\ competitive\ project"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{55}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{support}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{argue\ or\ speak\ in\ defense\ of\ "She\ supported\ the\ motion\ to\ strike"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{54}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cross}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{travel\ across\ or\ pass\ over\ "The\ caravan\ covered\ almost\ 100\ miles\ each\ day"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{54}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{include}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{allow\ participation\ in\ or\ the\ right\ to\ be\ part\ of\ permit\ to\ exercise\ the\ rights,\ functions,\ and\ responsibilities\ of\ "admit\ someone\ to\ the\ profession"\ "She\ was\ admitted\ to\ the\ New\ Jersey\ Bar"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{54}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{urge}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{push\ for\ something\ "The\ travel\ agent\ recommended\ strongly\ that\ we\ not\ travel\ on\ Thanksgiving\ Day"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{53}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{agree}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{consent\ or\ assent\ to\ a\ condition,\ or\ agree\ to\ do\ something\ "She\ agreed\ to\ all\ my\ conditions"\ "He\ agreed\ to\ leave\ her\ alone"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{53}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{do\ something\ repeatedly\ and\ showing\ no\ intention\ to\ stop\ "We\ continued\ our\ research\ into\ the\ cause\ of\ the\ illness"\ "The\ landlord\ persists\ in\ asking\ us\ to\ move"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{53}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{send}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{transfer\ "The\ spy\ sent\ the\ classified\ information\ off\ to\ Russia"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{53}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{support}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ behind\ approve\ of\ "He\ plumped\ for\ the\ Labor\ Party"\ "I\ backed\ Kennedy\ in\ 1960"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{52}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{erupt}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{erupt\ or\ intensify\ suddenly\ "Unrest\ erupted\ in\ the\ country"\ "Tempers\ flared\ at\ the\ meeting"\ "The\ crowd\ irrupted\ into\ a\ burst\ of\ patriotism"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{52}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{resolve}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reach\ a\ conclusion\ after\ a\ discussion\ or\ deliberation}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{52}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{view}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{deem\ to\ be\ "She\ views\ this\ quite\ differently\ from\ me"\ "I\ consider\ her\ to\ be\ shallow"\ "I\ dont\ see\ the\ situation\ quite\ as\ negatively\ as\ you\ do"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{50}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{occur}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ be\ found\ to\ exist\ "sexism\ occurs\ in\ many\ workplaces"\ "precious\ stones\ occur\ in\ a\ large\ area\ in\ Brazil"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{50}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{admit\ into\ a\ group\ or\ community\ "accept\ students\ for\ graduate\ study"\ "Well\ have\ to\ vote\ on\ whether\ or\ not\ to\ admit\ a\ new\ member"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{claim}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ as\ an\ undesirable\ consequence\ of\ some\ event\ or\ state\ of\ affairs\ "the\ accident\ claimed\ three\ lives"\ "The\ hard\ work\ took\ its\ toll\ on\ her"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fade}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{disappear\ gradually\ "The\ pain\ eventually\ passed\ off"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{intervene}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{get\ involved,\ so\ as\ to\ alter\ or\ hinder\ an\ action,\ or\ through\ force\ or\ threat\ of\ force\ "Why\ did\ the\ U.S.\ not\ intervene\ earlier\ in\ WW\ II"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{kill}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{deprive\ of\ life\ "AIDS\ has\ killed\ thousands\ in\ Africa"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{note}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ mention\ of\ "She\ observed\ that\ his\ presentation\ took\ up\ too\ much\ time"\ "They\ noted\ that\ it\ was\ a\ fine\ day\ to\ go\ sailing"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{49}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{suspend}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stop\ a\ process\ or\ a\ habit\ by\ imposing\ a\ freeze\ on\ it\ "Suspend\ the\ aid\ to\ the\ wartorn\ country"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{48}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{deploy}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ distribute\ systematically\ or\ strategically\ "The\ U.S.\ deploys\ its\ weapons\ in\ the\ Middle\ East"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{48}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{overthrow}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause\ the\ downfall\ of\ of\ rulers\ "The\ Czar\ was\ overthrown"\ "subvert\ the\ ruling\ class"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{48}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{resolve}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ to\ an\ end\ settle\ conclusively\ "The\ case\ was\ decided"\ "The\ judge\ decided\ the\ case\ in\ favor\ of\ the\ plaintiff"\ "The\ father\ adjudicated\ when\ the\ sons\ were\ quarreling\ over\ their\ inheritance"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{48}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{state}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{express\ in\ words\ "He\ said\ that\ he\ wanted\ to\ marry\ her"\ "tell\ me\ what\ is\ bothering\ you"\ "state\ your\ opinion"\ "state\ your\ name"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{47}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accept}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{give\ an\ affirmative\ reply\ to\ respond\ favorably\ to\ "I\ cannot\ accept\ your\ invitation"\ "I\ go\ for\ this\ resolution"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{47}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{form}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{develop\ into\ a\ distinctive\ entity\ "our\ plans\ began\ to\ take\ shape"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{47}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{use}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{avail\ oneself\ to\ "apply\ a\ principle"\ "practice\ a\ religion"\ "use\ care\ when\ going\ down\ the\ stairs"\ "use\ your\ common\ sense"\ "practice\ nonviolent\ resistance"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{46}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{allege}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{report\ or\ maintain\ "He\ alleged\ that\ he\ was\ the\ victim\ of\ a\ crime"\ "He\ said\ it\ was\ too\ late\ to\ intervene\ in\ the\ war"\ "The\ registrar\ says\ that\ I\ owe\ the\ school\ money"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{46}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{approve}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{give\ sanction\ to\ "I\ approve\ of\ his\ educational\ policies"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{46}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{arrive}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reach\ a\ destination\ arrive\ by\ movement\ or\ progress\ "She\ arrived\ home\ at\ 7\ oclock"\ "She\ didnt\ get\ to\ Chicago\ until\ after\ midnight"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{46}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{break}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ a\ rupture\ in\ the\ ranks\ of\ the\ enemy\ or\ ones\ own\ by\ quitting\ or\ fleeing\ "The\ ranks\ broke"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{46}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{resume}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ up\ or\ begin\ anew\ "We\ resumed\ the\ negotiations"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{45}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{negotiate}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{discuss\ the\ terms\ of\ an\ arrangement\ "They\ negotiated\ the\ sale\ of\ the\ house"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{45}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{protect}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{shield\ from\ danger,\ injury,\ destruction,\ or\ damage\ "Weatherbeater\ protects\ your\ roof\ from\ the\ rain"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{45}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{protract}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{lengthen\ in\ time\ cause\ to\ be\ or\ last\ longer\ "We\ prolonged\ our\ stay"\ "She\ extended\ her\ visit\ by\ another\ day"\ "The\ meeting\ was\ drawn\ out\ until\ midnight"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{44}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{follow}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ bring\ something\ about\ at\ a\ later\ time\ than\ "She\ followed\ dinner\ with\ a\ brandy"\ "He\ followed\ his\ lecture\ with\ a\ question\ and\ answer\ period"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{44}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{leave}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{remove\ oneself\ from\ an\ association\ with\ or\ participation\ in\ "She\ wants\ to\ leave"\ "The\ teenager\ left\ home"\ "She\ left\ her\ position\ with\ the\ Red\ Cross"\ "He\ left\ the\ Senate\ after\ two\ terms"\ "after\ 20\ years\ with\ the\ same\ company,\ she\ pulled\ up\ stakes"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accept}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{react\ favorably\ to\ consider\ right\ and\ proper\ "People\ did\ not\ accept\ atonal\ music\ at\ that\ time"\ "We\ accept\ the\ idea\ of\ universal\ health\ care"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{comply}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{act\ in\ accordance\ with\ someones\ rules,\ commands,\ or\ wishes\ "He\ complied\ with\ my\ instructions"\ "You\ must\ comply\ or\ else"\ "Follow\ these\ simple\ rules"\ "abide\ by\ the\ rules"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{generate}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ into\ existence\ "The\ new\ manager\ generated\ a\ lot\ of\ problems"\ "The\ computer\ bug\ generated\ chaos\ in\ the\ office"\ "The\ computer\ generated\ this\ image"\ "The\ earthquake\ generated\ a\ tsunami"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{increase}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{become\ bigger\ or\ greater\ in\ amount\ "The\ amount\ of\ work\ increased"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{initiate}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{set\ in\ motion,\ start\ an\ event\ or\ prepare\ the\ way\ for\ "Hitlers\ attack\ on\ Poland\ led\ up\ to\ World\ War\ II"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{meet}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{experience\ as\ a\ reaction\ "My\ proposal\ met\ with\ much\ opposition"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{refuse}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{show\ unwillingness\ towards\ "he\ declined\ to\ join\ the\ group\ on\ a\ hike"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{43}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{succeed}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{attain\ success\ or\ reach\ a\ desired\ goal\ "The\ enterprise\ succeeded"\ "We\ succeeded\ in\ getting\ tickets\ to\ the\ show"\ "she\ struggled\ to\ overcome\ her\ handicap\ and\ won"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{42}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{come}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ received\ "News\ came\ in\ of\ the\ massacre\ in\ Rwanda"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{42}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{launch}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{propel\ with\ force\ "launch\ the\ space\ shuttle"\ "Launch\ a\ ship"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{42}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{move}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{follow\ a\ procedure\ or\ take\ a\ course\ "We\ should\ go\ farther\ in\ this\ matter"\ "She\ went\ through\ a\ lot\ of\ trouble"\ "go\ about\ the\ world\ in\ a\ certain\ manner"\ "Messages\ must\ go\ through\ diplomatic\ channels"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{42}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reaffirm}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{affirm\ once\ again\ "He\ reaffirmed\ his\ faith\ in\ the\ church"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{42}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ into\ consideration\ for\ exemplifying\ purposes\ "Take\ the\ case\ of\ China"\ "Consider\ the\ following\ case"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{41}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{adopt}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ up\ the\ cause,\ ideology,\ practice,\ method,\ of\ someone\ and\ use\ it\ as\ ones\ own\ "She\ embraced\ Catholicism"\ "They\ adopted\ the\ Jewish\ faith"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{41}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{control}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{exercise\ authoritative\ control\ or\ power\ over\ "control\ the\ budget"\ "Command\ the\ military\ forces"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{41}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{proclaim}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{state\ or\ announce\ "I\ am\ not\ a\ Communist,\ he\ exclaimed"\ "The\ King\ will\ proclaim\ an\ amnesty"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{41}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{threaten}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{pose\ a\ threat\ to\ present\ a\ danger\ to\ "The\ pollution\ is\ endangering\ the\ crops"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{40}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bomb}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{throw\ bombs\ at\ or\ attack\ with\ bombs\ "The\ Americans\ bombed\ Dresden"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{40}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{last}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue\ to\ live\ and\ avoid\ dying\ "We\ went\ without\ water\ and\ food\ for\ 3\ days"\ "These\ superstitions\ survive\ in\ the\ backwaters\ of\ America"\ "The\ race\ car\ driver\ lived\ through\ several\ very\ serious\ accidents"\ "how\ long\ can\ a\ person\ last\ without\ food\ and\ water"\ "One\ crash\ victim\ died,\ the\ other\ lived"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{39}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{implement}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{pursue\ to\ a\ conclusion\ or\ bring\ to\ a\ successful\ issue\ "Did\ he\ go\ through\ with\ the\ treatment"\ "He\ implemented\ a\ new\ economic\ plan"\ "She\ followed\ up\ his\ recommendations\ with\ a\ written\ proposal"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{38}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{strengthen}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ strong\ or\ stronger\ "This\ exercise\ will\ strengthen\ your\ upper\ body"\ "strengthen\ the\ relations\ between\ the\ two\ countries"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{consider}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{think\ about\ carefully\ weigh\ "They\ considered\ the\ possibility\ of\ a\ strike"\ "Turn\ the\ proposal\ over\ in\ your\ mind"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{decide}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reach,\ make,\ or\ come\ to\ a\ decision\ about\ something\ "We\ finally\ decided\ after\ lengthy\ deliberations"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{do}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{carry\ on\ or\ function\ "We\ could\ do\ with\ a\ little\ more\ help\ around\ here"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause\ to\ move\ cause\ to\ be\ in\ a\ certain\ position\ or\ condition\ "He\ got\ his\ squad\ on\ the\ ball"\ "This\ let\ me\ in\ for\ a\ big\ surprise"\ "He\ got\ a\ girl\ into\ trouble"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{help}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{contribute\ to\ the\ furtherance\ of\ "This\ money\ will\ help\ the\ development\ of\ literacy\ in\ developing\ countries"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{hold}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ and\ maintain\ control\ over,\ often\ by\ violent\ means\ "The\ dissatisfied\ students\ held\ the\ Presidents\ office\ for\ almost\ a\ week"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{impose}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{impose\ and\ collect\ "levy\ a\ fine"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{indicate}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ a\ signal\ for\ or\ a\ symptom\ of\ "These\ symptoms\ indicate\ a\ serious\ illness"\ "Her\ behavior\ points\ to\ a\ severe\ neurosis"\ "The\ economic\ indicators\ signal\ that\ the\ euro\ is\ undervalued"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{37}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{order}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{issue\ commands\ or\ orders\ for}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{36}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{become}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{enter\ or\ assume\ a\ certain\ state\ or\ condition\ "He\ became\ annoyed\ when\ he\ heard\ the\ bad\ news"\ "It\ must\ be\ getting\ more\ serious"\ "her\ face\ went\ red\ with\ anger"\ "She\ went\ into\ ecstasy"\ "Get\ going"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{36}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{participate}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{become\ a\ participant\ be\ involved\ in\ "enter\ a\ race"\ "enter\ an\ agreement"\ "enter\ a\ drug\ treatment\ program"\ "enter\ negotiations"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{35}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{call}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{order,\ summon,\ or\ request\ for\ a\ specific\ duty\ or\ activity,\ work,\ role\ "He\ was\ already\ called\ 4\ times\ for\ jury\ duty"\ "They\ called\ him\ to\ active\ military\ duty"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{35}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause\ to\ do\ cause\ to\ act\ in\ a\ specified\ manner\ "The\ ads\ induced\ me\ to\ buy\ a\ VCR"\ "My\ children\ finally\ got\ me\ to\ buy\ a\ computer"\ "My\ wife\ made\ me\ buy\ a\ new\ sofa"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{35}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{enmesh}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{entangle\ or\ catch\ in\ or\ as\ if\ in\ a\ mesh}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{35}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{be\ confronted\ with\ "What\ do\ we\ have\ here"\ "Now\ we\ have\ a\ fine\ mess"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{35}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{give\ certain\ properties\ to\ something\ "get\ someone\ mad"\ "She\ made\ us\ look\ silly"\ "He\ made\ a\ fool\ of\ himself\ at\ the\ meeting"\ "Dont\ make\ this\ into\ a\ big\ deal"\ "This\ invention\ will\ make\ you\ a\ millionaire"\ "Make\ yourself\ clear"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{evacuate}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{move\ out\ of\ an\ unsafe\ location\ into\ safety\ "After\ the\ earthquake,\ residents\ were\ evacuated"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fire}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{drive\ out\ or\ away\ by\ or\ as\ if\ by\ fire\ "The\ soldiers\ were\ fired"\ "Surrender\ fires\ the\ cold\ skepticism"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{intend}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have\ in\ mind\ as\ a\ purpose\ "I\ mean\ no\ harm"\ "I\ only\ meant\ to\ help\ you"\ "She\ didnt\ think\ to\ harm\ me"\ "We\ thought\ to\ return\ early\ that\ night"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{lead}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{preside\ over\ "John\ moderated\ the\ discussion"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{refuse}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{resist\ immunologically\ the\ introduction\ of\ some\ foreign\ tissue\ or\ organ\ "His\ body\ rejected\ the\ liver\ of\ the\ donor"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{34}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{warn}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{notify\ of\ danger,\ potential\ harm,\ or\ risk\ "The\ director\ warned\ him\ that\ he\ might\ be\ fired"\ "The\ doctor\ warned\ me\ about\ the\ dangers\ of\ smoking"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{bring\ into\ a\ different\ state\ "this\ may\ land\ you\ in\ jail"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{provide}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{make\ a\ possibility\ or\ provide\ opportunity\ for\ permit\ to\ be\ attainable\ or\ cause\ to\ remain\ "This\ leaves\ no\ room\ for\ improvement"\ "The\ evidence\ allows\ only\ one\ conclusion"\ "allow\ for\ mistakes"\ "leave\ lots\ of\ time\ for\ the\ trip"\ "This\ procedure\ provides\ for\ lots\ of\ leeway"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{respond}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{show\ a\ response\ or\ a\ reaction\ to\ something}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{shoot}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{kill\ by\ firing\ a\ missile}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stop}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{hold\ back,\ as\ of\ a\ danger\ or\ an\ enemy\ check\ the\ expansion\ or\ influence\ of\ "Arrest\ the\ downward\ trend"\ "Check\ the\ growth\ of\ communism\ in\ South\ East\ Asia"\ "Contain\ the\ rebel\ movement"\ "Turn\ back\ the\ tide\ of\ communism"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{33}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{use}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{seek\ or\ achieve\ an\ end\ by\ using\ to\ ones\ advantage\ "She\ uses\ her\ influential\ friends\ to\ get\ jobs"\ "The\ presidents\ wife\ used\ her\ good\ connections"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accord}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{go\ together\ "The\ colors\ dont\ harmonize"\ "Their\ ideas\ concorded"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{broker}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{act\ as\ a\ broker}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{continue}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{span\ an\ interval\ of\ distance,\ space\ or\ time\ "The\ war\ extended\ over\ five\ years"\ "The\ period\ covered\ the\ turn\ of\ the\ century"\ "My\ land\ extends\ over\ the\ hills\ on\ the\ horizon"\ "This\ farm\ covers\ some\ 200\ acres"\ "The\ Archipelago\ continues\ for\ another\ 500\ miles"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fail}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fall\ short\ in\ what\ is\ expected\ "She\ failed\ in\ her\ obligations\ as\ a\ good\ daughterinlaw"\ "We\ must\ not\ fail\ his\ obligation\ to\ the\ victims\ of\ the\ Holocaust"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{flee}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{run\ away\ quickly\ "He\ threw\ down\ his\ gun\ and\ fled"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{involve}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{connect\ closely\ and\ often\ incriminatingly\ "This\ new\ ruling\ affects\ your\ business"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{32}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{maintain}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{keep\ in\ safety\ and\ protect\ from\ harm,\ decay,\ loss,\ or\ destruction\ "We\ preserve\ these\ archeological\ findings"\ "The\ old\ lady\ could\ not\ keep\ up\ the\ building"\ "children\ must\ be\ taught\ to\ conserve\ our\ national\ heritage"\ "The\ museum\ curator\ conserved\ the\ ancient\ manuscripts"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{adopt}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{choose\ and\ follow\ as\ of\ theories,\ ideas,\ policies,\ strategies\ or\ plans\ "She\ followed\ the\ feminist\ movement"\ "The\ candidate\ espouses\ Republican\ ideals"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{assist}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{give\ help\ or\ assistance\ be\ of\ service\ "Everyone\ helped\ out\ during\ the\ earthquake"\ "Can\ you\ help\ me\ carry\ this\ table"\ "She\ never\ helps\ around\ the\ house"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{fail}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{disappoint,\ prove\ undependable\ to\ abandon,\ forsake\ "His\ sense\ of\ smell\ failed\ him\ this\ time"\ "His\ strength\ finally\ failed\ him"\ "His\ children\ failed\ him\ in\ the\ crisis"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{follow}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{accept\ and\ follow\ the\ leadership\ or\ command\ or\ guidance\ of\ "Lets\ follow\ our\ great\ helmsman"\ "She\ followed\ a\ guru\ for\ years"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{force}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{take\ by\ force\ "Storm\ the\ fort"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{hold}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{to\ close\ within\ bounds,\ or\ otherwise\ limit\ or\ deprive\ of\ free\ movement\ "This\ holds\ the\ local\ until\ the\ express\ passengers\ change\ trains"\ "About\ a\ dozen\ animals\ were\ held\ inside\ the\ stockade"\ "The\ illegal\ immigrants\ were\ held\ at\ a\ detention\ center"\ "The\ terrorists\ held\ the\ journalists\ for\ ransom"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{monitor}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{keep\ tabs\ on\ keep\ an\ eye\ on\ keep\ under\ surveillance\ "we\ are\ monitoring\ the\ air\ quality"\ "the\ police\ monitor\ the\ suspects\ moves"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{move}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{change\ location\ move,\ travel,\ or\ proceed,\ also\ metaphorically\ "How\ fast\ does\ your\ new\ car\ go"\ "We\ travelled\ from\ Rome\ to\ Naples\ by\ bus"\ "The\ policemen\ went\ from\ door\ to\ door\ looking\ for\ the\ suspect"\ "The\ soldiers\ moved\ towards\ the\ city\ in\ an\ attempt\ to\ take\ it\ before\ night\ fell"\ "news\ travelled\ fast"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{order}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{give\ instructions\ to\ or\ direct\ somebody\ to\ do\ something\ with\ authority\ "I\ said\ to\ him\ to\ go\ home"\ "She\ ordered\ him\ to\ do\ the\ shopping"\ "The\ mother\ told\ the\ child\ to\ get\ dressed"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{31}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{receive}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{receive\ a\ specified\ treatment\ abstract\ "These\ aspects\ of\ civilization\ do\ not\ find\ expression\ or\ receive\ an\ interpretation"\ "His\ movie\ received\ a\ good\ review"\ "I\ got\ nothing\ but\ trouble\ for\ my\ good\ intentions"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{30}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{engage}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{carry\ on\ wars,\ battles,\ or\ campaigns\ "Napoleon\ and\ Hitler\ waged\ war\ against\ all\ of\ Europe"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{30}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{express}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{send\ by\ rapid\ transport\ or\ special\ messenger\ service\ "She\ expressed\ the\ letter\ to\ Florida"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{30}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{impose}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{impose\ something\ unpleasant\ "The\ principal\ visited\ his\ rage\ on\ the\ students"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{30}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{kill}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{mark\ for\ deletion,\ rub\ off,\ or\ erase\ "kill\ these\ lines\ in\ the\ Presidents\ speech"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{30}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{seek}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{try\ to\ locate\ or\ discover,\ or\ try\ to\ establish\ the\ existence\ of\ "The\ police\ are\ searching\ for\ clues"\ "They\ are\ searching\ for\ the\ missing\ man\ in\ the\ entire\ county"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{conclude}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{reach\ agreement\ on\ "They\ concluded\ an\ economic\ agreement"\ "We\ concluded\ a\ ceasefire"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{decide}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{cause\ to\ decide\ "This\ new\ development\ finally\ decided\ me"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{include}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{add\ as\ part\ of\ something\ else\ put\ in\ as\ part\ of\ a\ set,\ group,\ or\ category\ "We\ must\ include\ this\ chemical\ element\ in\ the\ group"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{plan}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have\ the\ will\ and\ intention\ to\ carry\ out\ some\ action\ "He\ plans\ to\ be\ in\ graduate\ school\ next\ year"\ "The\ rebels\ had\ planned\ turmoil\ and\ confusion"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{protest}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{express\ opposition\ through\ action\ or\ words\ "dissent\ to\ the\ laws\ of\ the\ country"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{receive}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{express\ willingness\ to\ have\ in\ ones\ home\ or\ environs\ "The\ community\ warmly\ received\ the\ refugees"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{regard}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{connect\ closely\ and\ often\ incriminatingly\ "This\ new\ ruling\ affects\ your\ business"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{say}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{report\ or\ maintain\ "He\ alleged\ that\ he\ was\ the\ victim\ of\ a\ crime"\ "He\ said\ it\ was\ too\ late\ to\ intervene\ in\ the\ war"\ "The\ registrar\ says\ that\ I\ owe\ the\ school\ money"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{stand}}} & \multicolumn{1}{>{\cellcolor[HTML]{EFEFEF}\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{have\ or\ maintain\ a\ position\ or\ stand\ on\ an\ issue\ "Where\ do\ you\ stand\ on\ the\ War"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedleft}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{29}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{undermine}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{6}{6}\selectfont{destroy\ property\ or\ hinder\ normal\ operations\ "The\ Resistance\ sabotaged\ railroad\ operations\ during\ the\ war"}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\clearpage

\hypertarget{coding-process}{%
\section{Coding Process}\label{coding-process}}

\hphantom{em}

\hypertarget{app:gui}{%
\subsection{Graphical User Interface}\label{app:gui}}

\hphantom{em}

\begin{figure}[H]
\caption{Graphical User Interface (GUI) for coding ICBe\label{fig:coding_gui}}
\centering{\includegraphics[width=17cm]{./gui_example}}
\end{figure}
\clearpage

\hypertarget{app:algo1}{%
\subsection{Coding Aggregation and Filtering}\label{app:algo1}}

\hphantom{em}

\begin{algorithm}[H]
\SetAlgoLined
\KwData{Coder-Sentence-Tokens}
\KwResult{Sentence-Events}
\ForEach{sentence in sentences}{
  \ForEach{concept in concepts}{
    \ForEach{token in tokens}{
      \lIf{expert votes==0}{reject; \tcc*[f]{Require at least one expert vote}}
      \uElseIf{expert votes >= expert voters/2}{accept; \tcc*[f]{And a majority of expert votes}}
      \uElseIf{novice votes >= novice voters/2}{accept; \tcc*[f]{Or a majority of novice votes}}
    }
    \If{sum(accepted) == 0}{
      sort tags by votes;
      head 1;
      accept; \tcc*[f]{If still no tag, take one with most votes}
    }
  }
  \ForEach{ActorSet in unique(ActorSets)}{
    emit event;
  }
}
\caption{Aggregating Coder-Tokens into Sentence-Events}
\end{algorithm}
\clearpage

\hypertarget{app:intercoderagreement}{%
\subsection{Intercoder Agreement}\label{app:intercoderagreement}}

\hphantom{em}

\begin{figure}[H]
\caption{Intercoder Agreement by Concept/Type \label{fig:intercoderagreement}}
\centering{\includegraphics[width=15cm]{./p_percent_chose_tag_by_concept.png}}
\end{figure}
\clearpage

\hypertarget{app:agreement_by_confidence}{%
\subsection{Intercoder Agreement by
Confidence}\label{app:agreement_by_confidence}}

\hphantom{em}

\begin{figure}[H]
\caption{Intercoder Agreement by Concept/Type \label{fig:intercoderagreement_confidence}}
\centering{\includegraphics[width=15cm]{./p_percent_chose_tag_by_confidence.png}}
\end{figure}
\clearpage

\hypertarget{recall}{%
\section{Recall}\label{recall}}

\hypertarget{synthetic-narratives}{%
\subsection{Synthetic Narratives}\label{synthetic-narratives}}

\doublespacing

A synthetic narrative is constructed as follows. First, compile a large
corpus, \(D\), of timelines, encyclopedia entries, journal articles,
news reports, websites, and government documents. Next, segment into
sentences, \(s \in S\), clean, filter, and break into simpler concise
sentences, \(s^* \in S^*\), using a large natural language model, BART
(Lewis et al. 2019), fine-tuned on summarization tasks,
distilbart-cnn-12-6 (Shleifer and Rush 2020). Then, calculate semantic
similarity between each \(S*\) by embedding them with a different large
language model, MPNet (Song et al., n.d.), fine tuned to identify
paraphrasing (Reimers and Gurevych 2019, 2020). Next, constructed an
undirected weighted graph of pair above 95th percentile cosine
similarity and hierarchically cluster them into potential events
\(e^* \in E^*\) (Clauset, Newman, and Moore 2004). The resulting set of
\(E^*\) are small enough to check by hand, summarize, and place in
chronological order. The relevant formulas and algorithm are shown
below.

\singlespacing

Cosine Similarity \begin{equation}
cosine\;similarity = \frac{a \cdot b}{ || a  ||\;|| b  || }
\end{equation}

Adjacency Matrix \begin{equation} \label{eq:sclrec2}
A_{vw}= \begin{cases} 1 \text{ if vertices v and w are connected} \\ 0 \text{ otherwise } \end{cases}
\end{equation}

Identity Function \begin{equation}
\delta(i,j)= \begin{cases} 1 \text{ if i=j} \\ 0 \text{ otherwise } \end{cases}
\end{equation}

Degree of a vertex \begin{equation}
k_v = \sum_{w}A_{vw}
\end{equation}

Modularity - Is the difference between observed within-community edges
and average expected edges on a random graph
\begin{equation} \label{eq:sclrec2}
Q = \frac{1}{2m}\sum_{vw}[A_{vw}-\frac{k_vk_w}{2m}]\delta(C_w,C_w)
\end{equation}

\begin{algorithm}[H]
\SetAlgoLined
\KwData{Document-Sentences}
\KwResult{Event-Mentions}
\tcc{ Embed text as a 768 dimensional semantic vector with a large language model}
\ForEach{document in Document}{
  \ForEach{sentence in sentence}{
    embedding=distilbart-cnn-6-6(sentence);
  }
}

similarity = cosine similarity(embeddings,embeddings);

\tcc{ Threshold similarity matrix and create weighted undirected graph }

\tcc{ Frame as adding edges so efficient nearest neighbor lookup can be used }

g=EmptyUndirectedWeightedGraph();\tcc*[f]{Sparse undirected graph to store edges}

\ForEach{v in 1:rows(similarity)}{
  \ForEach{w in 1:cols(similarity)}{ 
    \lIf{similarity[v,w]>0.96 and doc(V)!=doc(W)}{append edge v<->w to g}\tcc*[f]{If similar append as undirected edge to graph}
  }
}

\tcc{ Community detection with greedy hierarchically clustering (Clauset et al. 2004)}

gClustered=EmptyUndirectedGraph(vertices(g)); \tcc*[f]{Sparse undirected graph to store edges}
dendrogram=list()
\While{length(edges(g))>1}{

  \ForEach{edge in edges(g)}{
    Qpossible[edge]=Q(gClustered + edge)
  }

  \ForEach{ edge in edges(g) } {
    \tcc{ Add the edge with the maximum increase in modularity }
    \tcc{ Remove from original graph as a housekeeping trick }
    \lIf{ QPossible[edge]=max(QPossible) }{ append edge to gClustered and remove edge from g; append QPossible to dendrogram }
  }
}
\tcc{ Cut the dendrogram to maximize modularity and return event clusters}
\Return dendrogram where Q(dendrogram)==max(dendrogram) \tcc*[f]{}
\caption{Clustering and counting sentences about the same event}
\end{algorithm}
\clearpage

\hypertarget{synthetic-narrative-matches-cuban-missile}{%
\subsection{Synthetic Narrative Matches Cuban
Missile}\label{synthetic-narrative-matches-cuban-missile}}

\hphantom{em}

\begin{figure}[H]
\caption{ Cuban Missile Crisis Synthetic Narrative Recall Matches \label{fig:ft_cuba_recall}}
\centering{\includegraphics[height=19cm]{./ft_cuba_recall.png}}
\end{figure}
\clearpage

\hypertarget{synthetic-narrative-matches-crimea-donbas}{%
\subsection{Synthetic Narrative Matches
Crimea-Donbas}\label{synthetic-narrative-matches-crimea-donbas}}

\hphantom{em}

\begin{figure}[H]
\caption{ Crimea-Donbas Synthetic Narrative Recall Matches \label{fig:ft_crimea_recall}}
\centering{\includegraphics[height=19cm]{./ft_crimea_recall.png}}
\end{figure}
\clearpage

\hypertarget{precision}{%
\section{Precision}\label{precision}}

\hypertarget{crimea-donbas-narrative}{%
\subsection{Crimea-Donbas Narrative}\label{crimea-donbas-narrative}}

\hphantom{em}

\begin{figure}[H]
\caption{ Annexation of Crimea ICB Narrative and ICBe codings \label{fig:case_study_crimea_precision}}
\centering{\includegraphics[height=19cm]{./case_study_crimea_precision_part1.png}}
\end{figure}
\clearpage

\begin{figure}[H]
\centering{\includegraphics[height=11cm]{./case_study_crimea_precision_part2.png}}
\end{figure}
\clearpage

\hypertarget{stratified-sentence-examples}{%
\subsection{Stratified Sentence
Examples}\label{stratified-sentence-examples}}

\hphantom{em}
\begin{figure}[H]
\caption{ Stratified Sentence Examples  \label{fig:ft_sample_precision_precision1}}
\centering{\includegraphics[height=15cm]{"./ft_sample_precision_precision1.png"}} 
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision2.png"}}
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision3.png"}}
\end{figure}
\clearpage

\hphantom{em}
\begin{figure}[H]
\centering{\includegraphics[height=21cm]{"./ft_sample_precision_precision4.png"}}
\end{figure}

\hypertarget{cuban-missiles-crisis-maps}{%
\subsection{Cuban Missiles Crisis
Maps}\label{cuban-missiles-crisis-maps}}

\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Cuban Missiles}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=1\paperwidth]{./p_196_combined.png}
\label{fig:p_196_combined}
\end{adjustbox}
\end{figure}
\clearpage

\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Cuban Missiles - Phoenix Disaggregated}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=1\paperwidth]{./p_phoenix_plot_lv0_196.png}
\label{fig:p_phoenix_plot_lv0_196}
\end{adjustbox}
\end{figure}

\hypertarget{crimea-donbas-crisis-maps}{%
\subsection{Crimea-Donbas Crisis Maps}\label{crimea-donbas-crisis-maps}}

\begin{figure}[H]
\caption{Crisis Map - Crimea-Donbas - ICBe \label{fig:p_471_icbe}}
\centering{\includegraphics[width=17cm]{./p_471_icbe.png}}
\end{figure}

\hphantom{em}
\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Crimea-Donbas}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=.9\paperwidth]{./p_471_combined.png}
\label{fig:p_471_combined}
\end{adjustbox}
\end{figure}

\hphantom{em}
\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Crimea-Donbas Phoenix Disaggregated}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=1\paperwidth]{./p_phoenix_plot_lv0_471.png}
\label{fig:p_phoenix_plot_lv0_471}
\end{adjustbox}
\end{figure}

\hphantom{em}
\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Crimea-Donbas Terrier Disaggregated}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=1\paperwidth]{./p_terrier_plot_lv0_471.png}
\label{fig:p_terrier_plot_lv0_471}
\end{adjustbox}
\end{figure}

\hphantom{em}
\begin{figure}[H]
\begin{adjustbox}{
  addcode={\begin{minipage}{\width}
           \caption{Crisis Maps - Crimea-Donbas ICEWs Disaggregated}}
          {\end{minipage}},
  rotate=90}
\includegraphics[width=1\paperwidth]{./p_icews_plot_lv0_471.png}
\label{fig:p_icews_plot_lv0_471}
\end{adjustbox}
\end{figure}

\clearpage

\hypertarget{references}{%
\section{References}\label{references}}

\hphantom{em}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-clausetFindingCommunityStructure2004}{}}%
Clauset, Aaron, M. E. J. Newman, and Cristopher Moore. 2004. {``Finding
Community Structure in Very Large Networks.''} \emph{Physical Review E}
70 (6): 066111. \url{https://doi.org/10.1103/PhysRevE.70.066111}.

\leavevmode\vadjust pre{\hypertarget{ref-lewisBARTDenoisingSequencetoSequence2019}{}}%
Lewis, Mike, Yinhan Liu, Naman Goyal, Marjan Ghazvininejad, Abdelrahman
Mohamed, Omer Levy, Ves Stoyanov, and Luke Zettlemoyer. 2019. {``{BART}:
{Denoising Sequence-to-Sequence Pre-training} for {Natural Language
Generation}, {Translation}, and {Comprehension}.''} {arXiv}.
\url{https://doi.org/10.48550/arXiv.1910.13461}.

\leavevmode\vadjust pre{\hypertarget{ref-reimersSentenceBERTSentenceEmbeddings2019}{}}%
Reimers, Nils, and Iryna Gurevych. 2019. {``Sentence-{BERT}: {Sentence
Embeddings} Using {Siamese BERT-Networks}.''} \emph{arXiv:1908.10084
{[}Cs{]}}, August. \url{https://arxiv.org/abs/1908.10084}.

\leavevmode\vadjust pre{\hypertarget{ref-reimersMakingMonolingualSentence2020}{}}%
---------. 2020. {``Making {Monolingual Sentence Embeddings
Multilingual} Using {Knowledge Distillation}.''} \emph{arXiv:2004.09813
{[}Cs{]}}, October. \url{https://arxiv.org/abs/2004.09813}.

\leavevmode\vadjust pre{\hypertarget{ref-shleiferPretrainedSummarizationDistillation2020}{}}%
Shleifer, Sam, and Alexander M. Rush. 2020. {``Pre-Trained
{Summarization Distillation}.''} \emph{arXiv:2010.13002 {[}Cs{]}},
October. \url{https://arxiv.org/abs/2010.13002}.

\leavevmode\vadjust pre{\hypertarget{ref-songMPNetMaskedPermuted}{}}%
Song, Kaitao, Xu Tan, Tao Qin, Jianfeng Lu, and Tie-Yan Liu. n.d.
{``{MPNet}: {Masked} and {Permuted Pre-training} for {Language
Understanding},''} 14.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{paper.bib}

\end{document}